---
title: "CRDC Arrest Rate Analysis: Full sample"
format: html
knitr:
  opts_chunk:
    message: false
    warning: false
    error: false
    echo: false
    fig.path: export/figures/annualdescriptives-
execute:
  echo: false
  warning: false
  message: false
  error: false
params:
  year_full: "2021-22"
  target_name: "y2122"  # Default suffix, will be
---


```{r}
#| label: setup
#| include: false

library(ggplot2)
library(dplyr)
options(scipen = 9)
options(width = 120)

pretty_count <- function(x) {
  x <- round(x, digits = 0)
  x <- prettyNum(x, big.mark = ",")
  return(x)
}


pretty_per <- function (x, ndigit = 1)
{
    if (any(x >= 100) & !all(is.na(x))) {
        message("Values over 100 found, did you mean to use proportions?")
    }
    x <- format(round(x, digits = ndigit + 2) * 100, nsmall = ndigit)
    x <- paste0(x, "%")
    x[x == "NA%"] <- " - "
    x <- trimws(x)
    return(x)
}


# Add colors
color_fill_values <- c("WH" = "#c92d0e", "BL" = "#020684",
                        "AM" = "#ff8f43",
                       "TOTAL" = "#909090ff", "HI" = "#0791b6")
color_fill_values2 <- c("White" = "#c92d0e", "Black" = "#020684",
                        "Total" = "#909090ff", "Hispanic" = "#0791b6",
                        "Amer. Ind." = "#ff8f43")


# Source artifact-reproduction helpers
source("R/funs.R")
source("R/crdc_path.R"); source("R/model_registry.R"); source("R/paper_figures.R")
cv_apply_branding()   # civilytics theme on all figures (+ logo when magick present)

# Year suffix: from tar_render_rep params when present, else the default wave.
suffix <- if (exists("params") && !is.null(params$target_name)) params$target_name else "y2122"
# Load the model-input + district-geo stages for this wave (published artifacts)
sch_ref_arr  <- read_stage_df(sprintf("stages/crdc/model_data_%s.parquet", suffix))
ccd_dist_geo <- read_stage_df(sprintf("stages/crdc/ccd_dist_geo_%s.parquet", suffix))

#sch_referrals_long <- model_data
#sch_universe <- schuniverse
# sch_referrals_long <- params$model_data
# sch_universe <- params$schuniverse
# sch_ref_arr <- left_join(sch_universe,
#                                 sch_referrals_long,
#                                 by = join_by(COMBOKEY))
# rm(sch_referrals_long, sch_universe)

dist_ref_arr <- sch_ref_arr %>%
  group_by(YEAR,LEA_STATE, LEAID, RACE, SEX) %>%
  summarize(ARRESTS = sum(ARRESTS),
            REFERRALS = sum(REFERRALS),
            stu_enroll = sum(stu_enroll),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = ARRESTS / (stu_enroll / 1000),
         referral_rate = REFERRALS / (stu_enroll / 1000)) %>%
  group_by(YEAR, LEA_STATE, LEAID) |>
  mutate(total_enroll = sum(stu_enroll[RACE != "TOTAL" &
                                         SEX != "TOTAL"]))


# Used for 2017-18 to drop NYC but we won't do that for this year
#dist_ref_arr <- dist_ref_arr[dist_ref_arr$LEAID != "3620580",]
state_map <- tigris::states(cb  = TRUE, resolution = "20m") |> tigris::shift_geometry()


```

## Overview for School Year `r params$year_full`

```{r}
#| label: nat_summary
#| echo: false
#| warning: false
#| message: false
dist_ref_arr |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) -> nat_totals

nat_referral_rate <- nat_totals$referrals / (nat_totals$enrollment / 1000)
nat_arrest_rate <- nat_totals$arrests / (nat_totals$enrollment / 1000)
nat_totals
# FIXME: Report the rates also below as Number Needed to Harm (NNH)
```

In `r pretty_count(nat_totals$all_lea_count)` LEAs with data in the `r first(dist_ref_arr$YEAR)`
CRDC data. Of these, `r pretty_count(nat_totals$lea_any_arr)` reported any arrests, or
`r pretty_per(nat_totals$lea_any_arr / nat_totals$all_lea_count)` of all
reporting LEAs.

There were `r pretty_count(nat_totals$enrollment)` students enrolled in the CRDC sample. Of these,
`r pretty_count(nat_totals$referrals)` unique students were referred to law enforcement and `r pretty_count(nat_totals$arrests)`
were arrested. This equates to a rate of `r round(nat_totals$referrals / (nat_totals$enrollment / 1000), 2)` referrals per 1,000 students and
`r round(nat_totals$arrests / (nat_totals$enrollment / 1000), 2)` arrests
per 1,000 students.


### Geographic variation

```{r}
#| label: state_map
#| echo: false
#| warning: false
#| message: false
dist_ref_arr |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEA_STATE) |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000)) |>
    mutate(arrest_nnh = enrollment / arrests) -> state_totals

state_map2 <- inner_join(state_map, state_totals, by = join_by(STUSPS == LEA_STATE))

ggplot(data = state_map2, aes( fill = arrest_rate)) +
  geom_sf() +
  geom_sf_label(aes(label = round(arrest_rate, 1))) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  labs(title = "Arrests per 1,000 for all students by state") +
  theme_void()

rm(state_map2)
```



```{r}
#| label: state_map_arrests
#| echo: false
#| warning: false
#| message: false
dist_ref_arr |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEA_STATE) |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000)) |>
    mutate(arrest_nnh = enrollment / arrests) -> state_totals

state_map2 <- inner_join(state_map, state_totals, by = join_by(STUSPS == LEA_STATE))

ggplot(data = state_map2, aes( fill = arrests)) +
  geom_sf() +
  geom_sf_label(aes(label = civilytics::pretty_count(arrests))) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  labs(title = "Students arrested at school by state") +
  theme_void()


```


```{r}
#| label: state_map_arrestsnnh
#| echo: false
#| warning: false
#| message: false
dist_ref_arr |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEA_STATE) |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000)) |>
    mutate(arrest_nnh = enrollment / arrests) -> state_totals

state_map2 <- inner_join(state_map, state_totals, by = join_by(STUSPS == LEA_STATE))

ggplot(data = state_map2, aes( fill = arrest_nnh)) +
  geom_sf() +
  geom_sf_label(aes(label = civilytics::pretty_count(round(arrest_nnh, 0)))) +
  scale_fill_distiller(palette = "YlOrRd", direction = -1,
  breaks = c(50000, 8000, 3000, 1000, 500),
  trans = "log", labels = scales::comma) +
  labs(title = "Students per arrested student by state", fill = "NNH") +
  theme_void() +
  theme(legend.position.inside = c(0.1, 0.8), legend.position = "inside")

```

### Variation by student group

```{r}
#| label: nat_sg_summary
#| #| echo: false
#| warning: false
#| message: false
dist_ref_arr |>
    ungroup() |>
    filter(RACE != "TOTAL" & SEX != "TOTAL") |>
    group_by(RACE, SEX) |>
    summarize(
        lea_count = n_distinct(LEAID),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000),
            referral_rate = referrals / (enrollment/1000)) -> nat_sg_totals

nat_sg_totals |>
    filter(RACE %in% c("AM", "WH", "HI", "BL"))

```

Another way to characterize the arrest rate is to calculate the Number Needed
to Harm (NNH). This is used in epidemiology to make it easier to compare the
relative risk of low-incident rate health risks and is considered more intuitive
than presenting rates per 1,000.

Below, the NNH for student race/sex combinations is presented in a bar chart -
the lower the NNH, the higher the arrest rate.

```{r}
#| label: sg_summary_arrestNNH
#| fig-width: 9
#| warning: false
#| message: false

ggplot(nat_sg_totals |> filter(RACE %in% c("AM", "WH", "HI", "BL"))) +
  aes(x = crdc_race_recode(RACE), y = 1000 / arrest_rate,
  fill = crdc_race_recode(RACE)) +
    scale_fill_manual(values = color_fill_values2) +
  geom_col() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0))),
              vjust = -0.25) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14) +
  labs(x = "", y = "NNH (Arrests)", fill = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())

```


```{r}
#| label: sg_summary_arrestrate
#| fig-width: 9
#| warning: false
#| message: false

ggplot(nat_sg_totals |> filter(RACE %in% c("AM", "WH", "HI", "BL"))) +
  aes(x = crdc_race_recode(RACE), y = arrest_rate,
  fill = crdc_race_recode(RACE)) +
    scale_fill_manual(values = color_fill_values2) +
  geom_col() +
  geom_text(aes(label = civilytics::pretty_count(round(arrest_rate,2))),
              vjust = -0.25) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 16) +
  labs(title = "2021-22 National Arrest Rates by Select Student Groups",
  x = "", y = "Arrests per 1,000", fill = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())
```

```{r}
#| label: sg_summary_referralNNH
#| fig-width: 9
#| warning: false
#| message: false

ggplot(nat_sg_totals |> filter(RACE %in% c("AM", "WH", "HI", "BL"))) +
  aes(x = crdc_race_recode(RACE), y = 1000 / referral_rate,
  fill = crdc_race_recode(RACE)) +
    scale_fill_manual(values = color_fill_values2) +
  geom_col() +
  geom_text(aes(label = round(1000/referral_rate, 0))) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14) +
  labs(x = "", y = "NNH (Referrals)", fill = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())

```

## Distribution of Arrests by District Size

```{r}
#| label: arrests_by_size
#| echo: false
#| warning: false
#| message: false
thresh <- c(0, 1000, 10000, 20000,Inf)
labs <- c("0-999", "1000-9,999", "10,000-19,999","20,000+")
#
# # Estimate expected rate
 dist_ref_arr %>%
   ungroup() %>%
     filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
   mutate(popcut = cut(stu_enroll,
                       breaks = thresh,
                       labels = labs,
                       include.lowest= TRUE)) %>%
   group_by(popcut) %>%
 mutate(expec_arrest = pbinom(1, stu_enroll, nat_arrest_rate/1000,
                              lower.tail = FALSE)) %>%
   summarize(dists = n(),
             dists_w_arrests = sum(ARRESTS>0),
             expec_distw_arrests = round(sum(expec_arrest),1)) %>%
   mutate(dist_w_arrest_per = pretty_per(dists_w_arrests/dists),
          expect_w_arrest_per = pretty_per(expec_distw_arrests/dists))
```


## Arrest Concentration Plot - National


```{r}
#| label: arrests_concentration_plot
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
total_enroll_national <- sum(dist_ref_arr$total_enroll[
  dist_ref_arr$RACE == "TOTAL" & dist_ref_arr$SEX == "TOTAL"
])
dist_ref_arr %>%
  ungroup %>%
  filter(ARRESTS > 0) %>%
  filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
  arrange(desc(stu_enroll)) %>%
  mutate(total_arr = sum(ARRESTS)) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_enroll_national) %>%
  mutate(dist_per_cum = cumsum(dist_per)) %>%
  ggplot(aes(x = dist_per_cum, y = cum_sum_per)) +
  geom_line() +
  geom_hline(yintercept = 0.5) +
  scale_x_continuous("% of national student population",
                     labels = scales::percent) +
  scale_y_continuous("% of all arrests", labels = scales::percent) +
  theme_minimal(base_size = 12) +
  labs(title = "Share of total arrests by cumulative district enrollment",
       subtitle = "Horizontal line at 50% of all arrests") +
  theme(axis.ticks.x = element_blank(),
        panel.grid = element_blank())
```



```{r}
#| label: arrests_concentration_by_sg
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
dist_ref_arr %>%
    filter(RACE != "TOTAL" & SEX  == "TOTAL") %>%
    filter(RACE %in% c("BL", "WH", "HI", "AM")) |>
  group_by(RACE) %>%
  arrange(RACE, desc(ARRESTS)) %>%
  mutate(total_arr = sum(ARRESTS),
         total_pop = sum(stu_enroll)) %>%
  filter(ARRESTS > 0) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_pop) %>%
  mutate(dist_per_cum = cumsum(dist_per)) %>%
  ggplot(aes(x = dist_per_cum, y = cum_sum_per, color = RACE,
             group = RACE)) +
  geom_line() +
  geom_hline(yintercept = 0.5) +
  scale_x_continuous("% of national student population",
                     labels = scales::percent) +
  scale_y_continuous("% of all arrests", labels = scales::percent) +
 theme_minimal(base_size =12) +
  labs(title = "Share of cumulative arrests by student population",
       subtitle = "Horizontal line at 50% of all arrests",
       color = "Race/Eth.") +
  theme(axis.ticks.x = element_blank(),
        panel.grid = element_blank()) +
  theme(legend.position = "inside",
    legend.position.inside = c(0.8, 0.2))
```


## District Tables


### Counts

#### Overall

```{r}
#| label: dist_top20
#| echo: false
#| warning: false
#| message: false

# Need to get LEA name, state, enrollment

#sch_dir_info <- targets::tar_read("schuniverse_2122")
#sch_dir_info <- schuniverse

# dist_jj_count <- sch_dir_info |>
#     group_by(LEAID) |>
#     summarize(jj_count = sum(JJ == "Yes"))

# Collapse down by district, count JJ

#dist_ccd <- targets::tar_read("ccd_dist_geo_2122")
dist_ccd <- ccd_dist_geo

dist_ccd <- dist_ccd |>
    select(leaid, year, lea_name, enrollment, number_of_schools,
    highest_grade_offered, agency_type, latitude, longitude)
#civilytics::match_test(dist_ccd$leaid, dist_ref_arr$LEAID)

tab_df <- inner_join(dist_ref_arr, dist_ccd, by = join_by(LEAID == leaid))
#tab_df <- inner_join(tab_df, dist_jj_count)

tab_df <- tab_df |>
  rename(crdc_enroll = stu_enroll,
            ccd_enroll = enrollment)

tab_df |> filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    ungroup() |>
    arrange(desc(ARRESTS)) |>
    select(lea_name, LEA_STATE, ccd_enroll, ARRESTS, crdc_enroll, arrest_rate, highest_grade_offered, jj_count) |>

    slice_head(n = 20)


```

#### By group

```{r}
#| label: dist_top20_sorted
#| echo: false
#| warning: false
#| message: false

library(DT)

arrests_by_group <- tab_df |>
  filter(RACE %in% c("WH", "BL", "AM", "HI")) |>  # Keep TOTAL in SEX but not RACE
  group_by(RACE, SEX) |>
  slice_max(order_by = ARRESTS, n = 20) |>
  ungroup() |>
  select(RACE, SEX, lea_name, LEA_STATE, ARRESTS, arrest_rate, ccd_enroll, crdc_enroll)

# Create interactive table
datatable(arrests_by_group,
  filter = list(
    position = "top",
    clear = FALSE,
    list(
      position = "top",
      columns = c(0, 1) # Only RACE and SEX columns filterable
    )
  ),
  options = list(
    pageLength = 20,
    autoWidth = TRUE,
    order = list(list(4, 'desc')) # Sort by ARRESTS column
  )
)
```


### Rates

#### Overall

```{r}
#| label: dist_top20_by_rate
tab_df |> filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    ungroup() |>
    arrange(desc(arrest_rate)) |>
    select(lea_name, LEA_STATE, ccd_enroll, ARRESTS, crdc_enroll, arrest_rate, highest_grade_offered, jj_count) |>
    slice_head(n = 20)

```


#### By Group

```{r}
#| label: dist_top20_rates_sorted
#| echo: false
#| warning: false
#| message: false

arrestrates_by_group <- tab_df |>
  filter(RACE %in% c("WH", "BL", "AM", "HI")) |>  # Keep TOTAL in SEX but not RACE
  group_by(RACE, SEX) |>
  slice_max(order_by = arrest_rate, n = 20) |>
  ungroup() |>
  select(RACE, SEX, lea_name, LEA_STATE, arrest_rate, ARRESTS, ccd_enroll, crdc_enroll)

# Create interactive table
datatable(arrestrates_by_group,
  filter = list(
    position = "top",
    clear = FALSE,
    list(
      position = "top",
      columns = c(0, 1) # Only RACE and SEX columns filterable
    )
  ),
  options = list(
    pageLength = 20,
    autoWidth = TRUE,
    order = list(list(4, 'desc')) # Sort by ARRESTS column
  )
)
```


## Coefficient of Variation by Student Group


```{r}
#| label: cov_plot_prep
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9

# For clarity, let's look at total district size and then cut it and show CV for
# overall district size, and then for Black, White, and Hispanic students

dist_ref_arr %>%
  group_by(LEA_STATE, LEAID) %>%
  filter(stu_enroll > 0) |>
  mutate(total_enroll = sum(stu_enroll[SEX == "TOTAL" & RACE == "TOTAL"]),
         total_arrests = ARRESTS[SEX == "TOTAL" & RACE == "TOTAL"],
         total_referrals = REFERRALS[SEX == "TOTAL" & RACE == "TOTAL"]
  ) -> tabdf

total_enroll_quint <- quantile(tabdf$total_enroll[tabdf$SEX == "TOTAL" & tabdf$RACE == "TOTAL"],
         probs = seq(0, 1,0.2))

quint_labels <- c(
  paste0(format(total_enroll_quint[1], big.mark = ","), " - ",
         format(total_enroll_quint[2] - 1, big.mark = ",")),
  paste0(format(total_enroll_quint[2], big.mark = ","), " - ",
         format(total_enroll_quint[3] - 1, big.mark = ",")),
  paste0(format(total_enroll_quint[3], big.mark = ","), " - ",
         format(total_enroll_quint[4] - 1, big.mark = ",")),
  paste0(format(total_enroll_quint[4], big.mark = ","), " - ",
         format(total_enroll_quint[5] - 1, big.mark = ",")),
  paste0(">", format(total_enroll_quint[5], big.mark = ","))
)
# cut by enrollment quintiles

plotdf <- tabdf %>%
    ungroup() |>
  mutate(size = cut(total_enroll, total_enroll_quint, labels = quint_labels,
      ordered_result = TRUE)
  ) %>%
  #mutate(size = as.numeric(size)) %>%
  filter(SEX == "TOTAL") %>%
  filter(!is.na(size)) %>%
  group_by(size, RACE) %>%
  summarize(count = n(),
            all_enroll = sum(total_enroll),
            mean_arr_rate = mean(arrest_rate, na.rm =TRUE),
            sd_arr_rate = sd(arrest_rate, na.rm = TRUE),
            mean_ref_rate = mean(referral_rate, na.rm = TRUE),
            sd_ref_rate = sd(referral_rate, na.rm = TRUE)) %>%
  mutate(cov_arrests = sd_arr_rate / mean_arr_rate * 100,
         cov_referrals = sd_ref_rate / mean_ref_rate * 100)

plotdf$race_lab <- crdc_race_recode(plotdf$RACE)


```


```{r}
#| label: cov_plot_total
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9

ggplot(plotdf[plotdf$RACE== "TOTAL",],
       aes(x = size, y = cov_arrests/100, color = race_lab, group = race_lab)) +
  geom_smooth(method = "loess", show.legend = FALSE) +
  geom_point(size = 4) +
  scale_color_manual(values = color_fill_values2) +
  scale_y_continuous(labels = scales::label_percent(big.mark = ","),
  limits = c(0, 30)) +
  civilytics::theme_civilytics(font_size = 14) +
  theme(legend.position = "bottom",
  axis.title.x = element_text(margin = margin(t = 15))) +
  labs(x = "Total Enrollment Quintile",
      color = "",
       title = "CV for Arrest Rates by Total Enrollment",
       y = "CV (% scale)")

```

```{r}
#| label: cov_plot
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9



p1 <- ggplot(plotdf[plotdf$RACE %in% c("AM","BL", "WH", "TOTAL", "HI"),],
       aes(x = size, y = cov_arrests/100, color = race_lab, group = race_lab)) +
  geom_smooth(method = "loess", show.legend = FALSE) +
  geom_point(size = 4) +
  scale_color_manual(values = color_fill_values2) +
  scale_y_continuous(labels = scales::label_percent(big.mark = ","),
  limits = c(0, 30)) +
  civilytics::theme_civilytics(font_size = 14) +
  theme(legend.position = "bottom",
  axis.title.x = element_text(margin = margin(t = 15))) +
  labs(x = "Total Enrollment Quintile",
       title = "Arrests",
       y = "CV (% scale)",
       color = "Race / Ethnicity")

  # #theme(legend.position = c(0.8, 0.7))+
  # labs(x = "Total Enrollment Quantile",
  #      y = "CV (% scale)",
  #      title = "Figure 2: Coefficient of Variation for Arrest Rates by District Population",
  #      color = "Race")


p2 <- ggplot(plotdf[plotdf$RACE %in% c("AM", "BL", "WH", "TOTAL", "HI"),],
       aes(x = size, y = cov_referrals/100, color = race_lab, group = race_lab)) +
  geom_smooth(method = "loess", show.legend = FALSE) +
  geom_point(size = 4) +
  scale_color_manual(values = color_fill_values2) +
  scale_y_continuous(labels = scales::label_percent(big.mark = ","), limits = c(0, 25)) +
  civilytics::theme_civilytics(font_size = 14) +
  theme(legend.position = "bottom",
  axis.title.x = element_text(margin = margin(t = 15)))  +
  labs(x = "Total Enrollment Quintile",
       title = "Referrals",
       y = "",
       color = "Race / Ethnicity")

# +
#   labs(x = "Total Enrollment Quantile",
#        y = "CV (% scale)",
#        color = "Race")


library(patchwork)

pw <- p1 + p2


pw + plot_annotation(title = "Coefficient of Variation by District Total Enrollment") +
  plot_layout(guides = "collect") &
  civilytics::theme_civilytics(font_size = 12) +
  theme(legend.position = "bottom")

```
