---
title: "Applied Examples of Prediction Intervals"
format:
  html:
    embed-resources: true
knitr:
  opts_chunk:
    message: false
    warning: false
    error: false
    echo: false
    fig.path: export/figures/appliedexample-
execute:
  warning: false
  message: false
  error: false
  fig-width: 10
  fig-height: 6
---

## Introduction

Using statistical modeling we are able to obtain more precise estimates of the
arrest rate for districts and states. This report provides an example for
how these estimates can be applied to provide better demographic, geographic,
and temporal comparisons.

```{r}
#| label: setup
#| include: false

# Initialize our data
# Get some samples
# Prepare our data for plotting

library(brms)
library(dplyr)
library(ggplot2)
library(ggdist)
library(marginaleffects)
library(tidybayes)
library(ggridges)
library(duckdb)
library(patchwork)

source("R/funs.R")
source("R/crdc_path.R"); source("R/model_registry.R"); source("R/paper_figures.R")
cv_apply_branding()   # civilytics theme on all figures (+ logo when magick present)

# predicted_draws VIEW over the PUBLISHED draws parquet (CRDC_ARTIFACTS selects
# the source: local "export" for the owner, hf:// for a stranger). The existing
# get_prediction_summary(con, ...) calls below work unchanged.
h <- open_draws_view(); con <- h$con

tydata <- read_stage_df("stages/inputs/three_year_data.parquet")
rdata  <- read_stage_df("stages/inputs/recent_data.parquet")


```

To illustrate the usefulness of the approach we look at how model predictions
apply to three specific cases:

1. A district with 0 arrests but many students
2. A district that many arrests
3. A large diverse school district

We compare the predictions to the approximate interval calculated from the
data. In most cases we focus on the most recent wave of CRDC, 2021-22, but
where other years are used we discuss.

```{r}
#| label: get_data

# Get 100 highest arrests
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      arrange(desc(total_arrests)) |>
      ungroup() |>
      slice_head(n= 100) -> bigarr

# Get 100 largest districts
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      arrange(desc(total_enroll)) |>
      ungroup() |>
      slice_head(n= 100) -> big100

# Get largest districts with 0 arrests
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      filter(total_arrests == 0) |>
      arrange(desc(total_enroll)) |>
      ungroup() |>
      slice_head(n= 100) -> big0

# rdata |>
#   filter(YEAR == "21-22") |>
#   mutate(phat = (2+ARRESTS) / (4+stu_enroll)) |> # agcouli approximation ARRESTS +2 and stu_enroll + 4
#   mutate(phat_se = sqrt((phat * (1-phat)/(4+stu_enroll)))) |> # same phat but stu_enroll + 4 here
#   mutate(sd = phat_se * (4 + stu_enroll)) |> # calculate rate standard error to arrest scale
#   mutate(fitted_value = ARRESTS) |>
#   mutate(CVp = 100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))) |> # express CVp as percentage, this is a biased estimator in rare events
#   # agresti-coull confidence interval for arrests using the adjusted arrest rate
#   mutate(ci_upper = (phat + (1.96*phat_se)) * (4 + stu_enroll),
#         ci_lower = (phat - (1.96*phat_se)) * (4 + stu_enroll)) |>
#             # add rule of three correction
#   mutate(ci_upper = ifelse(ARRESTS == 0, 3, ci_upper),  # add 3
#         ci_lower = ifelse(ARRESTS == 0, 0, ci_lower)) |>
#   mutate(ci_lower = ifelse(ci_lower < 0, 0, ci_lower)) |>
#   mutate(model_id = "observed") -> obsv_data



agresti_coull <- function(numerator, denominator, confidence_level = 0.95) {
    adj_star <- qnorm(1 - (1-confidence_level)/2)
    # critical value
    if (numerator > 0) {
        num_star <- numerator + adj_star
        denom_star <- denominator + (2 * adj_star)
        phat <- num_star / denom_star
        phat_se <- sqrt((phat/denom_star) * (1-phat))
        ci_upper <- (phat + (adj_star * phat_se)) * denom_star
        ci_lower <- (phat - (adj_star * phat_se)) * denom_star
        sd <- mean(c(ci_upper, ci_lower)) / confidence_level
        sd <- phat_se * denom_star
    } else {

        ci_upper <- denominator*(-log(1-confidence_level)/denominator)
        ci_lower <- 0
        sd <- mean(c(ci_upper, ci_lower)) / confidence_level
        phat_se <- 0
        phat <- mean(c(ci_upper, ci_lower))
    }
    return(c(ci_upper, ci_lower, sd, phat_se, phat))
}


```

## Large district with 0 arrests

```{r}
#| label: zero_arrest_rates
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

focal_dist <- sample(unique(big0$LEAID), 1)
# focal_dist <- "4826580"
# focal_dist <- "1301110"
focal_dist <- "3412690"

obsv_focal <- rdata |> filter(LEAID == focal_dist)
# Make plot title
unique(obsv_focal$LEA_NAME)
unique(obsv_focal$LEA_STATE)
dist_name_for_title <- paste0(tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))),
                            " in ", state.name[state.abb==unique(obsv_focal$LEA_STATE)])


dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", YEAR = "21-22", LEAID = focal_dist)

dist_draws <- get_prediction_draws(con,  YEAR = "21-22", LEAID = focal_dist)


obsv_plot <- obsv_focal |> ungroup() |>
    summarize(arrests = sum(ARRESTS),
             enroll = sum(stu_enroll)) |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Frequentist")

model_plot <- dist_draws |>
    group_by(model_id, draw_id) |>
    summarize(arrestst = sum(pred)) |>
    group_by(model_id) |>
    summarize(arrests = median(arrestst),
            ci_upper = quantile(arrestst, 1 - (1 - 0.95)/2),
            ci_lower = quantile(arrestst, (1 - 0.95)/2),
            enroll = obsv_plot$enroll[1]) |>
    rename(type = model_id)


# let's wrap model types in two dimensions

add_model_time <- function(x) {
    y <- rep("One year", length(x))
    y[x %in% c("nat_m3_mod", "nat_m4_mod", "nat_m5_mod")] <- "Three year"
    y[x %in% c("sg_m3_mod", "sg_m4_mod", "sg_m5_mod")] <- "Three year"
    return(y)
}

add_model_form <- function(x) {
    y <- rep("Stratified", length(x))
    y[startsWith(x, "nat")] <- "Unified"
    return(y)
}

add_model_cov <- function(x) {
  y <- rep("Baseline", length(x))
  y[x %in% c("sg_m2_mod", "nat_m4_mod", "nat_m5_mod", "nat_m2_mod", "sg_m4_mod", "sg_m5_mod" )] <- "Covariate"
  return(y)
}


print_model_name <- function(x, lbreak = FALSE) {
    y <- rep("Stratified", length(x))
    y[startsWith(x, "nat")] <- "Unified"
    x <- gsub("_mod", "", x)
    x <- gsub("nat_m", "", x)
    x <- gsub("sg_m", "", x)
    if (lbreak) {
      y <- paste0(y, "\nModel:", x)
    }  else {
      y <- paste0(y, " Model:", x)

    }

    return(y)
}


model_palette <- c('Modeled' ="#000a9bff", 'Frequentist' = "#858585bb")


plotdf <- bind_rows(obsv_plot %>%
  tidyr::crossing(model_id = unique(model_plot$type)),
  model_plot |> mutate(model_id = type)) |>
    mutate(type = ifelse(type == "Frequentist", type, "Modeled"))

ggplot(plotdf, aes(x = model_id, y = arrests,
                    ymin = ci_upper, ymax = ci_lower, color = type,
                    group = type)) +
    geom_pointrange(position = position_dodge(width = 0.5)) +
    facet_wrap(add_model_cov(model_id)~add_model_time(model_id), scales = "free_y") +
    scale_x_discrete("Model", limits = rev,
      labels = function(x) print_model_name(x, lbreak = TRUE)) +
    coord_flip() +
    scale_color_manual(values = model_palette, guide =
              guide_legend(override.aes = list(linetype = c(0, 0)))) +
    labs(title = stringr::str_wrap(paste0("Predicted arrests 95% interval for ", dist_name_for_title), 90),
    y = "Predicted arrests", x = "Model type",
    color = "",
    subtitle = "Observed arrests are 0, frequentist interval given by the rule of 3.") +
    theme_ridges(grid = TRUE, font_size = 14) +
    theme(legend.position = "bottom",
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
    axis.text.y = element_text(angle = 90, hjust = 0.25))

```

In our first example, we look at how Bayesian prediction models differ from
frequentist intervals in the case of no observed arrests. The Paterson New Jersey
school district has `r pretty_count(obsv_focal$total_enroll[1])`
in our dataset - so having 0 arrests is suspicious.

In the case of no arrests, the observed interval is constant - from 0 to 3, and shown in gray
alongside the Bayesian prediction interval in blue for
each of our 10 models. The models are sorted into single-year specifications
on the left and three year specifications on the right. We see that the single-year
models, except for stratfiied model 1, have much narrower intervals than the
frequentist interval - confirming the advantage of the modeling methodology for increasing
precision.

However, these models still suggest that 0 arrests is the median estimate (the point).
Here the three year models disagree - the intervals are all wider
than the frequentist interval and the median estimated arrests is greater than 0.
In particular models without covariates (Unifed Model 3
and Stratified Model 3) suggest that more than 10 arrests are most liklely. The
reason becomes clear when we look at the multiple years of data and see that
in 15-16 Paterson had 47 arrests reported and in 17-18 it had 11 arrests reported.

```{r}
tydata[tydata$LEAID == focal_dist,] |> group_by(YEAR) |>
  summarize(students = sum(stu_enroll), referrals = sum(REFERRALS), arrests = sum(ARRESTS))
```

The three-year
models are incorporating this information, which expands their intervals. The
utility of the modeling approach becomes more clear when we look at the distribution
of the expected arrests for Paterson shown in the next figure. Here,
every draw of the model is plotted - for clarity, we show
only the covariate specification with level 1 covariates because
there are almost no meaningful differences from the version with level 1 and 2
covariates.

Here we see that the one-year models agree that 0
arrests is the most likely number for Paterson. The stratified model is less
convinced than the than unified model, giving a near equal chance (47.2%) of greater
than 0 arrests as it does to 0 arrests. However, adding a covariate brings the
stratified model closer to the unified model, increasing the chance of 0 arrests
to 76.4%. Interestingly, the covariate has much less effect on the unified model
only increasing the chance of 0 arrests from 82.2 to 84%. Even though these
models are fairly confident there are 0 arrests, they cannot rule out 1 arrest
with 95% confidence - which illustrates the intervals in the previous figure.

Things are more interesting with the three year models. Without covariates,
these models are very confident there are more than 0 arrests in Paterson, but
uncertain about exactly how many, with a relatively flat distribution of likely
arrest counts ranging from 8 to 15, and a substantial tail at 16 or more arrests
(the arrest counts are trunacated at 16 for visual clarity). These models
confidently rule out 0 arrests and cannot confidently rule out as many as 16
arrests.

Adding covariates greatly changes the picture, shifting the likely number of
arrests downward, with  3-4 arrests now being the most likely. These models
still rule out 0 arrests with 95% probability, but cannot rule out only 1 arrest.
The three year models all differ greatly from the one-year models, which we
expect incorporating the prior information about Paterson into the model. However,
we also see that the impact of the covariates is substantial, because  in 21-22
Paterson reported 0 referrals to law enforcement, so this absence of referrals
provides additional information to the model that the prediction should be
pushed toward 0.

Taken together, these models give us three ways to think about the likelihood of
0 arrests in Paterson New Jersey: 1) strongly align with the reported data and
greatly improve precision over the frequentist interval (one year models 1 and 2),
strongly weigh the historic pattern of arrests in Paterson and expect 10 or more arrests (model 3), or
find a middle ground between these two by taking into account the additional
information we have from reported referrals (model 4). With Bayesian intervals we
can choose the method that is most appropriate to our application.

```{r}
#| label: zero_arrest_counts_draws
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

plot_draws <- dist_draws |> group_by(model_id, draw_id) |>
              summarize(pred = sum(pred))
ndraws <- 500 # draws per model per group, 500 * 8

plot_draws <- plot_draws |>
              mutate(pred = ifelse(pred > 16, 16, pred))

p1 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
            filter(add_model_cov(model_id) == "Baseline") |>
            filter(add_model_time(model_id) == "Three year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 3.25, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.1))) +
    scale_x_continuous(breaks = c(5:16), limits = c(4, 17),
    expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "Three year models", subtitle = "Baseline", x ="", y = "") +
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))


p2 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
  filter(add_model_cov(model_id) == "Baseline") |>
     filter(add_model_time(model_id) == "One year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 4, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.2)), name = "Model type") +
    scale_x_continuous(breaks = c(0:8), limits = c(-0.5, 8),
    expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "One year models", subtitle = "Baseline", x = "", y = "Model type")+
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))

p3 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
            filter(add_model_cov(model_id) == "Covariate") |>
            filter(add_model_time(model_id) == "Three year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 3.25, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.1))) +
    scale_x_continuous(breaks = c(0:12), limits = c(-0.5, 13),
    expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "Three year models", subtitle = "Covariates",
            x = "Predicted Arrests", y ="") +
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))


p4 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
      filter(add_model_cov(model_id) == "Covariate") |>
     filter(add_model_time(model_id) == "One year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 4, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.2))) +
    scale_x_continuous(breaks = c(0:8), limits = c(-0.5, 8),
    expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "One year models", subtitle = "Covariates",
     y = "", x = "Predicted Arrests")+
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))

p2+ p1 + p4 + p3  + plot_annotation(
  title = paste0("All model draws for ", dist_name_for_title),
  caption = "Observed arrests in 2021-22 are 0.\nArrests are top-coded at 12 or more for visual clarity.\nValues with fewer than 5% predicted likelihood are not labeled.",
  subtitle = 'Frequency of predicted arrests from 500 draws of posterior for each model'
)

```

```{r}

dist_name_for_title
knitr::kable(obsv_focal |> select(-LEA_NAME))

```

## A district with many arrests



```{r}
#| label: high_arrest_count_intervals
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

focal_dist <- sample(unique(bigarr$LEAID), 1)

#focal_dist <- "3412690"
focal_dist <- "0102370"
obsv_focal <- rdata |> filter(LEAID == focal_dist)
# Make plot title
unique(obsv_focal$LEA_NAME)
unique(obsv_focal$LEA_STATE)
dist_name_for_title <- paste0(tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))),
                            " in ", state.name[state.abb==unique(obsv_focal$LEA_STATE)])

dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", YEAR = "21-22", LEAID = focal_dist)

dist_draws <- get_prediction_draws(con,  YEAR = "21-22", LEAID = focal_dist)


obsv_plot <- obsv_focal |> ungroup() |>
    summarize(arrests = sum(ARRESTS),
             enroll = sum(stu_enroll)) |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Frequentist")

model_plot <- dist_draws |>
    group_by(model_id, draw_id) |>
    summarize(arrestst = sum(pred)) |>
    group_by(model_id) |>
    summarize(arrests = median(arrestst),
            ci_upper = quantile(arrestst, 1 - (1 - 0.95)/2),
            ci_lower = quantile(arrestst, (1 - 0.95)/2),
            enroll = obsv_plot$enroll[1]) |>
    rename(type = model_id)




plotdf <- bind_rows(obsv_plot %>%
  tidyr::crossing(model_id = unique(model_plot$type)),
  model_plot |> mutate(model_id = type)) |>
    mutate(type = ifelse(type == "Frequentist", type, "Modeled"))

ggplot(plotdf, aes(x = model_id, y = arrests,
                    ymin = ci_upper, ymax = ci_lower, color = type,
                    group = type)) +
    geom_pointrange(position = position_dodge(width = 0.5)) +
    facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top",
     scales = "free_y") +
    scale_x_discrete("Model", limits = rev,
                labels = function(x) print_model_name(x, lbreak = TRUE)) +
      scale_color_manual(values = model_palette, guide =
              guide_legend(override.aes = list(linetype = c(0, 0)))) +
    coord_flip() +
    labs(title = stringr::str_wrap(paste0("Predicted arrests 95% interval for ", dist_name_for_title), 120),
    y = "Predicted arrests", x = "Model type",
    subtitle = "Frequentist interval shown in gray.",
    color = "") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom",
    axis.text.y = element_text(angle = 90))



```


Now we turn our attention to the other end of the scale - how our models
perform when there are many observed arrests. Consider the case of Mobile County
Alabama, a district with `r pretty_count(sum(obsv_focal$stu_enroll))` students
which reported `r pretty_count(sum(obsv_focal$ARRESTS))` arrests.
In all models except Unified models 4 and 5,  the observed interval, in gray,
is narrower than the model intervals, and in no case are the intervals in
complete disagreement. Again, we see the biggest difference among models comes
from either all of the one-year models (left side) which closely track the
frequentist interval, but are wide,
the three year models without covariates (top right) which tend toward more arrests
than the frequentist interval,
and the three year models with covariates (bottom right) which tend toward fewer
arrests.

However, as we saw above, we can go beyond intervals and look at the distribution
of likely arrests from the models and see how it compares to the frequentist interval.

We see a more interesting story when we plot the distribution of the draws from
each model as density ridges. Here, the 95% Bayesian highest posterior density
distribution is represneted in color, with the full posterior distribution shown
in gray. The frequentist interval is represented as the
purple point interval. We see the one-year models on the left are
very similar with the baseline models nearly perfectly recreating the frequentist
interval and the stratified models having a wider tail (gray density region).

None of the three-year models align with the one-year models. The baseline models (top right)
are shifted to the right (higher expected arrests) and the covariate models (bottom right)
 are shifted to the left (fewer expected arrests). In all cases the HPD from the
 Bayesian models still intersects with the frequentist interval. However,
 as we will look into with our next example, the difference in the intervals is
 not the same as the interval of the difference between them. Bayesian analysis
 allows us to work out the probability of differences exactly.

```{r}
#| label: high_arrest_count_draws_hpdplot
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10


plot_draws <- dist_draws |>
  filter(!(grepl("_m5", model_id))) |>
  group_by(model_id, draw_id) |>
              summarize(pred = sum(pred)) |>
  group_by(model_id) |>
  arrange(model_id, desc(pred)) |>
  mutate( ordvar = 1:n()) |>
  mutate(fillvar = abs(ordvar - 250))



# Starting with your filtered and summarized data
result <- dist_draws |>
  filter(!(grepl("_m5", model_id))) |>
  group_by(model_id, draw_id) |>
  summarize(pred = sum(pred), .groups = "drop") |>
  group_by(model_id) |>
  # Add rank (descending by pred value)
  mutate(
    rank = rank(-pred, ties.method = "min")
  )

# Calculate HPD intervals for each model_id using base R
# First, get unique model_ids
model_ids <- unique(result$model_id)

# Initialize vectors to store HPD bounds
lower_hpd_vec <- numeric(nrow(result))
upper_hpd_vec <- numeric(nrow(result))
within_hpd_vec <- logical(nrow(result))

# Loop through each model_id to calculate HPD
for (i in seq_along(model_ids)) {
  current_model <- model_ids[i]
  model_rows <- which(result$model_id == current_model)
  model_preds <- result$pred[model_rows]

  # Sort predictions for this model
  sorted_preds <- sort(model_preds)

  # Find the narrowest window containing 95% of values
  n_values <- length(sorted_preds)
  window_size <- ceiling(0.95 * n_values)

  # Calculate all possible windows
  window_widths <- numeric(n_values - window_size + 1)
  window_starts <- numeric(n_values - window_size + 1)
  window_ends <- numeric(n_values - window_size + 1)

  for (j in 1:(n_values - window_size + 1)) {
    window_starts[j] <- sorted_preds[j]
    window_ends[j] <- sorted_preds[j + window_size - 1]
    window_widths[j] <- window_ends[j] - window_starts[j]
  }

  # Find the narrowest window
  min_width_idx <- which.min(window_widths)
  lower_hpd <- window_starts[min_width_idx]
  upper_hpd <- window_ends[min_width_idx]

  # Assign HPD bounds and within_hpd flag to all rows for this model
  lower_hpd_vec[model_rows] <- lower_hpd
  upper_hpd_vec[model_rows] <- upper_hpd
  within_hpd_vec[model_rows] <- model_preds >= lower_hpd & model_preds <= upper_hpd
}

# Add HPD information to the result data frame
result$lower_hpd <- lower_hpd_vec
result$upper_hpd <- upper_hpd_vec
result$within_hpd <- within_hpd_vec

# Arrange by model_id and rank
result <- result |>
  arrange(model_id, rank)

# To filter only rows within HPD
hpd_only <- result |> filter(within_hpd)


ggplot(result, aes(x = pred, y = model_id, group = model_id)) +
  geom_density_ridges(scale = 0.95, rel_min_height = 0.01) +
    geom_density_ridges_gradient(data = result |> filter(within_hpd),
    aes(fill = after_stat(x)), scale = 0.95, gradient_lwd = 1.0, rel_min_height = 0.01) +
    geom_pointrange(data = obsv_plot %>%
  tidyr::crossing(model_id = unique(plot_draws$model_id)),
  aes(y = model_id, x = arrests, xmin = ci_lower, fill = NULL, xmax = ci_upper), color = I("#5a0052ff"),
  position = position_nudge(y = 0.15)) +
     scale_fill_viridis_c(name = "Arrests", option = "C", guide = guide_none()) +
        scale_y_discrete("Model", limits = rev,
                labels = function(x) print_model_name(x, lbreak = TRUE),
                expand = expansion(mult = c(0.1, 0.15))) +
    facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    labs(title = stringr::str_wrap(paste0("Predicted arrests probability density interval for ", dist_name_for_title), 120),
    x = "Predicted arrests", y = "Model type",
    color = "Model or observed",
    subtitle = stringr::str_wrap("Frequentist rate and interval shown in purple. Colored fill represents the 95% highest posterior density for each model, shaded to emphasize number of arrests.", 70)) +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom",
    panel.grid.major.y = element_line(color = "gray50"),
    axis.text.y = element_text(angle = 90, hjust = -0.5))


```

```{r}
dist_name_for_title
knitr::kable(obsv_focal |> select(-LEA_NAME))

# Three year data summary
tydata[tydata$LEAID == focal_dist,] |> group_by(YEAR) |>
  summarize(students = sum(stu_enroll), referrals = sum(REFERRALS), arrests = sum(ARRESTS))

```

## A large and diverse school district

One of the major motivations for modeling arrest rates is to improve our ability
to quantify demographic, geographic, and temporal differences in arrest rates
in the face of uncertainty posed by the sparse and rare nature of arrests. We
take the case of Clark County, NV where we model the arrest rates of three
demographic groups for comparison, White, Hispanic, and Black male students.
In the figure below the Bayesian posterior density is plotted against the frequentist
approximate interval (the point line). We can see a frequentist analysis allows us
to confirm that Black male students have a decidedly higher arrest rate than
Hispanic and White male students, but does not allow us to distinguish between
White and Hispanic students (the intervals overlap).

In all cases, our Bayesian models agree that Black students have a meaningfully
different arrest rate, though without covariates, unified models 1 and 3 (bottom row, top 2 panels)
believe the rate should be much lower than the frequentist estimate. In all cases,
we observe substantial overlap in the distribution of Hispanic and White students, though
we see in some cases, like the covariate three year model (bottom right panel),
there is more distinction between the two distributions.

However, the difference in these intervals is not the interval of the difference.
Fortunately, with our Bayesian posterior samples we can directly compute the distribution
of the difference between White and Hispanic students and plot it just like we have
plotted the predicted values here.

```{r}
#| label: group_comparison
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

focal_dist <- "2400480"
focal_dist <- "3200060"
#focal_dist <- sample(unique(big100$LEAID), 1)
obsv_focal <- rdata |> filter(LEAID == focal_dist)
# Make plot title
unique(obsv_focal$LEA_NAME)
unique(obsv_focal$LEA_STATE)
dist_name_for_title <- paste0(tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))),
                            " in ", state.name[state.abb==unique(obsv_focal$LEA_STATE)])

dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", YEAR = "21-22", LEAID = focal_dist)

dist_draws <- get_prediction_draws(con,  YEAR = "21-22", LEAID = focal_dist)


obsv_plot <- obsv_focal |> ungroup() |>
    filter(SEX == "M") |>
    filter(RACE %in% c("BL", "WH", "HI")) |>
    group_by(RACE) |>
    summarize(arrests = sum(ARRESTS),
             enroll = sum(stu_enroll)) |>
    rowwise() |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "observed")


model_plot <- dist_draws |>
    group_by(model_id) |>
    summarize(arrests = median(pred),
            ci_upper = quantile(pred, 1 - (1 - 0.95)/2),
            ci_lower = quantile(pred, (1 - 0.95)/2),
            enroll = obsv_plot$enroll[1]) |>
    rename(type = model_id)

plot_draws <- dist_draws |>
    filter(!grepl("_m5", model_id)) |>
    filter(SEX == "M") |>
    filter(RACE %in% c("BL", "WH", "HI")) |>
    mutate(enroll = ifelse(RACE == "BL", obsv_plot$enroll[obsv_plot$RACE == "BL"],
                            ifelse(RACE == "HI", obsv_plot$enroll[obsv_plot$RACE == "HI"],
                            obsv_plot$enroll[obsv_plot$RACE == "WH"])))

  group_palette <- c("WH" = "#09d2ecc2", "BL" = "#ab00a9f1", "HI" = "#00ab17f1")


ggplot(plot_draws, aes(x = (pred/(enroll/1000)), color= RACE, fill = RACE, y = model_id)) +
    geom_density_ridges(scale = 0.9, rel_min_height = 0.01, alpha = 1/5, bandwidth = 0.1) +
    geom_pointrange(data = obsv_plot %>%
    filter(RACE == "HI") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (arrests/(enroll/1000)), xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
    geom_pointrange(data = obsv_plot %>%
    filter(RACE == "BL")|>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (arrests/(enroll/1000)), xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.35)) +
    geom_pointrange(data = obsv_plot %>%
    filter(RACE == "WH")|>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (arrests/(enroll/1000)), xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.15)) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(c(0,0))) +
    facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    guides(color = guide_none()) +
    scale_color_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    scale_fill_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title), 120),
    x = "Arrest rate per 1,000", y = "",
    fill = "Student race:",
    subtitle = "Male students. Frequentist interval shown as point range.") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom",
    axis.text.y = element_text(angle = 90, hjust = -0.5))


```


```{r}
dist_name_for_title
knitr::kable(obsv_focal |> select(-LEA_NAME))

```

Our next figure shows the predicted arrest distributions on the top and then
the distribution of the differnces in orange below. This figure illustrates how
looking at the differences in the predicted arrest distribution can be misleading
in understanding the modeled differnce in arrest rates. Our models see a minimum
84% and maximum 100% probability that Hispanic male student are arrested at a
higher rate than White students, and that this arrest rate difference can likely
be as a high 1 additional arrest per 1,000 students.

```{r}
#| label: clarkdifferencescomposite
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

diff_a <- plot_draws |>
   filter(!grepl("_m5", model_id)) |>
   group_by(model_id, draw_id) |>
  summarize(group_b = pred[RACE == "BL"]/enroll[RACE=="BL"],
            group_w = pred[RACE == "WH"]/enroll[RACE=="WH"]) |>
    mutate(diffv = group_b - group_w)

diff_b <- plot_draws |>
     filter(!grepl("_m5", model_id)) |>
  group_by(model_id, draw_id) |>
  summarize(group_h = pred[RACE == "HI"]/enroll[RACE=="HI"],
            group_w = pred[RACE == "WH"]/enroll[RACE=="WH"]) |>
    mutate(diffv = group_h - group_w)


# ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id)) +
#     geom_density_ridges(scale = 0.9, alpha = 3/5, bandwidth = 0.1)+
#     geom_vline(xintercept = 0)


annotate_df <- diff_b |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 1) |>
  mutate(diff_per = diffcount / total)

plot_draws <- plot_draws |> filter(!grepl("_m5", model_id))

p1 <- ggplot(diff_b, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-0.5, to = 1.5, scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df,
        size = 2.5,
        position = position_nudge(y = 0.4, x = 0),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +

    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 12) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Model estimated difference (", "\u0394", ") between Hispanic and White student arrest rates in ", dist_name_for_title), 90),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    ) +
      scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE), expand = expansion(c(0,0))) +
      facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    theme(legend.position = "bottom",
    axis.text.y = element_text(angle = 90, hjust = 0.25))


p2 <- ggplot(plot_draws |>
  filter(RACE != "BL"), aes(x = (pred/(enroll/1000)),
      color= RACE, fill = RACE, y = model_id)) +
    geom_density_ridges(scale = 0.9, rel_min_height = 0.01, alpha = 1/5, bandwidth = 0.1) +
    geom_pointrange(data = obsv_plot %>%
    filter(RACE == "HI") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (arrests/(enroll/1000)), xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
   geom_pointrange(data = obsv_plot %>%
    filter(RACE == "WH")|>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (arrests/(enroll/1000)), xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.15)) +

    scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE), expand = expansion(c(0,0))) +
  facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    guides(color = guide_none()) +
    scale_color_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    scale_fill_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title), 90),
    x = "Arrest rate per 1,000", y = "",
    fill = "Student race:",
    subtitle = "Male students only. Frequentist interval shown by point interval.") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 12) +
    theme(legend.position = "bottom",
    axis.text.y = element_text(angle = 90, hjust = 0.25))



p2 / p1


```


## Multicomparison

Compare Baltimore City to Baltimore County
Baltimore City to itself last year
Broward White Female to Hispanic female



```{r}

focal_dist <- c("2400090", "2400120")

obsv_focal <- tydata |> filter(LEAID == focal_dist) |>
  group_by(YEAR, LEA_STATE, LEAID, LEA_NAME) |>
  summarize(enroll = sum(stu_enroll),
            enroll2 = first(total_enroll),
            referrals = sum(REFERRALS),
            arrests  = sum(ARRESTS)
            )


dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", LEAID = focal_dist)

dist_test <- dist_test |> filter(add_model_time(model_id) == "three year")
dist_draws <- get_prediction_draws(con,  LEAID = focal_dist)
dist_draws <- inner_join(dist_draws, obsv_focal |> select(YEAR:LEA_NAME, enroll, arrests))
dist_test <- inner_join(dist_test, obsv_focal |> select(YEAR:LEA_NAME, enroll, arrests))

plot_draws <- dist_draws |> group_by(model_id, LEAID, YEAR, draw_id) |>
              summarize(pred = sum(pred))

plot_draws <- inner_join(plot_draws, obsv_focal |> select(LEA_NAME, LEAID, YEAR, enroll, arrests))

dist_name_for_title <- gsub(" Public Schools", "",
                      tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))))



```

Here we apply the same approach as above but to compare arrest rates from two
different school districts - Baltimore City Public Schools (orange) and Baltimore County
Public Schools (blue). Here we see a curious pattern - if we look at our one
year models, the frequentist interval and Bayesian intervals strongly agree that
Baltimore City has a higher arrest rate than Baltimore County, with the most
likely arrest rate in Baltimore County being 0. Covariates make little difference
in these estimates.

The three year models have a very different picture - the Bayesian models see
Baltimore County as having not just a higher arrest rate, but a much higher
arrest rate than Baltimore City. The Baltimore City estimates are relatively
unchanged, though covariates seem to shift the distribution slightly closer to
0 (unified and stratified models 4).

When we turn to our direct estimate of the differences, we see this pattern play
out - the one year models are 100% certain in all specifications that Baltimore
City has a higher arrest rate than Baltimore County, and that difference is
most likely narrowly between just greater than 0 and 1. In contract, the 3 year
models are equally sure that Baltimore City has no chance of a higher arrest
rate than Baltimore County, and in fact, that it is most likely that Baltimore
County's arrest rate is 1-2 arrests per 1,000 greater than Baltimore City.
When we pool all of our models together, we are left with a 50% chance one
of the Baltimore's has a greater arrest rate than the other (inset, bottom left).

```{r}
#| label: baltimorecityandcountydifferences
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

obsv_plot <-  obsv_focal |>
  rowwise() |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Observed")


diff_a <- plot_draws |> filter(YEAR == "21-22") |>
  filter(!grepl("_m5", model_id)) |>
  group_by(model_id, draw_id) |>
  summarize(group_city = pred[LEAID == "2400090"]/enroll[LEAID=="2400090"],
            group_county = pred[LEAID == "2400120"]/enroll[LEAID=="2400120"]) |>
    mutate(diffv = group_city - group_county)

annotate_df <- diff_a |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 1) |>
  mutate(diff_per = diffcount / total)

plot_draws <-
  plot_draws |> filter(YEAR == "21-22") |>  filter(!grepl("_m5", model_id))

p2 <- ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-3, to = 1.5,  rel_min_height = 0.01, scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df |> filter(model_id %in% c("sg_m2_mod", "sg_m4_mod")),
        size = 3,
        position = position_nudge(y = 0.4, x = 0),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE)) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Model estimated difference (", "\u0394", ") between Baltimore City and Baltimore County student arrest rates in 21-22"), 120),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    ) +
      facet_wrap(~add_model_time(model_id), scales = "free_y") +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, size = 10, hjust = -0.5))

p1 <- ggplot(plot_draws,
    aes(x = (pred/(enroll/1000)),
      color= LEA_NAME, fill = LEA_NAME, y = model_id)) +
    geom_density_ridges(scale = 0.9, rel_min_height = 0.01, alpha = 2/5, bandwidth = 0.1) +
    geom_pointrange(data = obsv_plot |> filter(YEAR == "21-22") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (arrests/(enroll/1000)),
      xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = LEA_NAME, alpha = 1),
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(0,0)) +
    facet_wrap(~add_model_time(model_id), scales = "free_y") +
    guides(color = guide_none()) +
    scale_color_manual(values = c("#D55E0050", "#0072B250"), labels = function(x) stringr::str_wrap(tools::toTitleCase(tolower(x)), 80)) +
    scale_fill_manual(values = c("#D55E0050", "#0072B250"), labels = function(x) stringr::str_wrap(tools::toTitleCase(tolower(x)), 80)) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title[1], " and ", dist_name_for_title[2]), 120),
    x = "Arrest rate per 1,000", y = "Model type", fill = "", color = "") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, size = 10, hjust = -0.5))



diff_b <- diff_a |> mutate(model_id = "All Models Pooled") |>  mutate(model_id == as.character(model_id))
annotate_df <- diff_b |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 1) |>
  mutate(diff_per = diffcount / total)

p3 <- ggplot(diff_b,
    aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-3, to = 1.5, rel_min_height = 0.01, scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_text(data = annotate_df,
        size = 2.5,
        position = position_nudge(y = 1.2, x = -2.25),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 10) +
    scale_y_discrete(expand = expansion(0, 0)) +
    labs(x = "Arrest rate per 1,000",
    title = "All models pooled",
    y = "",
    ) +
    theme(legend.position = "bottom",
          axis.text.y = element_blank())


p1 / (p2 + inset_element(p3, left = 0, bottom = 0.2, top = 0.9, right = 0.24))





```

To better understand what is going on we can use our models to make temporal
comparisons in Baltimore County - what does the trend look like.

```{r}
#| label: baltimorecountytrend
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

# Temporal for Baltimore County

plot_draws <- dist_draws |> group_by(model_id, LEAID, YEAR, draw_id) |>
              summarize(pred = sum(pred))

plot_draws <- inner_join(plot_draws, obsv_focal |> select(LEA_NAME, LEAID, YEAR, enroll, arrests))

plot_draws <- plot_draws |> filter(add_model_time(model_id) == "Three year") |>
  filter(LEAID == "2400120") |>
  filter(!grepl("_m5", model_id))


dist_name_for_title <- gsub(" Public Schools", "",
                      tools::toTitleCase(tolower(unique(plot_draws$LEA_NAME))))


p1 <- ggplot(plot_draws ,
    aes(x = (pred/(enroll/1000)),
      color= I("black"), fill = YEAR, y = YEAR)) +
    geom_density_ridges(scale = 2.5, rel_min_height = 0.01, alpha = 9/10, bandwidth = 0.05) +

    geom_point(data = obsv_plot |> filter(LEAID == "2400120") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = YEAR, x = (arrests/(enroll/1000)),
   color = I("black")), size = 6, alpha = 1,
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
     geom_pointrange(data = obsv_plot |> filter(LEAID == "2400120") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = YEAR, x = (arrests/(enroll/1000)),
      xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = YEAR), size = 1, alpha = 1,
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
    scale_y_discrete() +
    facet_wrap(~print_model_name(model_id)) +
    guides(color = guide_none()) +
    guides(fill = guide_none()) +
    scale_color_brewer(palette = "Blues", direction = 1) +
    scale_fill_brewer(palette = "Blues", direction = 1) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title[1], " over time"), 120),
  subtitle = "Observed value shown by point",
    x = "Arrest rate per 1,000", y ="", fill = "", color = "") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom")


diff_a <-plot_draws |> filter(add_model_time(model_id) == "Three year") |>
  filter(LEAID == "2400120") |>
  group_by(model_id, draw_id) |>
  summarize(group_t = pred[YEAR == "21-22"]/enroll[YEAR=="21-22"],
            groupprev = pred[YEAR == "17-18"]/enroll[YEAR == "17-18"]) |>
    mutate(diffv = group_t - groupprev)

annotate_df <- diff_a |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv < 0),
            diffv = -5.5) |>
  mutate(diff_per = diffcount / total)

p2 <- ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient( rel_min_height = 0.01,
    from = -16, to = 1,
    scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_text(data = annotate_df |> filter(model_id == "sg_m4_mod"),
        size = 3,
        position = position_nudge(y = 0.7, x = 0),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "<0): ", pretty_per(diff_per)))) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE)) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Model estimated difference (", "\u0394", ") between ", dist_name_for_title, " and 2021-22 and 2017-18"), 120),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    ) +
      facet_wrap(~add_model_time(model_id), scales = "free_y") +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = -0.5))


p3 <- ggplot(obsv_plot, aes(x = YEAR, y = enroll, group = LEA_NAME, color = LEA_NAME)) +
  geom_point() + geom_line() +
  scale_color_manual(values = c("#d55c00ff", "#0071b2f6"), labels = function(x) stringr::str_wrap(tools::toTitleCase(tolower(x)), 50)) +
  scale_y_continuous("Enrollment", labels = scales::comma, limits = c(0,NA)) +
  guides(color = guide_legend(nrow = 2, override.aes = list(linetype = c(0, 0)))) +

  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE) +
  labs(color = "", x = "CRDC Year", title = "Reported enrollment") +
  theme(legend.position = "bottom")


p1 / (p2 + inset_element(p3, left = 0.67, bottom = 0.2, top = 0.8, right = 1))



```




```{r}
# Small groups
focal_dist <- "1200180"
obsv_focal <- rdata |> filter(LEAID == focal_dist) |>
  filter(SEX == "F") |>
  filter(RACE %in% c("WH", "HI"))

dist_draws <- get_prediction_draws(con,  LEAID = focal_dist, YEAR = "21-22",
                            SEX = "F", RACE = c("WH", "HI"))

plot_draws <- inner_join(dist_draws, obsv_focal |> select(YEAR:RACE, stu_enroll, ARRESTS)) |>
  filter(!grepl("_m5", model_id))

dist_name_for_title <- gsub(" Public Schools", "",
                      tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))))

obsv_plot <- obsv_focal |>
  rowwise() |>
    mutate(ci_upper = agresti_coull(ARRESTS, stu_enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(ARRESTS, stu_enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Observed")


```



```{r}
#| label: browardsmallgroups
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

p1 <- ggplot(plot_draws, aes(x = (pred/(stu_enroll/1000)),
      color= RACE, fill = RACE, y = model_id)) +
    geom_density_ridges(scale = 0.9, rel_min_height = 0.01, alpha = 1/5, bandwidth = 0.1) +
    geom_pointrange(data = obsv_plot %>%
    filter(RACE == "HI") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (ARRESTS/(stu_enroll/1000)), xmin = ci_lower/(stu_enroll/1000), xmax = ci_upper/(stu_enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.35)) +
   geom_pointrange(data = obsv_plot %>%
    filter(RACE == "WH")|>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (ARRESTS/(stu_enroll/1000)), xmin = ci_lower/(stu_enroll/1000), xmax = ci_upper/(stu_enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.2)) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE), expand = expansion(c(0,0))) +
      facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    guides(color = guide_none()) +
    scale_color_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    scale_fill_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title), 60),
    x = "Arrest rate per 1,000", y = "",
    fill = "Student race:",
    subtitle = "Female students only.") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = -0.5))

diff_a <- plot_draws |>
  group_by(model_id, draw_id) |>
  summarize(group_h = pred[RACE == "HI"]/stu_enroll[RACE=="HI"],
            group_w = pred[RACE == "WH"]/stu_enroll[RACE=="WH"]) |>
    mutate(diffv = group_h - group_w)


annotate_df <- diff_a |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 0.7) |>
  mutate(diff_per = diffcount / total)
#annotate_df

p2 <- ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-0.75, to = 1, scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df,
        position = position_nudge(y = 0.6, x = 0),
        size = 2.5,
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE), expand = expansion(c(0,0))) +
      facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Model estimated difference (", "\u0394", ") between Hispanic and White student arrest rates in ", dist_name_for_title), 90),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    )+
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = -0.5))



diff_b <- diff_a |> mutate(model_id = "All Models Pooled") |>  mutate(model_id == as.character(model_id))

annotate_df <- diff_b |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 0.5) |>
  mutate(diff_per = diffcount / total)

p3 <- ggplot(diff_b,
    aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-1, to = 1, rel_min_height = 0.01, scale = 1, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df,
        size = 3.5,
        position = position_nudge(y = 0.7, x = 0.15),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    scale_y_discrete(expand = expansion(0, 0)) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Pooled model estimated differences (", "\u0394", ") between ",
    dist_name_for_title, " student arrest rates in 21-22"), 90),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    ) +
    theme(legend.position = "bottom", axis.text.y = element_blank())


#(p1 / p2)

#p1 / p2
p1 / p3


```

We can

## State example

```{r}
#| label: comparestates
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10
state_draws <- get_state_prediction_draws(con, LEA_STATE = "ALL", YEAR = "21-22",
                          model = c("nat_m1_mod", "nat_m2_mod", "nat_m3_mod", "nat_m4_mod",
                                    "sg_m1_mod", "sg_m2_mod", "sg_m3_mod", "sg_m4_mod"),
                          RACE = "ALL", SEX = "ALL")

state_obsv <- rdata |> group_by(LEA_STATE, YEAR) |>
              summarize(arrests = sum(ARRESTS),
                      enroll = sum(stu_enroll))
state_obsv <- state_obsv |>
   rowwise() |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "observed")

facet_state <- function(x) {
  y <- rep(1, length(x))
  y[ x %in% state.abb[13:25]] <- 3
  y[ x %in% state.abb[26:37]] <- 2
  y[ x %in% state.abb[38:50]] <- 4
  y[x == "DC"] <- 1
  y[x == "PR"] <- 3
  y[x == "VI"] <- 4
  return(y)
}

plot_draws <- inner_join(state_draws, state_obsv |> select(LEA_STATE, YEAR, arrests, enroll))
plot_draws <- plot_draws |>
                mutate(arr_rate = 1000 * fitted_value / enroll)
state_obsv <- state_obsv |>
            mutate(arr_rate = 1000 * arrests / enroll)

rm(state_draws); gc()

```


```{r}
#| label: plotstates
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10
 ggplot(plot_draws |> filter(model_id %in% c("nat_m3_mod", "sg_m2_mod")),
        aes(x = arr_rate, y = LEA_STATE, fill = model_id)) +
    geom_density_ridges(scale = 1.5, rel_min_height = 0.01, alpha = 3/5, bandwidth = 0.1) +
    geom_pointrange(data = state_obsv ,
   aes(y = LEA_STATE, x = arr_rate,fill = NULL, xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000)),
   show.legend = FALSE,
   position = position_nudge(y = 0.35)) +
   facet_wrap(~facet_state(LEA_STATE), scales = "free_y") +
   scale_x_continuous(expand = expansion(0,0)) +
   scale_y_discrete(limits = rev) +
  scale_fill_manual(values = c("#d55c00a9", "#0071b2c5"),
  labels = function(x) print_model_name(x, lbreak=FALSE)) +
       labs(title = "State arrest rate estimates, 2021-22",
    x = "Arrest rate per 1,000", y = "",
    fill = "Model:") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(  strip.background = element_blank(),
  strip.text.x = element_blank(), legend.position = "bottom")
```



```{r}
#| label: prepfirsttable
#| echo: false
#| warning: false
#| message: false


state_table_draws <- get_state_prediction_draws(con, LEA_STATE = c("AK", "CO"),
                                                      YEAR = "21-22",
                          model = c("nat_m1_mod", "nat_m2_mod", "nat_m3_mod", "nat_m4_mod",
                                    "sg_m1_mod", "sg_m2_mod", "sg_m3_mod", "sg_m4_mod"),
                          RACE = c("BL", "AM"), SEX = c("M", "F"))


state_tab_obsv <- rdata |> filter(LEA_STATE %in% c("AK", "CO")) |>
                    group_by(LEA_STATE, YEAR, RACE, SEX) |>
              summarize(arrests = sum(ARRESTS),
                      enroll = sum(stu_enroll))

state_tab_obsv <- state_tab_obsv |>
   rowwise() |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "observed")



plot_draws <- inner_join(state_table_draws, state_tab_obsv |>
                    select(LEA_STATE, RACE, SEX, YEAR, arrests, enroll))
plot_draws <- plot_draws |>
                mutate(arr_rate = 1000 * fitted_value / enroll)
state_tab_obsv <- state_tab_obsv |>
            filter(RACE %in% c("BL", "AM")) |>
            mutate(arr_rate = 1000 * arrests / enroll) |>
            mutate(rowid = paste(LEA_STATE, RACE, SEX, sep = "-"))


plot_draws$rowid <- paste(plot_draws$LEA_STATE, plot_draws$RACE, plot_draws$SEX, sep = "-")

keep_ids <- c("AK-BL-M", "AK-BL-F", "AK-AM-M", "CO-AM-M")
parse_row_ids <- function(x) {
  y <- x
  y[x == "AK-BL-M"] <- "AK Black \nMale"
  y[x == "AK-BL-F"] <- "AK Black \nFemale"
  y[x == "AK-AM-M"] <- "AK Amer. Indian \nAlaska Native Male"
  y[x == "CO-AM-M"] <- "CO Amer. Indian \nAlaska Native Male"
  return(y)

}
```


```{r}
#| label: recreatefirsttable
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10
p1 <- ggplot(plot_draws |> filter(rowid %in% keep_ids) |>
        filter(model_id %in% c("nat_m3_mod", "sg_m2_mod")),
        aes(x = fitted_value/(enroll/1000), y = rowid, fill = model_id)) +
    geom_density_ridges(scale = 1, from = 0, to = 9,
                      rel_min_height = 0.01, alpha = 3/5, bandwidth = 0.4) +
    geom_pointrange(data = state_tab_obsv  |> filter(rowid %in% keep_ids) ,
   aes(y = rowid, x = arr_rate,fill = NULL, xmin = 1000*ci_lower/enroll, xmax = 1000*ci_upper/enroll),
   show.legend = FALSE,
   position = position_nudge(y = 0.35)) +
    scale_y_discrete(expand = expansion(0, 0), labels = function(x) parse_row_ids(x)) +
      scale_fill_manual(values = c("#d55c00a9", "#0071b2c5"),
            labels = function(x) print_model_name(x, lbreak=FALSE)) +
    #coord_flip() +
   #facet_wrap(~LEA_STATE, scales = "free_y") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 12) +
    labs( x = "Arrests per 1,000", y ="", fill = "Model",
    title = "Bayesian modeled arrest rate predictions for selected state demographic groups") +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = 0.5))



diff_a <- plot_draws |> filter(rowid %in% c("AK-AM-M", "CO-AM-M")) |>
  ungroup() |> select(model_id, draw_id, rowid, fitted_value, enroll) |>
        filter(model_id %in% c("nat_m3_mod", "sg_m2_mod")) |>
   group_by(model_id,draw_id) |>
  summarize(group_b = fitted_value[rowid == "AK-AM-M"]/enroll[rowid == "AK-AM-M"],
            group_a = fitted_value[rowid == "CO-AM-M"]/enroll[rowid == "CO-AM-M"]) |>
    mutate(diffv =  group_a - group_b)

annotate_df <- diff_a |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 2) |>
  mutate(diff_per = diffcount / total)


p2 <- ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-1.25, to = 4, alpha = 4/5, rel_min_height = 0.02, bandwidth = 0.2) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df,
        size = 4.5,
        position = position_nudge(y = 0.65, x = 0),
    aes(y = model_id, x = diffv, label = paste0("Pr(CO > AK):\n", pretty_per(diff_per))))  +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 12) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap("Difference between Colorado and Alaska American Indian / Alaska Native Male Arrest Rates", 55),
    subtitle = "No difference shown as red vertical line.", y = "") +
      scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE),
      expand = expansion(c(0,0))) +
    theme(legend.position = "bottom",
    axis.text.y = element_text(angle = 90, hjust = -0.5))


diff_b <- plot_draws |> filter(rowid %in% c("AK-AM-M", "AK-BL-M")) |>
  ungroup() |> select(model_id, draw_id, rowid, fitted_value, enroll) |>
        filter(model_id %in% c("nat_m3_mod", "sg_m2_mod")) |>
   group_by(model_id,draw_id) |>
  summarize(group_a = fitted_value[rowid == "AK-AM-M"]/enroll[rowid == "AK-AM-M"],
            group_b = fitted_value[rowid == "AK-BL-M"]/enroll[rowid == "AK-BL-M"]) |>
    mutate(diffv =  group_a - group_b)

annotate_df2 <- diff_b |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 1.25) |>
  mutate(diff_per = diffcount / total)



p3 <- ggplot(diff_b, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-1.25, alpha = 4/5, rel_min_height = 0.02, bandwidth = 0.2) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df2,
        size = 4.5,
        position = position_nudge(y = 0.5, x = 0),
    aes(y = model_id, x = diffv, label = paste0("Pr( Amer. Ind > Black):\n", pretty_per(diff_per))))  +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 12) +
    labs(x = "Arrest rate per 1,000", y = "",
  title = stringr::str_wrap("Difference between Black and American Indian / Alaska Native male students within Alaska", 55)) +
            # subtitle = "No difference shown as red vertical line.", y = "") +
        scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE),
        expand = expansion(c(0,0))) +
     theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = -0.5))


p1 / (p2 + p3)

```

```{r}
#| label: cleanup
#| include: false

close_draws_view(h)
```
