---
title: "Applied Examples of Prediction Intervals"
format:
  html:
    embed-resources: true
knitr:
  opts_chunk:
    message: false
    warning: false
    error: false
    echo: false
    fig.path: export/figures/appliedexample-
execute:
  warning: false
  message: false
  error: false
  fig-width: 10
  fig-height: 6
---

## Introduction

Using statistical modeling we are able to obtain more precise estimates of the
arrest rate for districts and states. This report provides an example for
how these estimates can be applied to provide better demographic, geographic,
and temporal comparisons.

```{r}
#| label: setup
#| include: false

# Initialize our data
# Get some samples
# Prepare our data for plotting

library(brms)
library(dplyr)
library(ggplot2)
library(ggdist)
library(marginaleffects)
library(tidybayes)
library(ggridges)
library(duckdb)
library(patchwork)

source("R/funs.R")
source("R/crdc_path.R"); source("R/model_registry.R"); source("R/paper_figures.R")
cv_apply_branding()   # civilytics theme on all figures (+ logo when magick present)

# predicted_draws VIEW over the PUBLISHED draws parquet (CRDC_ARTIFACTS selects
# the source: local "export" for the owner, hf:// for a stranger). The existing
# get_prediction_summary(con, ...) calls below work unchanged.
h <- open_draws_view(); con <- h$con

tydata <- read_stage_df("stages/inputs/three_year_data.parquet")
rdata  <- read_stage_df("stages/inputs/recent_data.parquet")

# Model-labelling helpers (add_model_time/_form/_cov, print_model_name) come from
# R/paper_figures.R (sourced above). The examples below use the colour palettes
# as variables; define them from the shared builders so they no longer depend on
# the removed case-study chunks.
model_palette <- wp_model_palette()
group_palette <- wp_group_palette()

```

The three foundational single-district case studies (a zero-arrest district, a
high-arrest district, and a diverse-district group comparison) now live in the
white paper. This report collects the additional applied examples that build on
them:

1. Comparing two districts (Baltimore City vs. Baltimore County)
2. Comparing a single district over time (Baltimore County across CRDC waves)
3. A small-groups demographic disparity (Hispanic vs. White female students in
   Broward County)
4. A national, all-states overview of modeled arrest rates

We compare the predictions to the approximate interval calculated from the
data. In most cases we focus on the most recent wave of CRDC, 2021-22, but
where other years are used we discuss.

```{r}
#| label: get_data

# Get 100 highest arrests
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      arrange(desc(total_arrests)) |>
      ungroup() |>
      slice_head(n= 100) -> bigarr

# Get 100 largest districts
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      arrange(desc(total_enroll)) |>
      ungroup() |>
      slice_head(n= 100) -> big100

# Get largest districts with 0 arrests
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      filter(total_arrests == 0) |>
      arrange(desc(total_enroll)) |>
      ungroup() |>
      slice_head(n= 100) -> big0

# rdata |>
#   filter(YEAR == "21-22") |>
#   mutate(phat = (2+ARRESTS) / (4+stu_enroll)) |> # agcouli approximation ARRESTS +2 and stu_enroll + 4
#   mutate(phat_se = sqrt((phat * (1-phat)/(4+stu_enroll)))) |> # same phat but stu_enroll + 4 here
#   mutate(sd = phat_se * (4 + stu_enroll)) |> # calculate rate standard error to arrest scale
#   mutate(fitted_value = ARRESTS) |>
#   mutate(CVp = 100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))) |> # express CVp as percentage, this is a biased estimator in rare events
#   # agresti-coull confidence interval for arrests using the adjusted arrest rate
#   mutate(ci_upper = (phat + (1.96*phat_se)) * (4 + stu_enroll),
#         ci_lower = (phat - (1.96*phat_se)) * (4 + stu_enroll)) |>
#             # add rule of three correction
#   mutate(ci_upper = ifelse(ARRESTS == 0, 3, ci_upper),  # add 3
#         ci_lower = ifelse(ARRESTS == 0, 0, ci_lower)) |>
#   mutate(ci_lower = ifelse(ci_lower < 0, 0, ci_lower)) |>
#   mutate(model_id = "observed") -> obsv_data



agresti_coull <- function(numerator, denominator, confidence_level = 0.95) {
    adj_star <- qnorm(1 - (1-confidence_level)/2)
    # critical value
    if (numerator > 0) {
        num_star <- numerator + adj_star
        denom_star <- denominator + (2 * adj_star)
        phat <- num_star / denom_star
        phat_se <- sqrt((phat/denom_star) * (1-phat))
        ci_upper <- (phat + (adj_star * phat_se)) * denom_star
        ci_lower <- (phat - (adj_star * phat_se)) * denom_star
        sd <- mean(c(ci_upper, ci_lower)) / confidence_level
        sd <- phat_se * denom_star
    } else {

        ci_upper <- denominator*(-log(1-confidence_level)/denominator)
        ci_lower <- 0
        sd <- mean(c(ci_upper, ci_lower)) / confidence_level
        phat_se <- 0
        phat <- mean(c(ci_upper, ci_lower))
    }
    return(c(ci_upper, ci_lower, sd, phat_se, phat))
}


```

The single-district case studies that previously appeared here -- a
zero-arrest district (Paterson, NJ), a high-arrest district (Mobile County,
AL), and a diverse-district group comparison (Clark County, NV) -- are now the
canonical worked examples in the "Applied Examples" section of
`white_paper.qmd` (rendered there via the `wp_fig_*` builders in
`R/paper_figures.R`). They are not duplicated here. What remains below are the
examples unique to this document: cross-district and temporal comparisons and
a small-groups disparity.

## Multicomparison

Compare Baltimore City to Baltimore County
Baltimore City to itself last year
Broward White Female to Hispanic female



```{r}

focal_dist <- c("2400090", "2400120")

obsv_focal <- tydata |> filter(LEAID == focal_dist) |>
  group_by(YEAR, LEA_STATE, LEAID, LEA_NAME) |>
  summarize(enroll = sum(stu_enroll),
            enroll2 = first(total_enroll),
            referrals = sum(REFERRALS),
            arrests  = sum(ARRESTS)
            )


dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", LEAID = focal_dist)

dist_test <- dist_test |> filter(add_model_time(model_id) == "three year")
dist_draws <- get_prediction_draws(con,  LEAID = focal_dist)
dist_draws <- inner_join(dist_draws, obsv_focal |> select(YEAR:LEA_NAME, enroll, arrests))
dist_test <- inner_join(dist_test, obsv_focal |> select(YEAR:LEA_NAME, enroll, arrests))

plot_draws <- dist_draws |> group_by(model_id, LEAID, YEAR, draw_id) |>
              summarize(pred = sum(pred))

plot_draws <- inner_join(plot_draws, obsv_focal |> select(LEA_NAME, LEAID, YEAR, enroll, arrests))

dist_name_for_title <- gsub(" Public Schools", "",
                      tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))))



```

Here we apply the same approach as above but to compare arrest rates from two
different school districts - Baltimore City Public Schools (orange) and Baltimore County
Public Schools (blue). Here we see a curious pattern - if we look at our one
year models, the frequentist interval and Bayesian intervals strongly agree that
Baltimore City has a higher arrest rate than Baltimore County, with the most
likely arrest rate in Baltimore County being 0. Covariates make little difference
in these estimates.

The three year models have a very different picture - the Bayesian models see
Baltimore County as having not just a higher arrest rate, but a much higher
arrest rate than Baltimore City. The Baltimore City estimates are relatively
unchanged, though covariates seem to shift the distribution slightly closer to
0 (unified and stratified models 4).

When we turn to our direct estimate of the differences, we see this pattern play
out - the one year models are 100% certain in all specifications that Baltimore
City has a higher arrest rate than Baltimore County, and that difference is
most likely narrowly between just greater than 0 and 1. In contract, the 3 year
models are equally sure that Baltimore City has no chance of a higher arrest
rate than Baltimore County, and in fact, that it is most likely that Baltimore
County's arrest rate is 1-2 arrests per 1,000 greater than Baltimore City.
When we pool all of our models together, we are left with a 50% chance one
of the Baltimore's has a greater arrest rate than the other (inset, bottom left).

```{r}
#| label: baltimorecityandcountydifferences
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

obsv_plot <-  obsv_focal |>
  rowwise() |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Observed")


diff_a <- plot_draws |> filter(YEAR == "21-22") |>
  filter(!grepl("_m5", model_id)) |>
  group_by(model_id, draw_id) |>
  summarize(group_city = pred[LEAID == "2400090"]/enroll[LEAID=="2400090"],
            group_county = pred[LEAID == "2400120"]/enroll[LEAID=="2400120"]) |>
    mutate(diffv = group_city - group_county)

annotate_df <- diff_a |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 1) |>
  mutate(diff_per = diffcount / total)

plot_draws <-
  plot_draws |> filter(YEAR == "21-22") |>  filter(!grepl("_m5", model_id))

p2 <- ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-3, to = 1.5,  rel_min_height = 0.01, scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df |> filter(model_id %in% c("sg_m2_mod", "sg_m4_mod")),
        size = 3,
        position = position_nudge(y = 0.4, x = 0),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE)) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Model estimated difference (", "\u0394", ") between Baltimore City and Baltimore County student arrest rates in 21-22"), 120),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    ) +
      facet_wrap(~add_model_time(model_id), scales = "free_y") +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, size = 10, hjust = -0.5))

p1 <- ggplot(plot_draws,
    aes(x = (pred/(enroll/1000)),
      color= LEA_NAME, fill = LEA_NAME, y = model_id)) +
    geom_density_ridges(scale = 0.9, rel_min_height = 0.01, alpha = 2/5, bandwidth = 0.1) +
    geom_pointrange(data = obsv_plot |> filter(YEAR == "21-22") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (arrests/(enroll/1000)),
      xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = LEA_NAME, alpha = 1),
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(0,0)) +
    facet_wrap(~add_model_time(model_id), scales = "free_y") +
    guides(color = guide_none()) +
    scale_color_manual(values = c("#D55E0050", "#0072B250"), labels = function(x) stringr::str_wrap(tools::toTitleCase(tolower(x)), 80)) +
    scale_fill_manual(values = c("#D55E0050", "#0072B250"), labels = function(x) stringr::str_wrap(tools::toTitleCase(tolower(x)), 80)) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title[1], " and ", dist_name_for_title[2]), 120),
    x = "Arrest rate per 1,000", y = "Model type", fill = "", color = "") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, size = 10, hjust = -0.5))



diff_b <- diff_a |> mutate(model_id = "All Models Pooled") |>  mutate(model_id == as.character(model_id))
annotate_df <- diff_b |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 1) |>
  mutate(diff_per = diffcount / total)

p3 <- ggplot(diff_b,
    aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-3, to = 1.5, rel_min_height = 0.01, scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_text(data = annotate_df,
        size = 2.5,
        position = position_nudge(y = 1.2, x = -2.25),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 10) +
    scale_y_discrete(expand = expansion(0, 0)) +
    labs(x = "Arrest rate per 1,000",
    title = "All models pooled",
    y = "",
    ) +
    theme(legend.position = "bottom",
          axis.text.y = element_blank())


p1 / (p2 + inset_element(p3, left = 0, bottom = 0.2, top = 0.9, right = 0.24))





```

To better understand what is going on we can use our models to make temporal
comparisons in Baltimore County - what does the trend look like.

```{r}
#| label: baltimorecountytrend
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

# Temporal for Baltimore County

plot_draws <- dist_draws |> group_by(model_id, LEAID, YEAR, draw_id) |>
              summarize(pred = sum(pred))

plot_draws <- inner_join(plot_draws, obsv_focal |> select(LEA_NAME, LEAID, YEAR, enroll, arrests))

plot_draws <- plot_draws |> filter(add_model_time(model_id) == "Three year") |>
  filter(LEAID == "2400120") |>
  filter(!grepl("_m5", model_id))


dist_name_for_title <- gsub(" Public Schools", "",
                      tools::toTitleCase(tolower(unique(plot_draws$LEA_NAME))))


p1 <- ggplot(plot_draws ,
    aes(x = (pred/(enroll/1000)),
      color= I("black"), fill = YEAR, y = YEAR)) +
    geom_density_ridges(scale = 2.5, rel_min_height = 0.01, alpha = 9/10, bandwidth = 0.05) +

    geom_point(data = obsv_plot |> filter(LEAID == "2400120") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = YEAR, x = (arrests/(enroll/1000)),
   color = I("black")), size = 6, alpha = 1,
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
     geom_pointrange(data = obsv_plot |> filter(LEAID == "2400120") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = YEAR, x = (arrests/(enroll/1000)),
      xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000),
   color = YEAR), size = 1, alpha = 1,
   show.legend = FALSE,
   position = position_nudge(y = 0.25)) +
    scale_y_discrete() +
    facet_wrap(~print_model_name(model_id)) +
    guides(color = guide_none()) +
    guides(fill = guide_none()) +
    scale_color_brewer(palette = "Blues", direction = 1) +
    scale_fill_brewer(palette = "Blues", direction = 1) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title[1], " over time"), 120),
  subtitle = "Observed value shown by point",
    x = "Arrest rate per 1,000", y ="", fill = "", color = "") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom")


diff_a <-plot_draws |> filter(add_model_time(model_id) == "Three year") |>
  filter(LEAID == "2400120") |>
  group_by(model_id, draw_id) |>
  summarize(group_t = pred[YEAR == "21-22"]/enroll[YEAR=="21-22"],
            groupprev = pred[YEAR == "17-18"]/enroll[YEAR == "17-18"]) |>
    mutate(diffv = group_t - groupprev)

annotate_df <- diff_a |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv < 0),
            diffv = -5.5) |>
  mutate(diff_per = diffcount / total)

p2 <- ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient( rel_min_height = 0.01,
    from = -16, to = 1,
    scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_text(data = annotate_df |> filter(model_id == "sg_m4_mod"),
        size = 3,
        position = position_nudge(y = 0.7, x = 0),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "<0): ", pretty_per(diff_per)))) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE)) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Model estimated difference (", "\u0394", ") between ", dist_name_for_title, " and 2021-22 and 2017-18"), 120),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    ) +
      facet_wrap(~add_model_time(model_id), scales = "free_y") +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = -0.5))


p3 <- ggplot(obsv_plot, aes(x = YEAR, y = enroll, group = LEA_NAME, color = LEA_NAME)) +
  geom_point() + geom_line() +
  scale_color_manual(values = c("#d55c00ff", "#0071b2f6"), labels = function(x) stringr::str_wrap(tools::toTitleCase(tolower(x)), 50)) +
  scale_y_continuous("Enrollment", labels = scales::comma, limits = c(0,NA)) +
  guides(color = guide_legend(nrow = 2, override.aes = list(linetype = c(0, 0)))) +

  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE) +
  labs(color = "", x = "CRDC Year", title = "Reported enrollment") +
  theme(legend.position = "bottom")


p1 / (p2 + inset_element(p3, left = 0.67, bottom = 0.2, top = 0.8, right = 1))



```




```{r}
# Small groups
focal_dist <- "1200180"
obsv_focal <- rdata |> filter(LEAID == focal_dist) |>
  filter(SEX == "F") |>
  filter(RACE %in% c("WH", "HI"))

dist_draws <- get_prediction_draws(con,  LEAID = focal_dist, YEAR = "21-22",
                            SEX = "F", RACE = c("WH", "HI"))

plot_draws <- inner_join(dist_draws, obsv_focal |> select(YEAR:RACE, stu_enroll, ARRESTS)) |>
  filter(!grepl("_m5", model_id))

dist_name_for_title <- gsub(" Public Schools", "",
                      tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))))

obsv_plot <- obsv_focal |>
  rowwise() |>
    mutate(ci_upper = agresti_coull(ARRESTS, stu_enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(ARRESTS, stu_enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Observed")


```



```{r}
#| label: browardsmallgroups
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

p1 <- ggplot(plot_draws, aes(x = (pred/(stu_enroll/1000)),
      color= RACE, fill = RACE, y = model_id)) +
    geom_density_ridges(scale = 0.9, rel_min_height = 0.01, alpha = 1/5, bandwidth = 0.1) +
    geom_pointrange(data = obsv_plot %>%
    filter(RACE == "HI") |>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (ARRESTS/(stu_enroll/1000)), xmin = ci_lower/(stu_enroll/1000), xmax = ci_upper/(stu_enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.35)) +
   geom_pointrange(data = obsv_plot %>%
    filter(RACE == "WH")|>
   tidyr::crossing(model_id = unique(plot_draws$model_id)),
   aes(y = model_id, x = (ARRESTS/(stu_enroll/1000)), xmin = ci_lower/(stu_enroll/1000), xmax = ci_upper/(stu_enroll/1000),
   color = RACE),
   show.legend = FALSE,
   position = position_nudge(y = 0.2)) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE), expand = expansion(c(0,0))) +
      facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    guides(color = guide_none()) +
    scale_color_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    scale_fill_manual(values = group_palette, labels = function(x) crdc_race_recode(x)) +
    labs(title = stringr::str_wrap(paste0("Arrest rate probability density for ", dist_name_for_title), 60),
    x = "Arrest rate per 1,000", y = "",
    fill = "Student race:",
    subtitle = "Female students only.") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = -0.5))

diff_a <- plot_draws |>
  group_by(model_id, draw_id) |>
  summarize(group_h = pred[RACE == "HI"]/stu_enroll[RACE=="HI"],
            group_w = pred[RACE == "WH"]/stu_enroll[RACE=="WH"]) |>
    mutate(diffv = group_h - group_w)


annotate_df <- diff_a |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 0.7) |>
  mutate(diff_per = diffcount / total)
#annotate_df

p2 <- ggplot(diff_a, aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-0.75, to = 1, scale = 1.5, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df,
        position = position_nudge(y = 0.6, x = 0),
        size = 2.5,
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak=TRUE), expand = expansion(c(0,0))) +
      facet_wrap(add_model_cov(model_id) ~ add_model_time(model_id),
    strip.position = "top", scales = "free_y") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Model estimated difference (", "\u0394", ") between Hispanic and White student arrest rates in ", dist_name_for_title), 90),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    )+
    theme(legend.position = "bottom", axis.text.y = element_text(angle = 90, hjust = -0.5))



diff_b <- diff_a |> mutate(model_id = "All Models Pooled") |>  mutate(model_id == as.character(model_id))

annotate_df <- diff_b |> group_by(model_id) |>
  summarize(total = n(),
            diffcount = sum(diffv > 0),
            diffv = 0.5) |>
  mutate(diff_per = diffcount / total)

p3 <- ggplot(diff_b,
    aes(x = (diffv * 1000),  y = model_id, fill = after_stat(x))) +
    geom_density_ridges_gradient(from=-1, to = 1, rel_min_height = 0.01, scale = 1, alpha = 4/5, bandwidth = 0.05) +
    scale_fill_distiller(name = "Diff.", direction = 1, palette = "YlOrRd",
            guide = guide_none()) +
    geom_vline(xintercept = 0, linetype = 3, color = I("red"), linewidth = 2) +
    geom_text(data = annotate_df,
        size = 3.5,
        position = position_nudge(y = 0.7, x = 0.15),
    aes(y = model_id, x = diffv, label = paste0("Pr(", "\u0394", "> 0): ", pretty_per(diff_per)))) +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    scale_y_discrete(expand = expansion(0, 0)) +
    labs(x = "Arrest rate per 1,000",
    title = stringr::str_wrap(paste0("Pooled model estimated differences (", "\u0394", ") between ",
    dist_name_for_title, " student arrest rates in 21-22"), 90),
    subtitle = "No difference shown as red vertical line.",
    y = "",
    ) +
    theme(legend.position = "bottom", axis.text.y = element_blank())


#(p1 / p2)

#p1 / p2
p1 / p3


```

We can

## State example

```{r}
#| label: comparestates
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10
state_draws <- get_state_prediction_draws(con, LEA_STATE = "ALL", YEAR = "21-22",
                          model = c("nat_m1_mod", "nat_m2_mod", "nat_m3_mod", "nat_m4_mod",
                                    "sg_m1_mod", "sg_m2_mod", "sg_m3_mod", "sg_m4_mod"),
                          RACE = "ALL", SEX = "ALL")

state_obsv <- rdata |> group_by(LEA_STATE, YEAR) |>
              summarize(arrests = sum(ARRESTS),
                      enroll = sum(stu_enroll))
state_obsv <- state_obsv |>
   rowwise() |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "observed")

facet_state <- function(x) {
  y <- rep(1, length(x))
  y[ x %in% state.abb[13:25]] <- 3
  y[ x %in% state.abb[26:37]] <- 2
  y[ x %in% state.abb[38:50]] <- 4
  y[x == "DC"] <- 1
  y[x == "PR"] <- 3
  y[x == "VI"] <- 4
  return(y)
}

plot_draws <- inner_join(state_draws, state_obsv |> select(LEA_STATE, YEAR, arrests, enroll))
plot_draws <- plot_draws |>
                mutate(arr_rate = 1000 * fitted_value / enroll)
state_obsv <- state_obsv |>
            mutate(arr_rate = 1000 * arrests / enroll)

rm(state_draws); gc()

```


```{r}
#| label: plotstates
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10
 ggplot(plot_draws |> filter(model_id %in% c("nat_m3_mod", "sg_m2_mod")),
        aes(x = arr_rate, y = LEA_STATE, fill = model_id)) +
    geom_density_ridges(scale = 1.5, rel_min_height = 0.01, alpha = 3/5, bandwidth = 0.1) +
    geom_pointrange(data = state_obsv ,
   aes(y = LEA_STATE, x = arr_rate,fill = NULL, xmin = ci_lower/(enroll/1000), xmax = ci_upper/(enroll/1000)),
   show.legend = FALSE,
   position = position_nudge(y = 0.35)) +
   facet_wrap(~facet_state(LEA_STATE), scales = "free_y") +
   scale_x_continuous(expand = expansion(0,0)) +
   scale_y_discrete(limits = rev) +
  scale_fill_manual(values = c("#d55c00a9", "#0071b2c5"),
  labels = function(x) print_model_name(x, lbreak=FALSE)) +
       labs(title = "State arrest rate estimates, 2021-22",
    x = "Arrest rate per 1,000", y = "",
    fill = "Model:") +
    coord_cartesian(clip = "off") +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(  strip.background = element_blank(),
  strip.text.x = element_blank(), legend.position = "bottom")
```

The recreation of the Table 1 state demographic-group differences is now part
of the white paper's "Comparing state-level demographic group differences"
(`wp_fig_state_differences`); only the all-states overview above is kept here.


```{r}
#| label: cleanup
#| include: false

close_draws_view(h)
```
