---
title: "Combined Model Data EDA"
format: html
execute:
  warning: false
  message: false
  error: false
  fig-width: 10
  fig-height: 6
---

## Overview

This document explores the combined data across all years (2015-16, 2017-18, 2021-22) from the Civil Rights Data Collection.

```{r}
#| label: setup
#| include: false
library(targets)
library(dplyr)
library(ggplot2)
library(knitr)


options(scipen = 9)
options(width = 120)

# Load the combined data
combined_data <- targets::tar_read(combined_model_data)

# Overwrite name
combined_data |>
  select(LEAID, LEA_NAME, YEAR) |>
  distinct_all() |>
  group_by(LEAID) |>
  mutate(last_name = ifelse(any(YEAR == "21-22"),
                            LEA_NAME[YEAR == "21-22"],
                            ifelse(any(YEAR == "17-18"),
                            LEA_NAME[YEAR == "17-18"],
                            LEA_NAME[YEAR == "15-16"]))) |>
  ungroup() |>
  select(LEAID, last_name) |>
  distinct_all() -> lea_names_canonical

combined_data <- left_join(combined_data |> select(-LEA_NAME),
                          lea_names_canonical |>
                            rename(LEA_NAME = last_name),
                          by = join_by(LEAID))

rm(lea_names_canonical)


# Load state shapes
state_map <- tigris::states(cb  = TRUE, resolution = "20m") |> tigris::shift_geometry()



# Add colors
color_fill_values <- c("WH" = "#c92d0e", "BL" = "#020684",
                        "AM" = "#ff8f43",
                       "TOTAL" = "#909090ff", "HI" = "#0791b6")
color_fill_values2 <- c("White" = "#c92d0e", "Black" = "#020684",
                        "Total" = "#909090ff", "Hispanic" = "#0791b6",
                        "Amer. Ind." = "#ff8f43")


```

## Basic Data Structure

```{r}
#| label: data-structure
# Basic structure and dimensions
cat("Data dimensions:", nrow(combined_data), "rows,", ncol(combined_data), "columns\n\n")

# Column names and types
str(combined_data)

# First few rows
head(combined_data) |> kable()
```

## Summary Statistics by Year

```{r}
#| label: year-summary
# Summary statistics by year
year_summary <- combined_data |>
  group_by(YEAR) |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    mean_enrollment = mean(stu_enroll, na.rm = TRUE),
    mean_arrests = mean(ARRESTS, na.rm = TRUE),
    mean_referrals = mean(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )

year_summary |> knitr::kable(digits = 2)



```

### Summary Plot of Trends

```{r}

year_summary <- combined_data |>
  group_by(YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(year_summary, aes(x = YEAR, y = total_arrests)) +
  geom_point() +
  geom_line(group = 1) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  labs(x = "School Year", y = "CRDC Arrests", title = "Arrest Trends Across CRDC Collections") +
  civilytics::theme_civilytics(font_size =12)

ggplot(year_summary, aes(x = YEAR, y = total_arrests/(total_enrollment/1000))) +
  geom_point() +
  geom_line(group = 1) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  labs(x = "School Year", y = "CRDC Arrests per 1,000", title = "Arrest Rate Trends Across CRDC Collections") +
  civilytics::theme_civilytics(font_size =12)




```

## State Variation Summing over 3 years

```{r}
#| label: state_map_arrest_rate
#| echo: false
#| warning: false
#| message: false

state_summary <- combined_data |>
  group_by(LEA_STATE) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate =  total_arrests / (total_enrollment / 1000)) |>
    mutate(arrest_nnh = total_enrollment / total_arrests)



state_map2 <- inner_join(state_map, state_summary, by = join_by(STUSPS == LEA_STATE))

ggplot(data = state_map2, aes( fill = arrest_rate)) +
  geom_sf() +
  geom_sf_label(aes(label = round(arrest_rate, 1))) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  labs(title = "Arrests per 1,000 for all students by state") +
  theme_void()

#rm(state_map2)

```



```{r}
#| label: state_map_arrestsnnh
#| echo: false
#| warning: false
#| message: false
ggplot(data = state_map2, aes( fill = arrest_nnh)) +
  geom_sf() +
  geom_sf_label(aes(label = civilytics::pretty_count(round(arrest_nnh, 0)))) +
  scale_fill_distiller(palette = "YlOrRd", direction = -1,
  breaks = c(50000, 8000, 3000, 1000, 500),
  trans = "log", labels = scales::comma) +
  labs(title = "Students per arrested student by state", fill = "NNH") +
  theme_void() +
  theme(legend.position.inside = c(0.1, 0.8), legend.position = "inside")
```

## Race and Sex Distribution

```{r}
#| label: race-sex-distribution
# Distribution across race and sex categories
race_sex_summary <- combined_data |>
  filter(RACE != "TOTAL", SEX != "TOTAL") |>
  group_by(RACE, SEX) |>
  summarise(
    n_observations = n(),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate = total_arrests / (total_enrollment / 1000)) |>
  arrange(desc(total_arrests))

race_sex_summary |> knitr::kable(digits = 3)
```



```{r}
#| label: nnh_arrest_counts_3yr

ggplot(race_sex_summary |> filter(RACE %in% c("AM", "WH", "HI", "BL"))) +
  aes(x = crdc_race_recode(RACE), y = 1000 / arrest_rate,
  fill = crdc_race_recode(RACE)) +
    scale_fill_manual(values = color_fill_values2) +
  geom_col() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0))),
              vjust = -0.25) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14) +
  labs(x = "", y = "NNH (Arrests)", fill = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())

```

### Trends by group

```{r}
#| label: race-sex-arrestratetrends
race_sex_trends <- combined_data |>
  filter(RACE != "TOTAL", SEX != "TOTAL") |>
  group_by(YEAR, RACE, SEX) |>
  summarise(
    n_observations = n(),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate = total_arrests / (total_enrollment / 1000)) |>
  arrange(desc(total_arrests))



ggplot(race_sex_trends |> filter(RACE %in% c("AM", "WH", "HI", "BL")) |>
  mutate(group_var = paste(RACE, SEX, sep = "_"))) +
  aes(x = YEAR, y = 1000 / arrest_rate, group = group_var,
  color = crdc_race_recode(RACE)) +
    scale_color_manual(values = color_fill_values2) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0)),
                color = NULL),
              vjust = -0.25, show.legend = FALSE) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14) +
  labs(x = "", y = "NNH (Arrests)", color = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4000),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())


ggplot(race_sex_trends |> filter(RACE %in% c("AM", "WH", "HI", "BL")) |>
  mutate(group_var = paste(RACE, SEX, sep = "_"))) +
  aes(x = YEAR, y = arrest_rate, group = group_var,
  color = crdc_race_recode(RACE)) +
    scale_color_manual(values = color_fill_values2) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(round(arrest_rate, 2)),
                color = NULL),
              vjust = -0.25, show.legend = FALSE) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14) +
  labs(x = "", y = "Arrests per 1,000", color = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())


```


## State Trends

```{r}
state_trend_summary <- combined_data |>
  filter(LEA_STATE %in% c("KS", "CA")) |>
  group_by(LEA_STATE, YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate =  total_arrests / (total_enrollment / 1000)) |>
    mutate(arrest_nnh = total_enrollment / total_arrests)

ggplot(state_trend_summary, aes(x = YEAR, y = total_arrests, group = LEA_STATE)) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(total_arrests)), nudge_y = -200) +
  geom_text(data = state_trend_summary |> filter(YEAR == "21-22"),
            aes(label = LEA_STATE), nudge_x = 0.25) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4000),
                    labels = scales::comma) +
    labs(x = "CRDC Wave", y = "Total arrests", title = "State-level total arrest trend comparison") +
    civilytics::theme_civilytics(font_size = 14)



```
## LEA Comparison

If we aggregate to the LEA level, what patterns do we see?

```{r}

combined_data |>
    filter(total_enroll > 0) |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEAID) |>
    summarize(
          nyears = n_distinct(YEAR),
              nnames = n_distinct(LEA_NAME)
    ) -> lea_samp_summary

table(lea_samp_summary$nyears) |> prop.table()
nrow(lea_samp_summary)



combined_data |>
    filter(total_enroll > 0) |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEAID) |>
    mutate(
          nyears = n_distinct(YEAR)
    ) |>
      group_by(nyears) |>
      summarize(
                obs = n_distinct(LEAID),
                enrollment = sum(stu_enroll),
                ARRESTS = sum(ARRESTS)) -> lea_samp_summary


(lea_samp_summary$enrollment / lea_samp_summary$nyears) / sum(lea_samp_summary$enrollment / lea_samp_summary$nyears)


```


### Rankings over time

```{r}


combined_data |>
#  filter(total_enroll > 0) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  group_by(YEAR, LEAID, LEA_NAME, LEA_STATE) |>
  summarize(enroll = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS, na.rm = TRUE)) |>
  group_by(YEAR) |>
  mutate(arrest_rank = rank(arrests)) |>
  #arrange(desc(arrest_rank), YEAR) |>
  tidyr::pivot_wider(
        values_from = c(enroll, arrests, arrest_rank),
                    names_from = YEAR) |>
  ungroup() |>
  filter(!is.na(`arrests_21-22`)) |>
  arrange(desc(`arrests_21-22`)) |>
  #mutate(LEA_NAME = stringr::str_trunc(LEA_NAME, width = 24)) |>
  mutate(arrest_rate_2122 = round(`arrests_21-22` / (`enroll_21-22` / 1000), 2)) |>
  select(LEA_STATE, LEA_NAME, `enroll_21-22`, `arrests_21-22`,
          arrest_rate_2122,  `arrests_17-18`, `arrests_15-16`) -> arrest_rank_leaderboard


  write.csv(arrest_rank_leaderboard, file = "export/arrest_rank_leader_table.csv")



  arrest_rank_leaderboard |>
    rename(State = LEA_STATE,
           `LEA Name` = LEA_NAME,
           `Enrollment (21-22)` = `enroll_21-22`,
           `Arrests (21-22)` = `arrests_21-22`,
           `Arrest Rate (21-22)` = `arrest_rate_2122`,
           `Arrests (17-18)` = `arrests_17-18`,
           `Arrests (15-16)` = `arrests_15-16`
           ) |>
     head(20) |> knitr::kable()


```

### Rankings of arrest rates

```{r}


combined_data |>
#  filter(total_enroll > 0) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  group_by(YEAR, LEAID, LEA_NAME, LEA_STATE) |>
  summarize(enroll = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS, na.rm = TRUE),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = round(arrests / (enroll / 1000), 1)) |>
  group_by(YEAR) |>
  #arrange(desc(arrest_rank), YEAR) |>
  tidyr::pivot_wider(
        values_from = c(enroll, arrests, arrest_rate, jj_count),
                    names_from = YEAR) |>
  ungroup() |>
  filter(!is.na(`arrests_21-22`)) |>
  arrange(desc(`arrest_rate_21-22`)) |>
  #mutate(LEA_NAME = stringr::str_trunc(LEA_NAME, width = 24)) |>
  select(LEA_STATE, LEA_NAME, `jj_count_21-22`,
        `arrest_rate_21-22`, `arrest_rate_17-18`,
        `arrest_rate_15-16`,
        `enroll_21-22`, `arrests_21-22`,
        `arrests_17-18`,
        `arrests_15-16`) -> arrest_rate_rank_leaderboard


  arrest_rate_rank_leaderboard |>
    select(-`jj_count_21-22`) |>
    rename(State = LEA_STATE,
           `LEA Name` = LEA_NAME,
           `Enrollment (21-22)` = `enroll_21-22`,
           `Arrests (21-22)` = `arrests_21-22`,
           `Arrest Rate (21-22)` = `arrest_rate_21-22`,
           `Arrest Rate (17-18)` = `arrest_rate_17-18`,
           `Arrest Rate (15-16)` = `arrest_rate_15-16`,
           `Arrests (17-18)` = `arrests_17-18`,
           `Arrests (15-16)` = `arrests_15-16`
           ) |>
     head(20) |> knitr::kable()

```


```{r}
#| label: badcells
#| eval: false
#| include: false
# Check waterbury
combined_data |> filter(LEAID == "0634320") |> View()
old_data <- targets::tar_read("model_data_y1516")
old_data |> filter(LEAID == "0634320") |> View()

old_data |> filter(LEA_STATE == "CA") |>
  filter(SEX == "F", RACE == "BL") |>
  group_by(LEAID, LEA_NAME) |>
  summarize(stu_enroll = sum(stu_enroll),
            nschls = n(),
            arrests = sum(ARRESTS)) |>
  filter(stu_enroll > 2000)
```
