---
title: "CRDC Arrest Analysis Results"
format: html
editor: visual
---

# Setup and Data Loading

```{r}
#| label: setup
#| message: false
#| warning: false

library(targets)
library(brms)
library(dplyr)
library(ggplot2)
library(ggdist)
library(marginaleffects)
library(tidybayes)
library(ggridges)
library(duckdb)

# Create export/figures directory if it doesn't exist
if (!dir.exists("export/figures")) {
  dir.create("export/figures", recursive = TRUE)
}
```

```{r}
#| label: load-data
#| message: false
#| warning: false

# Connect to database
con <- DBI::dbConnect(duckdb::duckdb(), "export/db/crdc_arrests.duckdb")

# Load target data
tydata <- targets::tar_read(three_year_data)$data
rdata <- targets::tar_read(recent_data)$data

# Source custom functions
source("R/funs.R")
```

# National Descriptive Statistics (2021-22)

## Overall Totals

```{r}
#| label: compute-overall-totals

y2122 <- targets::tar_read("full_crdc_data_y2122")
```

```{r}
#| label: display-total-arrests

# Total students arrested
sum(y2122$ARRESTS[y2122$RACE == "TOTAL" & y2122$SEX=="TOTAL"])
```

```{r}
#| label: display-total-enrollment

# Enrollment
sum(y2122$stu_enroll[y2122$RACE == "TOTAL" & y2122$SEX=="TOTAL"])
```

```{r}
#| label: display-arrest-rate

34846/(48596489/1000)
```

## By Race

```{r}
#| label: compute-race-totals

nat_sg_totals_race <- y2122 |>
    ungroup() |>
    filter(RACE != "TOTAL" & SEX != "TOTAL") |>
    group_by(RACE) |>
    summarize(
        lea_count = n_distinct(LEAID),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000),
            referral_rate = referrals / (enrollment/1000))
```

```{r}
#| label: display-race-totals

nat_sg_totals_race |>
    filter(RACE %in% c("AM", "WH", "HI", "BL"))
```

## By Sex

```{r}
#| label: compute-sex-totals

nat_sg_totals_sex <- y2122 |>
    ungroup() |>
    filter(RACE != "TOTAL" & SEX != "TOTAL") |>
    group_by(SEX) |>
    summarize(
        lea_count = n_distinct(LEAID),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000),
            referral_rate = referrals / (enrollment/1000))
```

```{r}
#| label: display-sex-totals

nat_sg_totals_sex
```

## District-Level Totals

```{r}
#| label: compute-district-totals

nat_totals <- y2122 |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS))

nat_referral_rate <- nat_totals$referrals / (nat_totals$enrollment / 1000)
nat_arrest_rate <- nat_totals$arrests / (nat_totals$enrollment / 1000)
```

```{r}
#| label: display-district-totals

nat_totals
```

# Sample Continuity Analysis

```{r}
#| label: load-historical-data

y1718 <- targets::tar_read("full_crdc_data_y1718")
y1516 <- targets::tar_read("full_crdc_data_y1516")
```

```{r}
#| label: match-test-1718-2122

match_test(y1718$LEAID, y2122$LEAID)
```

```{r}
#| label: match-test-1516-1718

match_test(y1516$LEAID, y1718$LEAID[y1718$LEAID %in% unique(y2122$LEAID)])
```

# District Concentration Analysis

## District Size Categories

```{r}
#| label: compute-district-categories

thresh <- c(0, 1000, 10000, 20000, Inf)
labs <- c("0-999", "1,000-9,999", "10,000-19,999","20,000+")

dist_ref_arr <- y2122 %>%
  group_by(YEAR, LEA_STATE, LEAID, RACE, SEX) %>%
  summarize(ARRESTS = sum(ARRESTS),
            REFERRALS = sum(REFERRALS),
            stu_enroll = sum(stu_enroll),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = ARRESTS / (stu_enroll / 1000),
         referral_rate = REFERRALS / (stu_enroll / 1000)) %>%
  group_by(YEAR, LEA_STATE, LEAID) |>
  mutate(total_enroll = sum(stu_enroll[RACE != "TOTAL" &
                                         SEX != "TOTAL"]))
```

## Expected vs Observed Arrests

```{r}
#| label: compute-expected-arrests

dist_size_summary <- dist_ref_arr %>%
   ungroup() %>%
     filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
   mutate(popcut = cut(stu_enroll,
                       breaks = thresh,
                       labels = labs,
                       include.lowest= TRUE)) %>%
   group_by(popcut) %>%
 mutate(expec_arrest = pbinom(1, stu_enroll, nat_arrest_rate/1000,
                              lower.tail = FALSE)) %>%
   summarize(dists = n(),
             dists_w_arrests = sum(ARRESTS>0),
             expec_distw_arrests = round(sum(expec_arrest),1)) %>%
   mutate(dist_w_arrest_per = pretty_per(dists_w_arrests/dists),
          expect_w_arrest_per = pretty_per(expec_distw_arrests/dists))
```

```{r}
#| label: display-expected-arrests

dist_size_summary
```

## Concentration Visualization

```{r}
#| label: compute-concentration-data

total_enroll_national <- sum(dist_ref_arr$total_enroll[
  dist_ref_arr$RACE == "TOTAL" & dist_ref_arr$SEX == "TOTAL"
])

concentration_plot_data <- dist_ref_arr %>%
  ungroup %>%
  filter(ARRESTS > 0) %>%
  filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
  arrange(desc(stu_enroll)) %>%
  mutate(total_arr = sum(ARRESTS)) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_enroll_national) %>%
  mutate(dist_per_cum = cumsum(dist_per))
```

```{r}
#| label: plot-concentration
#| fig-cap: "Share of total arrests by cumulative district enrollment"
#| fig-width: 10
#| fig-height: 6

concentration_plot <- concentration_plot_data %>%
  ggplot(aes(x = dist_per_cum, y = cum_sum_per)) +
  geom_line() +
  geom_hline(yintercept = 0.5) +
  scale_x_continuous("% of national student population",
                     labels = scales::percent) +
  scale_y_continuous("% of all arrests", labels = scales::percent) +
  theme_minimal(base_size = 12) +
  labs(title = "Share of total arrests by cumulative district enrollment",
       subtitle = "Horizontal line at 50% of all arrests") +
  theme(axis.ticks.x = element_blank(),
        panel.grid = element_blank())

print(concentration_plot)

# Save high-resolution figure
ggsave("export/figures/arrest_concentration_by_district_enrollment.png",
       plot = concentration_plot,
       width = 10, height = 6, dpi = 300, bg = "white")
```

## Enrollment by Arrest Status

```{r}
#| label: compute-enrollment-by-arrest

enrollment_by_arrest <- dist_ref_arr %>%
  ungroup %>%
  filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
  mutate(has_arrest = ifelse(ARRESTS > 0, "Arrests", "No arrests")) |>
  group_by(has_arrest) |>
  summarize(enrollment = sum(total_enroll)) |>
  mutate(enrollment_per = enrollment / total_enroll_national)
```

```{r}
#| label: display-enrollment-by-arrest

enrollment_by_arrest
```

# Sample Restriction Impact

## Section 504 Data

```{r}
#| label: compute-504-data

sch_ref <- targets::tar_read("lerefs_y2122")

section_504_summary <- sch_ref %>%
  select(!matches("TOT_")) %>%
  select(matches("_504_")) |>
  ungroup() |>
  summarize_all(~ sum(crdc_sub(.x), na.rm = TRUE))
```

```{r}
#| label: display-504-data

section_504_summary
```

## Missing Data Patterns

```{r}
#| label: compute-missing-patterns

# Drop non COMBOKEY identifier columns
sch_ref <- sch_ref %>% select(COMBOKEY:last_col())
names(sch_ref) <- gsub("_IDEA_", "_", names(sch_ref))

# Check for reserve values
missing_pattern_table <- sch_ref %>%
  select(!matches("TOT_")) %>%
  select(!matches("_504_")) %>%
  select(!matches("_LEP_")) %>%
  select(!matches("_EL_")) %>%
  tidyr::pivot_longer(cols = matches("SCH_DIS"), names_prefix = "SCH_",
              names_to = c("DISAB", "REF_TYPE", "RACE", "SEX"),
              names_sep = "_") %>%
  group_by(COMBOKEY, DISAB, REF_TYPE, RACE, SEX) |>
  mutate(value_mod = ifelse(value < 0, TRUE, FALSE)) |>
  pull(value_mod) |>
  table()
```

```{r}
#| label: display-missing-patterns

missing_pattern_table
```

```{r}
#| label: display-missing-proportions

prop.table(missing_pattern_table)
```

```{r}
#| label: cleanup-sch-ref

rm(sch_ref)
```

## Section 504 Enrollment

```{r}
#| label: compute-504-enrollment

enrollment <- targets::tar_read("popcounts_y2122")

enrollment_504_summary <- enrollment %>%
  select(!matches("TOT_")) %>%
  select(matches("_504")) |>
  select(1:2) |>
  summarize_all(~ sum(crdc_sub(.x), na.rm = TRUE))
```

```{r}
#| label: display-504-enrollment

enrollment_504_summary
```

```{r}
#| label: cleanup-enrollment

rm(enrollment)
```

## Grade Level Restrictions

```{r}
#| label: compute-grade-restrictions

combined_data <- targets::tar_read("combined_model_data") |> filter(YEAR == "21-22")
combined_data2 <- combined_data |> filter(highest_grade_offered >= 7)

grade_restriction_summary <- combined_data |>
  filter(YEAR == "21-22") |>
  filter(RACE == "TOTAL") |>
  filter(SEX == "TOTAL") |>
  mutate(hs = ifelse(highest_grade_offered >= 7, "Yes", "No")) |>
  group_by(hs) |>
  summarize(enrollment = sum(enrollment, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-grade-restrictions

grade_restriction_summary
```

## CCD Matching

```{r}
#| label: compute-ccd-matching

ccd <- targets::tar_read("ccd_sch_geo_y2122")
crdc <- targets::tar_read("full_crdc_data_y2122")

ccd$COMBOKEY <- stringr::str_pad(ccd$ncessch_num, width = 12,
                                side = "left", pad = "0")
ccd <- ccd |> select(COMBOKEY, highest_grade_offered, lowest_grade_offered,
                     latitude, longitude, enrollment)
```

```{r}
#| label: display-ccd-match-test

match_test(crdc$COMBOKEY, ccd$COMBOKEY)
```

```{r}
#| label: compute-ccd-match-summary

ccd_match_summary <- crdc |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  ungroup() |>
  mutate(in_ccd = ifelse(COMBOKEY %in% unique(ccd$COMBOKEY), "Yes", "No")) |>
  group_by(in_ccd) |>
  summarize(enrollment = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-ccd-match-summary

ccd_match_summary
```

```{r}
#| label: cleanup-ccd-crdc

rm(ccd, crdc)
gc()
```

## Enrollment Thresholds

```{r}
#| label: compute-enrollment-thresholds

combined_data <- targets::tar_read("combined_model_data") |> filter(YEAR == "21-22")
combined_data <- crdc_lea_collapse(combined_data)

enrollment_threshold_summary <- combined_data |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  mutate(filtered_out = ifelse(total_enroll < 30 | stu_enroll == 0, "Yes", "No")) |>
  group_by(filtered_out) |>
  summarize(dists = n(),
            enrollment = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-enrollment-thresholds

enrollment_threshold_summary
```

## Incomplete Enrollments

```{r}
#| label: compute-incomplete-enrollments

enrollment <- targets::tar_read("popcounts_y2122")
zz <- incomplete_enrollments(enrollment, year = "21-22")
zb <- targets::tar_read("schenrollraw_y2122")

incomplete_proportion <- nrow(zz) / sum(nrow(zz), nrow(zb))
```

```{r}
#| label: display-incomplete-enrollments

incomplete_proportion
```

# Model Predictions and Performance

## Observed Data Preparation

```{r}
#| label: compute-observed-data

obsv_data <- rdata |>
  filter(YEAR == "21-22") |>
  filter(stu_enroll > 0) |>
  mutate(phat = (2+ARRESTS) / (4+stu_enroll)) |> # agresti-coull approximation
  mutate(phat_se = sqrt((phat * (1-phat)/(4+stu_enroll)))) |>
  mutate(sd = phat_se * (4 + stu_enroll)) |>
  mutate(fitted_value = ARRESTS) |>
  mutate(CVp = 100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))) |>
  # agresti-coull confidence interval for arrests
  mutate(ci_upper = (phat + (1.96*phat_se)) * (4 + stu_enroll),
        ci_lower = (phat - (1.96*phat_se)) * (4 + stu_enroll)) |>
  # add rule of three correction
  mutate(ci_upper = ifelse(ARRESTS == 0, 3, ci_upper),
        ci_lower = ifelse(ARRESTS == 0, 0, ci_lower)) |>
  mutate(ci_lower = ifelse(ci_lower < 0, 0, ci_lower)) |>
  mutate(model_id = "observed")
```

## Model Predictions

```{r}
#| label: get-model-predictions

dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", YEAR = "21-22")
dist_test <- dist_test |> mutate(CVp = 100 * (sd / fitted_value))
dist_test <- dist_test |> select(model_id, LEAID, LEA_STATE, YEAR, RACE, SEX,
                                  fitted_value, sd, CVp, ci_lower, ci_upper)
```

## Prediction Adjustments

```{r}
#| label: join-predictions

plotdf <- inner_join(
  dist_test,
  obsv_data |>
    select(LEA_STATE:stu_enroll, sd, CVp, ci_upper, ci_lower) |>
    rename(obsv_sd = sd, obsv_CVP = CVp, obs_upper = ci_upper, obs_lower = ci_lower),
  by = join_by(LEAID, LEA_STATE, RACE, SEX)
)
```

```{r}
#| label: adjust-zero-variance

plotdf <- plotdf |>
  mutate(
    fitted_constant = ifelse(sd == 0, 1, 0),
    sd = ifelse(sd == 0, {
      phat <- (fitted_value + 2) / (stu_enroll+4)
      phat_se <- sqrt((phat * (1-phat)/(4+stu_enroll)))
      phat_se * (4 + stu_enroll)
    }, sd)
  ) |>
  mutate(CVp = 100 * (sd / fitted_value)) |>
  mutate(CVp = ifelse(!is.finite(CVp), {
    phat <- (fitted_value + 2) / (stu_enroll+4)
    100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))
  }, CVp))
```

## Performance Metrics

```{r}
#| label: compute-performance-metrics

plotdf <- plotdf |>
  mutate(
    fitted_value = fitted_value / (stu_enroll/1000),
    sd = sd / (stu_enroll/1000),
    obsv_sd = obsv_sd / (stu_enroll / 1000),
    ci_lower = ci_lower / (stu_enroll / 1000),
    ci_upper = ci_upper / (stu_enroll/1000),
    arrest_rate = ARRESTS / (stu_enroll / 1000),
    obs_upper = obs_upper / (stu_enroll / 1000),
    obs_lower = obs_lower / (stu_enroll / 1000)
  ) |>
  mutate(
    covered = if_else(ci_lower <= arrest_rate & ci_upper >= arrest_rate, 1, 0),
    improved_cv = ifelse(CVp < obsv_CVP & is.finite(obsv_CVP), 1, 0),
    obsv_precision = 1 / obsv_sd^2,
    fit_precision = 1 / sd^2
  ) |>
  mutate(
    improved_precision = ifelse(obsv_precision > fit_precision, 0, 1),
    narrower_interval = ifelse((ci_upper - ci_lower) < (obs_upper - obs_lower), 1, 0)
  )
```

## Overall Model Summary

```{r}
#| label: compute-overall-summary

overall_model_summary <- plotdf |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  ) |>
  select(-improved_interval, -improved_precision, -covered)
```

```{r}
#| label: display-overall-summary

overall_model_summary
```

# Model Performance Diagnostics

## Cross-tabulations

```{r}
#| label: crosstab-interval-precision

table(plotdf$narrower_interval, plotdf$improved_precision)
```

```{r}
#| label: crosstab-interval-cv

table(plotdf$narrower_interval, plotdf$improved_cv)
```

## Spot Checks

```{r}
#| label: spotcheck-narrower-intervals

plotdf |>
  filter(narrower_interval == 1) |>
  sample_n(10)
```

```{r}
#| label: spotcheck-narrower-positive

plotdf |>
  filter(narrower_interval == 1 & ci_upper > 0) |>
  sample_n(10)
```

```{r}
#| label: spotcheck-narrower-nonconstant

plotdf |>
  filter(narrower_interval == 1 & ci_upper > 0 & fitted_constant == 0) |>
  sample_n(10)
```

```{r}
#| label: spotcheck-narrower-constant

plotdf |>
  filter(narrower_interval == 1 & fitted_constant == 1) |>
  sample_n(10)
```

```{r}
#| label: crosstab-nonconstant

table(plotdf$narrower_interval[plotdf$fitted_constant == 0],
      plotdf$improved_precision[plotdf$fitted_constant == 0])
```

```{r}
#| label: constant-proportions

table(plotdf$fitted_constant) |> prop.table()
```

## Non-Constant Models

```{r}
#| label: compute-nonconstant-summary

nonconstant_summary <- plotdf |>
  filter(fitted_constant == 0) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  ) |>
  select(-meanCVP, -meanfitCVP, -improved_cv, -improved_precision, -improved_interval)
```

```{r}
#| label: display-nonconstant-summary

nonconstant_summary
```

## Non-Zero Arrests

```{r}
#| label: compute-nonzero-summary

nonzero_summary <- plotdf |>
  filter(ARRESTS > 0) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    interval_delta = median((((obs_upper - obs_lower) - (ci_upper - ci_lower))) /
                            (obs_upper - obs_lower)),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  ) |>
  select(-meanCVP, -meanfitCVP, -improved_cv, -improved_precision, -improved_interval)
```

```{r}
#| label: display-nonzero-summary

nonzero_summary
```

```{r}
#| label: load-model-objects

sg_m4 <- targets::tar_read("sg_m2_mod")
nat_m2 <- targets::tar_read("nat_m2_mod")
```

# Subgroup Analysis - Large Districts

## Define Subgroups

```{r}
#| label: define-big100

big100 <- rdata |>
  group_by(LEA_STATE, LEAID, LEA_NAME) |>
  summarize(
    total_enroll = sum(stu_enroll),
    total_arrests = sum(ARRESTS)
  ) |>
  arrange(desc(total_enroll)) |>
  ungroup() |>
  slice_head(n = 100)
```

```{r}
#| label: define-big0

big0 <- rdata |>
  group_by(LEA_STATE, LEAID, LEA_NAME) |>
  summarize(
    total_enroll = sum(stu_enroll),
    total_arrests = sum(ARRESTS)
  ) |>
  filter(total_arrests == 0) |>
  arrange(desc(total_enroll)) |>
  ungroup() |>
  slice_head(n = 100)
```

```{r}
#| label: define-bigarresters

bigarresters <- rdata |>
  group_by(LEA_STATE, LEAID, LEA_NAME) |>
  summarize(
    total_enroll = sum(stu_enroll),
    total_arrests = sum(ARRESTS)
  ) |>
  arrange(desc(total_arrests)) |>
  ungroup() |>
  slice_head(n = 100)
```

## Large Districts Performance

```{r}
#| label: compute-big100-summary

big100_summary <- plotdf |>
  filter(LEAID %in% big100$LEAID) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    interval_delta = median((((obs_upper - obs_lower) - (ci_upper - ci_lower))) /
                            (obs_upper - obs_lower)),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  )
```

```{r}
#| label: display-big100-summary

big100_summary
```

## High-Arrest Districts

```{r}
#| label: compute-bigarresters-summary

bigarresters_summary <- plotdf |>
  filter(LEAID %in% bigarresters$LEAID) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    interval_delta = median((((obs_upper - obs_lower) - (ci_upper - ci_lower))) /
                            (obs_upper - obs_lower)),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  )
```

```{r}
#| label: display-bigarresters-summary

bigarresters_summary
```

## Zero-Arrest Districts

```{r}
#| label: compute-big0-expected

expected_rate <- mean(big100$total_arrests / big100$total_enroll)

big0_summary <- plotdf |>
  filter(LEAID %in% big0$LEAID) |>
  ungroup() |>
  group_by(LEAID, LEA_STATE, YEAR, model_id, RACE, SEX) |>
  mutate(expected_value = expected_rate * stu_enroll) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    model_fitted = sum(fitted_value),
    model_min = sum(ci_lower),
    model_max = sum(ci_upper),
    expected = sum(expected_value)
  )
```

```{r}
#| label: display-big0-summary

big0_summary
```

```{r}
#| label: get-big0-draws

big0_draws <- get_prediction_draws(con, LEAID = unique(big0$LEAID), YEAR = "21-22")
```

```{r}
#| label: compute-big0-draws-summary

big0_draws_summary <- big0_draws |>
  group_by(YEAR, model_id, draw_id) |>
  summarize(fit = sum(pred)) |>
  group_by(YEAR, model_id) |>
  summarize(
    med = median(fit),
    low = quantile(fit, 0.025),
    hi = quantile(fit, 0.975)
  )
```

```{r}
#| label: display-big0-draws-summary

big0_draws_summary
```

# Model Computation Statistics

## National Models

```{r}
#| label: compute-nat-m1-stats

calculate_model_stats(targets::tar_read("nat_m1_mod"))
```

```{r}
#| label: compute-nat-m2-stats

calculate_model_stats(targets::tar_read("nat_m2_mod"))
```

```{r}
#| label: compute-nat-m3-stats

calculate_model_stats(targets::tar_read("nat_m3_mod"))
```

```{r}
#| label: compute-nat-m4-stats

calculate_model_stats(targets::tar_read("nat_m4_mod"))
```

```{r}
#| label: compute-nat-m5-stats

calculate_model_stats(targets::tar_read("nat_m5_mod"))
```

## Subgroup Models

```{r}
#| label: compute-sg-m1-stats

sg_m1_stats <- calculate_model_stats(targets::tar_read("sg_m1_mod")) |>
  summarize(
    runtime = sum(runtime_minutes),
    parameters = sum(parameters),
    data_rows = sum(data_rows),
    ndraws = sum(ndraws)
  )
```

```{r}
#| label: display-sg-m1-stats

sg_m1_stats
```

```{r}
#| label: compute-sg-m2-stats

sg_m2_stats <- calculate_model_stats(targets::tar_read("sg_m2_mod")) |>
  summarize(
    runtime = sum(runtime_minutes),
    parameters = sum(parameters),
    data_rows = sum(data_rows),
    ndraws = sum(ndraws)
  )
```

```{r}
#| label: display-sg-m2-stats

sg_m2_stats
```

```{r}
#| label: compute-sg-m3-stats

sg_m3_stats <- calculate_model_stats(targets::tar_read("sg_m3_mod")) |>
  summarize(
    runtime = sum(runtime_minutes),
    parameters = sum(parameters),
    data_rows = sum(data_rows),
    ndraws = sum(ndraws)
  )
```

```{r}
#| label: display-sg-m3-stats

sg_m3_stats
```

```{r}
#| label: compute-sg-m4-stats

sg_m4_stats <- calculate_model_stats(targets::tar_read("sg_m4_mod")) |>
  summarize(
    runtime = sum(runtime_minutes),
    parameters = sum(parameters),
    data_rows = sum(data_rows),
    ndraws = sum(ndraws)
  )
```

```{r}
#| label: display-sg-m4-stats

sg_m4_stats
```

```{r}
#| label: compute-sg-m5-stats

sg_m5_stats <- calculate_model_stats(targets::tar_read("sg_m5_mod")) |>
  summarize(
    runtime = sum(runtime_minutes),
    parameters = sum(parameters),
    data_rows = sum(data_rows),
    ndraws = sum(ndraws)
  )
```

```{r}
#| label: display-sg-m5-stats

sg_m5_stats
```

## HMC Diagnostics

```{r}
#| label: hmc-diag-nat-m1

rstan::check_hmc_diagnostics(targets::tar_read("nat_m1_mod")$fit)
```

```{r}
#| label: hmc-diag-nat-m2

rstan::check_hmc_diagnostics(targets::tar_read("nat_m2_mod")$fit)
```

```{r}
#| label: hmc-diag-nat-m3

rstan::check_hmc_diagnostics(targets::tar_read("nat_m3_mod")$fit)
```

```{r}
#| label: hmc-diag-nat-m4

rstan::check_hmc_diagnostics(targets::tar_read("nat_m4_mod")$fit)
```

```{r}
#| label: hmc-diag-nat-m5

rstan::check_hmc_diagnostics(targets::tar_read("nat_m5_mod")$fit)
```

# State-Level Analysis

## State Predictions

```{r}
#| label: get-state-predictions

state_test <- get_state_prediction_summary(con)
state_test <- state_test |> mutate(CVp = 100 * (sd / fitted_value))
state_test <- state_test |> select(model_id, LEA_STATE, YEAR, RACE, SEX,
                                    fitted_value, sd, CVp, ci_lower, ci_upper)
```

## State Observed Data

```{r}
#| label: compute-state-observed

state_obsv_data <- rdata |>
  filter(YEAR == "21-22") |>
  group_by(LEA_STATE, YEAR, RACE, SEX) |>
  summarize(
    ARRESTS = sum(ARRESTS),
    stu_enroll = sum(stu_enroll)
  ) |>
  mutate(phat = (2+ARRESTS) / (4+stu_enroll)) |>
  mutate(phat_se = sqrt((phat * (1-phat)/(4+stu_enroll)))) |>
  mutate(sd = phat_se * (4+stu_enroll)) |>
  mutate(fitted_value = ARRESTS) |>
  mutate(CVp = 100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))) |>
  mutate(
    ci_upper = (phat + (1.96*phat_se)) * stu_enroll,
    ci_lower = (phat - (1.96*phat_se)) * stu_enroll
  ) |>
  mutate(
    ci_upper = ifelse(ARRESTS == 0, 3/stu_enroll, ci_upper),
    ci_lower = ifelse(ARRESTS == 0, 0, ci_lower)
  ) |>
  mutate(model_id = "observed")
```

## State Performance

```{r}
#| label: compute-state-performance

state_plotdf <- left_join(
  state_test,
  state_obsv_data |>
    select(LEA_STATE:stu_enroll, sd, CVp, ci_upper, ci_lower) |>
    rename(obsv_sd = sd, obsv_CVP = CVp, obs_upper = ci_upper, obs_lower = ci_lower),
  by = join_by(LEA_STATE, YEAR, RACE, SEX)
)

state_performance <- state_plotdf |>
  filter(YEAR == "21-22") |>
  mutate(
    fitted_value = fitted_value / (stu_enroll/1000),
    sd = sd / (stu_enroll/1000),
    obsv_sd = obsv_sd / (stu_enroll / 1000),
    ci_lower = ci_lower / (stu_enroll / 1000),
    ci_upper = ci_upper / (stu_enroll/1000),
    arrest_rate = ARRESTS / (stu_enroll / 1000),
    obs_upper = obs_upper / (stu_enroll / 1000),
    obs_lower = obs_lower / (stu_enroll / 1000)
  ) |>
  mutate(
    covered = if_else(ci_lower <= arrest_rate & ci_upper >= arrest_rate, 1, 0),
    improved_cv = ifelse(CVp < obsv_CVP & is.finite(obsv_CVP), 1, 0),
    obsv_precision = 1 / obsv_sd^2,
    fit_precision = 1 / sd^2
  ) |>
  mutate(improved_precision = ifelse(obsv_precision > fit_precision, 0, 1)) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_improved_pre = improved_precision / count
  )
```

```{r}
#| label: display-state-performance

state_performance
```
