---
title: "CRDC Arrest Analysis Results"
format:
  html:
    embed-resources: true
knitr:
  opts_chunk:
    message: false
    warning: false
    error: false
    echo: false
    fig.path: export/figures/results-
execute:
  warning: false
  message: false
  error: false
  fig-width: 10
  fig-height: 6
---

# Introduction

This script takes the completed pipeline data analysis and produces the key
statistics, tables, and figures in the final report. This is the code that
was used to generate the results in the final paper.

# Setup and Data Loading

```{r}
#| label: setup
#| message: false
#| warning: false

library(targets)
library(brms)
library(dplyr)
library(ggplot2)
library(ggdist)
library(marginaleffects)
library(tidybayes)
library(ggridges)
library(duckdb)
library(knitr)

# Create export/figures directory if it doesn't exist
if (!dir.exists("export/figures")) {
  dir.create("export/figures", recursive = TRUE)
}
```

```{r}
#| label: load-data
#| message: false
#| warning: false

# Source the artifact-reproduction helpers + project functions
source("R/funs.R")
source("R/crdc_path.R"); source("R/model_registry.R"); source("R/paper_figures.R")
cv_apply_branding()   # civilytics theme on all figures (+ logo when magick present)

# Connect to a `predicted_draws` VIEW over the PUBLISHED draws parquet.
# CRDC_ARTIFACTS selects the source: local "export" (owner) or hf:// (stranger).
# get_prediction_summary()/get_state_prediction_summary() work unchanged on `con`.
h <- open_draws_view(); con <- h$con

# Load model-input data from the published staged artifacts (not the targets store)
tydata <- read_stage_df("stages/inputs/three_year_data.parquet")
rdata  <- read_stage_df("stages/inputs/recent_data.parquet")
```

# Descriptive Statistics from Most Recent CRDC Wave

These are the statistics for the most recent wave of data in the analysis.

## Overall Totals

Total students arrested:

```{r}
#| label: display-total-arrests

y2122 <- read_stage_df("stages/crdc/full_crdc_data_y2122.parquet")

# Total students arrested
sum(y2122$ARRESTS[y2122$RACE == "TOTAL" & y2122$SEX=="TOTAL"])
```

Total student enrollment:
```{r}
#| label: display-total-enrollment

# Enrollment
sum(y2122$stu_enroll[y2122$RACE == "TOTAL" & y2122$SEX=="TOTAL"])
```

Student arrest rate:

```{r}
#| label: display-arrest-rate

sum(y2122$ARRESTS[y2122$RACE == "TOTAL" & y2122$SEX=="TOTAL"]) /
  (sum(y2122$stu_enroll[y2122$RACE == "TOTAL" & y2122$SEX=="TOTAL"]) / 1000)
```

## Disaggregation By Race

Calculate and produce disaggregated arrest data by student race.

```{r}
#| label: compute-race-totals

nat_sg_totals_race <- y2122 |>
    ungroup() |>
    filter(RACE != "TOTAL" & SEX != "TOTAL") |>
    group_by(RACE) |>
    summarize(
        lea_count = n_distinct(LEAID),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000),
            referral_rate = referrals / (enrollment/1000))
```

```{r}
#| label: display-race-totals

nat_sg_totals_race |>
    filter(RACE %in% c("AM", "WH", "HI", "BL")) |>
    knitr::kable()
```

## Disaggregation By Sex

```{r}
#| label: compute-sex-totals

nat_sg_totals_sex <- y2122 |>
    ungroup() |>
    filter(RACE != "TOTAL" & SEX != "TOTAL") |>
    group_by(SEX) |>
    summarize(
        lea_count = n_distinct(LEAID),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000),
            referral_rate = referrals / (enrollment/1000))
```

```{r}
#| label: display-sex-totals

nat_sg_totals_sex |> knitr::kable()
```

## National summary statistics

```{r}
#| label: compute-district-totals

nat_totals <- y2122 |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS))

nat_referral_rate <- nat_totals$referrals / (nat_totals$enrollment / 1000)
nat_arrest_rate <- nat_totals$arrests / (nat_totals$enrollment / 1000)
```

```{r}
#| label: display-district-totals

nat_totals |> knitr::kable()
```

Arrest rates

```{r}
nat_referral_rate
nat_arrest_rate
```

# Sample Continuity Analysis for CRDC Waves

Look at how many LEAs are in multiple years of data and how many LEAs drop
out of the data when we add successive waves of the data.

```{r}
#| label: load-historical-data

y1718 <- read_stage_df("stages/crdc/full_crdc_data_y1718.parquet")
y1516 <- read_stage_df("stages/crdc/full_crdc_data_y1516.parquet")
```

```{r}
#| label: match-test-1718-2122

match_test(y1718$LEAID, y2122$LEAID)
```

```{r}
#| label: match-test-1516-1718

match_test(y1516$LEAID, y1718$LEAID[y1718$LEAID %in% unique(y2122$LEAID)])
```

# District Concentration Analysis

For this analysis we look at the most recent CRDC wave.

## District Size Categories

```{r}
#| label: compute-district-categories

thresh <- c(0, 1000, 10000, 20000, Inf)
labs <- c("0-999", "1,000-9,999", "10,000-19,999","20,000+")

dist_ref_arr <- y2122 %>%
  group_by(YEAR, LEA_STATE, LEAID, RACE, SEX) %>%
  summarize(ARRESTS = sum(ARRESTS),
            REFERRALS = sum(REFERRALS),
            stu_enroll = sum(stu_enroll),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = ARRESTS / (stu_enroll / 1000),
         referral_rate = REFERRALS / (stu_enroll / 1000)) %>%
  group_by(YEAR, LEA_STATE, LEAID) |>
  mutate(total_enroll = sum(stu_enroll[RACE != "TOTAL" &
                                         SEX != "TOTAL"]))
```

## Expected vs Observed Arrests

```{r}
#| label: compute-expected-arrests

dist_size_summary <- dist_ref_arr %>%
   ungroup() %>%
     filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
   mutate(popcut = cut(stu_enroll,
                       breaks = thresh,
                       labels = labs,
                       include.lowest= TRUE)) %>%
   group_by(popcut) %>%
 mutate(expec_arrest = pbinom(1, stu_enroll, nat_arrest_rate/1000,
                              lower.tail = FALSE)) %>%
   summarize(dists = n(),
             dists_w_arrests = sum(ARRESTS>0),
             expec_distw_arrests = round(sum(expec_arrest),1)) %>%
   mutate(dist_w_arrest_per = pretty_per(dists_w_arrests/dists),
          expect_w_arrest_per = pretty_per(expec_distw_arrests/dists))
```

```{r}
#| label: display-expected-arrests

dist_size_summary |> knitr::kable()
```

## Concentration Visualization

This visualization shows the relative concentration of arrests compared in LEAs
compared to the concentration of population in LEAs.
```{r}
#| label: compute-concentration-data

total_enroll_national <- sum(dist_ref_arr$total_enroll[
  dist_ref_arr$RACE == "TOTAL" & dist_ref_arr$SEX == "TOTAL"
])

concentration_plot_data <- dist_ref_arr %>%
  ungroup %>%
  filter(ARRESTS > 0) %>%
  filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
  arrange(desc(stu_enroll)) %>%
  mutate(total_arr = sum(ARRESTS)) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_enroll_national) %>%
  mutate(dist_per_cum = cumsum(dist_per))
```

```{r}
#| label: plot-concentration
#| fig-cap: "Share of total arrests by cumulative district enrollment"
#| fig-width: 10
#| fig-height: 6

concentration_plot <- concentration_plot_data %>%
  ggplot(aes(x = dist_per_cum, y = cum_sum_per)) +
  geom_line() +
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5, linetype = 2) +
  scale_x_continuous("% of national student population",
                     labels = scales::percent, limits = c(0, 0.7)) +
  scale_y_continuous("% of all arrests", labels = scales::percent) +
  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE) +
  labs(title = "Share of total arrests by cumulative district enrollment",
       subtitle = "Horizontal line at 50% of all arrests, vertical line at 50% of all students") +
  theme(axis.ticks.x = element_blank(),
        panel.grid = element_blank())

print(concentration_plot)

# Save high-resolution figure
ggsave("export/figures/arrest_concentration_by_district_enrollment.png",
       plot = concentration_plot,
       width = 10, height = 6, dpi = 300, bg = "white")
```

## Enrollment by Arrest Status

```{r}
#| label: compute-enrollment-by-arrest

enrollment_by_arrest <- dist_ref_arr %>%
  ungroup %>%
  filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
  mutate(has_arrest = ifelse(ARRESTS > 0, "Arrests", "No arrests")) |>
  group_by(has_arrest) |>
  summarize(enrollment = sum(total_enroll)) |>
  mutate(enrollment_per = enrollment / total_enroll_national)

enrollment_by_arrest |> knitr::kable()
```


# Sample Restriction Impact

Here we tabulate the figures about sample restrictions in the paper.

## Section 504 Data

```{r}
#| label: compute-504-data

sch_ref <- read_stage_df("stages/crdc/lerefs_y2122.parquet")

section_504_summary <- sch_ref %>%
  select(!matches("TOT_")) %>%
  select(matches("_504_")) |>
  ungroup() |>
  summarize_all(~ sum(crdc_sub(.x), na.rm = TRUE))
```

```{r}
#| label: display-504-data

section_504_summary |> knitr::kable()
```

## Missing Data Patterns

Law enforcement referral missingness data.

```{r}
#| label: compute-missing-patterns

# Drop non COMBOKEY identifier columns
sch_ref <- sch_ref %>% select(COMBOKEY:last_col())
names(sch_ref) <- gsub("_IDEA_", "_", names(sch_ref))

# Check for reserve values
missing_pattern_table <- sch_ref %>%
  select(!matches("TOT_")) %>%
  select(!matches("_504_")) %>%
  select(!matches("_LEP_")) %>%
  select(!matches("_EL_")) %>%
  tidyr::pivot_longer(cols = matches("SCH_DIS"), names_prefix = "SCH_",
              names_to = c("DISAB", "REF_TYPE", "RACE", "SEX"),
              names_sep = "_") %>%
  group_by(COMBOKEY, DISAB, REF_TYPE, RACE, SEX) |>
  mutate(value_mod = ifelse(value < 0, TRUE, FALSE)) |>
  pull(value_mod) |>
  table()
```

```{r}
#| label: display-missing-patterns

missing_pattern_table
```

```{r}
#| label: display-missing-proportions

prop.table(missing_pattern_table)
```

```{r}
#| label: cleanup-sch-ref

rm(sch_ref)
```

## Section 504 Enrollment

```{r}
#| label: compute-504-enrollment

enrollment <- read_stage_df("stages/crdc/popcounts_y2122.parquet")

enrollment_504_summary <- enrollment %>%
  select(!matches("TOT_")) %>%
  select(matches("_504")) |>
  select(1:2) |>
  summarize_all(~ sum(crdc_sub(.x), na.rm = TRUE))
```

```{r}
#| label: display-504-enrollment

enrollment_504_summary |> knitr::kable()
```

```{r}
#| label: cleanup-enrollment

rm(enrollment)
```

## Grade Level Restrictions

This reports the impact of restricting the analysis to schools that offer
a grade 7 or above.

```{r}
#| label: compute-grade-restrictions

combined_data <- read_stage_df("stages/inputs/combined_model_data.parquet") |> filter(YEAR == "21-22")
combined_data2 <- combined_data |> filter(highest_grade_offered >= 7)

grade_restriction_summary <- combined_data |>
  filter(YEAR == "21-22") |>
  filter(RACE == "TOTAL") |>
  filter(SEX == "TOTAL") |>
  mutate(hs = ifelse(highest_grade_offered >= 7, "Yes", "No")) |>
  group_by(hs) |>
  summarize(enrollment = sum(enrollment, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-grade-restrictions

grade_restriction_summary |> knitr::kable()
```

## CCD Matching

By matching to CCD data we further restrict the CRDC universe. This analysis
shows the impact of that.

```{r}
#| label: compute-ccd-matching

ccd <- read_stage_df("stages/crdc/ccd_sch_geo_y2122.parquet")
crdc <- read_stage_df("stages/crdc/full_crdc_data_y2122.parquet")

ccd$COMBOKEY <- stringr::str_pad(ccd$ncessch_num, width = 12,
                                side = "left", pad = "0")
ccd <- ccd |> select(COMBOKEY, highest_grade_offered, lowest_grade_offered,
                     latitude, longitude, enrollment)
```

```{r}
#| label: display-ccd-match-test

match_test(crdc$COMBOKEY, ccd$COMBOKEY)
```

```{r}
#| label: compute-ccd-match-summary

ccd_match_summary <- crdc |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  ungroup() |>
  mutate(in_ccd = ifelse(COMBOKEY %in% unique(ccd$COMBOKEY), "Yes", "No")) |>
  group_by(in_ccd) |>
  summarize(enrollment = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-ccd-match-summary

ccd_match_summary |> knitr::kable()
```

```{r}
#| label: cleanup-ccd-crdc

rm(ccd, crdc)
gc()
```

## Enrollment Thresholds

We restrict the sample by enrollment size, and this reports the impact of that
restriction.

```{r}
#| label: compute-enrollment-thresholds

combined_data <- read_stage_df("stages/inputs/combined_model_data.parquet") |> filter(YEAR == "21-22")
combined_data <- crdc_lea_collapse(combined_data)

enrollment_threshold_summary <- combined_data |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  mutate(filtered_out = ifelse(total_enroll < 30 | stu_enroll == 0, "Yes", "No")) |>
  group_by(filtered_out) |>
  summarize(dists = n(),
            enrollment = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-enrollment-thresholds

enrollment_threshold_summary |> knitr::kable()
```

## Censored or truncated enrollment data

Here we report the impact of suppressed records on enrollment.


```{r}
#| label: compute-incomplete-enrollments

enrollment <- read_stage_df("stages/crdc/popcounts_y2122.parquet")
zz <- incomplete_enrollments(enrollment, year = "21-22")
zb <- read_stage_df("stages/crdc/schenrollraw_y2122.parquet")

incomplete_proportion <- nrow(zz) / sum(nrow(zz), nrow(zb))
incomplete_proportion
```

# Model Predictions and Performance

Now we calculate the performance results comparing the models to non-model
methods for computing arrest rates.

## Observed Data Preparation

```{r}
#| label: compute-observed-data

obsv_data <- rdata |>
  filter(YEAR == "21-22") |>
  filter(stu_enroll > 0) |>
  mutate(phat = (2+ARRESTS) / (4+stu_enroll)) |> # agresti-coull approximation
  mutate(phat_se = sqrt((phat * (1-phat)/(4+stu_enroll)))) |>
  mutate(sd = phat_se * (4 + stu_enroll)) |>
  mutate(fitted_value = ARRESTS) |>
  mutate(CVp = 100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))) |>
  # agresti-coull confidence interval for arrests
  mutate(ci_upper = (phat + (1.96*phat_se)) * (4 + stu_enroll),
        ci_lower = (phat - (1.96*phat_se)) * (4 + stu_enroll)) |>
  # add rule of three correction
  mutate(ci_upper = ifelse(ARRESTS == 0, 3, ci_upper),
        ci_lower = ifelse(ARRESTS == 0, 0, ci_lower)) |>
  mutate(ci_lower = ifelse(ci_lower < 0, 0, ci_lower)) |>
  mutate(model_id = "observed")

dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", YEAR = "21-22")
dist_test <- dist_test |> mutate(CVp = 100 * (sd / fitted_value))
dist_test <- dist_test |> select(model_id, LEAID, LEA_STATE, YEAR, RACE, SEX,
                                  fitted_value, sd, CVp, ci_lower, ci_upper)

plotdf <- inner_join(
  dist_test,
  obsv_data |>
    select(LEA_STATE:stu_enroll, sd, CVp, ci_upper, ci_lower) |>
    rename(obsv_sd = sd, obsv_CVP = CVp, obs_upper = ci_upper, obs_lower = ci_lower),
  by = join_by(LEAID, LEA_STATE, RACE, SEX)
)

plotdf <- plotdf |>
  mutate(
    fitted_constant = ifelse(sd == 0, 1, 0),
    sd = ifelse(sd == 0, {
      phat <- (fitted_value + 2) / (stu_enroll+4)
      phat_se <- sqrt((phat * (1-phat)/(4+stu_enroll)))
      phat_se * (4 + stu_enroll)
    }, sd)
  ) |>
  mutate(CVp = 100 * (sd / fitted_value)) |>
  mutate(CVp = ifelse(!is.finite(CVp), {
    phat <- (fitted_value + 2) / (stu_enroll+4)
    100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))
  }, CVp))
```

## Performance Metrics

This code calculates our performance metrics on the arrest rate scale, translating
it from the arrest scale.

```{r}
#| label: compute-performance-metrics

plotdf <- plotdf |>
  mutate(
    fitted_value = fitted_value / (stu_enroll/1000),
    sd = sd / (stu_enroll/1000),
    obsv_sd = obsv_sd / (stu_enroll / 1000),
    ci_lower = ci_lower / (stu_enroll / 1000),
    ci_upper = ci_upper / (stu_enroll/1000),
    arrest_rate = ARRESTS / (stu_enroll / 1000),
    obs_upper = obs_upper / (stu_enroll / 1000),
    obs_lower = obs_lower / (stu_enroll / 1000)
  ) |>
  mutate(
    covered = if_else(ci_lower <= arrest_rate & ci_upper >= arrest_rate, 1, 0),
    improved_cv = ifelse(CVp < obsv_CVP & is.finite(obsv_CVP), 1, 0),
    obsv_precision = 1 / obsv_sd^2,
    fit_precision = 1 / sd^2
  ) |>
  mutate(
    improved_precision = ifelse(obsv_precision > fit_precision, 0, 1),
    narrower_interval = ifelse((ci_upper - ci_lower) <= (obs_upper - obs_lower), 1, 0)  # "equal or better" (matches white_paper)
  )
```

## Overall Model Summary

The paper reviews how the models do compared to a counterfactual of applying
non-model based uncertainty intervals to rate statistics.

```{r}
#| label: compute-overall-summary

overall_model_summary <- plotdf |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  ) |>
  select(-improved_interval, -improved_precision, -covered)
```

```{r}
#| label: display-overall-summary

overall_model_summary |> knitr::kable()
```

# Model Performance Diagnostics

## Comparing models to non-modeled intervals


```{r}
#| label: crosstab-interval-precision

table(narrower_interval = plotdf$narrower_interval,
      improved_precision =plotdf$improved_precision)
```

```{r}
#| label: crosstab-interval-cv

table(narrower_interval = plotdf$narrower_interval,
      improved_precision = plotdf$improved_cv)
```

## Looking at constant models

Some fitted values are constants with no variation. We want to look at these
cases separately.


```{r}
#| label: constant-proportions

table(fitted_constant = plotdf$fitted_constant) |> prop.table() |> knitr::kable()

```

When we exclude constant prediction results and see how models perform to
non-modeled intervals.

```{r}
#| label: crosstab-nonconstant

table(narrower = plotdf$narrower_interval[plotdf$fitted_constant == 0],
      higher_precsion = plotdf$improved_precision[plotdf$fitted_constant == 0])
```


## Non-Constant Models

Now we look in more depth at results from observations without constant
predicted values.

```{r}
#| label: compute-nonconstant-summary

nonconstant_summary <- plotdf |>
  filter(fitted_constant == 0) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  ) |>
  select(-meanCVP, -meanfitCVP, -improved_cv, -improved_precision, -improved_interval)
```

```{r}
#| label: display-nonconstant-summary

nonconstant_summary |> knitr::kable()
```

## Non-Zero Arrests

Now we look at results for the subset of records for which at least one
arrest is observed.

```{r}
#| label: compute-nonzero-summary

nonzero_summary <- plotdf |>
  filter(ARRESTS > 0) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    interval_delta = median((((obs_upper - obs_lower) - (ci_upper - ci_lower))) /
                            (obs_upper - obs_lower)),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  ) |>
  select(-meanCVP, -meanfitCVP, -improved_cv, -improved_precision, -improved_interval)
```

```{r}
#| label: display-nonzero-summary

nonzero_summary |> knitr::kable()
```

# Subgroup Analysis

## Define Subgroups

Create largest 100 districts

```{r}
#| label: define-big100

big100 <- rdata |>
  group_by(LEA_STATE, LEAID, LEA_NAME) |>
  summarize(
    total_enroll = sum(stu_enroll),
    total_arrests = sum(ARRESTS)
  ) |>
  arrange(desc(total_enroll)) |>
  ungroup() |>
  slice_head(n = 100)
```

Create largest 100 districts with 0 arrests

```{r}
#| label: define-big0

big0 <- rdata |>
  group_by(LEA_STATE, LEAID, LEA_NAME) |>
  summarize(
    total_enroll = sum(stu_enroll),
    total_arrests = sum(ARRESTS)
  ) |>
  filter(total_arrests == 0) |>
  arrange(desc(total_enroll)) |>
  ungroup() |>
  slice_head(n = 100)
```

Create districts with most arrests

```{r}
#| label: define-bigarresters

bigarresters <- rdata |>
  group_by(LEA_STATE, LEAID, LEA_NAME) |>
  summarize(
    total_enroll = sum(stu_enroll),
    total_arrests = sum(ARRESTS)
  ) |>
  arrange(desc(total_arrests)) |>
  ungroup() |>
  slice_head(n = 100)
```

## Large Districts Performance

Look at how models perform on the 100 largest districts by student enrollment

```{r}
#| label: compute-big100-summary

big100_summary <- plotdf |>
  filter(LEAID %in% big100$LEAID) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    interval_delta = median((((obs_upper - obs_lower) - (ci_upper - ci_lower))) /
                            (obs_upper - obs_lower)),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  )
```

```{r}
#| label: display-big100-summary

big100_summary |> knitr::kable()
```

## High-Arrest Districts

```{r}
#| label: compute-bigarresters-summary

bigarresters_summary <- plotdf |>
  filter(LEAID %in% bigarresters$LEAID) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    obsv_interval_med = median((obs_upper - obs_lower)),
    fit_interval_med = median((ci_upper - ci_lower)),
    interval_delta = median((((obs_upper - obs_lower) - (ci_upper - ci_lower))) /
                            (obs_upper - obs_lower)),
    improved_precision = sum(improved_precision),
    improved_interval = sum(narrower_interval),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_narrower = improved_interval / count,
    per_improved_pre = improved_precision / count
  )
```

```{r}
#| label: display-bigarresters-summary

bigarresters_summary |> knitr::kable()
```

## Zero-Arrest Districts

```{r}
#| label: compute-big0-expected

expected_rate <- mean(big100$total_arrests / big100$total_enroll)

big0_summary <- plotdf |>
  filter(LEAID %in% big0$LEAID) |>
  ungroup() |>
  group_by(LEAID, LEA_STATE, YEAR, model_id, RACE, SEX) |>
  mutate(expected_value = expected_rate * stu_enroll) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    model_fitted = sum(fitted_value),
    model_min = sum(ci_lower),
    model_max = sum(ci_upper),
    expected = sum(expected_value)
  )
```

```{r}
#| label: display-big0-summary

big0_summary |> knitr::kable()
```

```{r}
#| label: get-big0-draws

big0_draws <- get_prediction_draws(con, LEAID = unique(big0$LEAID), YEAR = "21-22")
```

```{r}
#| label: compute-big0-draws-summary

big0_draws_summary <- big0_draws |>
  group_by(YEAR, model_id, draw_id) |>
  summarize(fit = sum(pred)) |>
  group_by(YEAR, model_id) |>
  summarize(
    med = median(fit),
    low = quantile(fit, 0.025),
    hi = quantile(fit, 0.975)
  )
```

```{r}
#| label: display-big0-draws-summary

big0_draws_summary |> knitr::kable()
```

# Model Computation Statistics

Here we compute the model statistics.

## National Models

```{r}
#| label: compute-nat-model-stats

# Pooled-model stats from the published diagnostics artifact (precomputed
# calculate_model_stats(); labels via the model registry).
read_stage_df("stages/diagnostics/model_stats.parquet") |>
  filter(model_id %in% crdc_pooled_ids()) |>
  with_model_labels() |>
  transmute(Model = model_label, ndraws, chains, threads, thin, leas,
            parameters, worst_rhat, min_bulk_ess, mean_bulk_ess,
            runtime_minutes, data_rows) |>
  knitr::kable()

```


## Subgroup Models

```{r}
#| label: compute-sg-stats

# Student-group model stats from the published diagnostics artifact: each
# sg_m#_mod stored its per-group rows; sum them per model (matches the original
# disaggregated_*_avg aggregation), labelled via the registry.
read_stage_df("stages/diagnostics/model_stats.parquet") |>
  filter(grepl("^sg_", model_id)) |>
  with_model_labels() |>
  group_by(model_label) |>
  summarize(
    runtime    = sum(runtime_minutes),
    parameters = sum(parameters),
    data_rows  = sum(data_rows),
    ndraws     = sum(ndraws),
    .groups = "drop"
  )
```

## HMC Diagnostics

```{r}
#| label: hmc-diagnostics

# Live pooled-model HMC diagnostics when the fits are available locally/cached
# (stages/models/pooled_m#.qs2); otherwise show the published diagnostics table.
# crdc_cached_path() never downloads, so this won't force a ~2.9 GB pull.
reg <- crdc_model_registry()
any_live <- FALSE
for (id in crdc_pooled_ids()) {
  spec <- reg$spec[match(id, reg$id)]
  f <- crdc_cached_path(sprintf("stages/models/pooled_%s.qs2", spec))
  if (!is.na(f)) {
    any_live <- TRUE
    cat("\n###", crdc_model_label(id), "\n")
    rstan::check_hmc_diagnostics(qs2::qs_read(f)$fit)
  }
}
if (!any_live) {
  cat("Pooled fits not present locally; showing the published diagnostics table.\n")
  knitr::kable(read_stage_df("stages/diagnostics/hmc_diagnostics.parquet"))
}
```

# State-Level Analysis

## State Predictions

```{r}
rm(big0_draws, dist_ref_arr, dist_test, enrollment,
      obsv_data, combined_data, combined_data2, plotdf); gc()
```

```{r}
#| label: get-state-predictions

state_test <- get_state_prediction_summary(con)
state_test <- state_test |> mutate(CVp = 100 * (sd / fitted_value))
state_test <- state_test |> select(model_id, LEA_STATE, YEAR, RACE, SEX,
                                    fitted_value, sd, CVp, ci_lower, ci_upper)
```

## State Observed Data

```{r}
#| label: compute-state-observed

state_obsv_data <- rdata |>
  filter(YEAR == "21-22") |>
  group_by(LEA_STATE, YEAR, RACE, SEX) |>
  summarize(
    ARRESTS = sum(ARRESTS),
    stu_enroll = sum(stu_enroll)
  ) |>
  mutate(phat = (2+ARRESTS) / (4+stu_enroll)) |>
  mutate(phat_se = sqrt((phat * (1-phat)/(4+stu_enroll)))) |>
  mutate(sd = phat_se * (4+stu_enroll)) |>
  mutate(fitted_value = ARRESTS) |>
  mutate(CVp = 100 * sqrt(((1-phat)/(phat*(4+stu_enroll))))) |>
  mutate(
    ci_upper = (phat + (1.96*phat_se)) * stu_enroll,
    ci_lower = (phat - (1.96*phat_se)) * stu_enroll
  ) |>
  mutate(
    ci_upper = ifelse(ARRESTS == 0, 3/stu_enroll, ci_upper),
    ci_lower = ifelse(ARRESTS == 0, 0, ci_lower)
  ) |>
  mutate(model_id = "observed")
```

## State Performance

```{r}
#| label: compute-state-performance

state_plotdf <- left_join(
  state_test,
  state_obsv_data |>
    select(LEA_STATE:stu_enroll, sd, CVp, ci_upper, ci_lower) |>
    rename(obsv_sd = sd, obsv_CVP = CVp, obs_upper = ci_upper, obs_lower = ci_lower),
  by = join_by(LEA_STATE, YEAR, RACE, SEX)
)

state_performance <- state_plotdf |>
  filter(YEAR == "21-22") |>
  mutate(
    fitted_value = fitted_value / (stu_enroll/1000),
    sd = sd / (stu_enroll/1000),
    obsv_sd = obsv_sd / (stu_enroll / 1000),
    ci_lower = ci_lower / (stu_enroll / 1000),
    ci_upper = ci_upper / (stu_enroll/1000),
    arrest_rate = ARRESTS / (stu_enroll / 1000),
    obs_upper = obs_upper / (stu_enroll / 1000),
    obs_lower = obs_lower / (stu_enroll / 1000)
  ) |>
  mutate(
    covered = if_else(ci_lower <= arrest_rate & ci_upper >= arrest_rate, 1, 0),
    improved_cv = ifelse(CVp < obsv_CVP & is.finite(obsv_CVP), 1, 0),
    obsv_precision = 1 / obsv_sd^2,
    fit_precision = 1 / sd^2
  ) |>
  mutate(improved_precision = ifelse(obsv_precision > fit_precision, 0, 1)) |>
  group_by(model_id, YEAR) |>
  summarize(
    count = n(),
    meanCVP = mean(obsv_CVP[is.finite(obsv_CVP)], na.rm = TRUE),
    meanfitCVP = mean(CVp[is.finite(CVp)], na.rm = TRUE),
    obsv_precision = mean(obsv_precision),
    fit_precision = mean(fit_precision),
    improved_cv = sum(improved_cv),
    improved_precision = sum(improved_precision),
    covered = sum(covered)
  ) |>
  mutate(
    per_covered = covered/count,
    per_improved_cv = improved_cv / count,
    per_improved_pre = improved_precision / count
  )
```

```{r}
#| label: display-state-performance

state_performance |> knitr::kable()
```

```{r}
#| label: cleanup
#| include: false

close_draws_view(h)
```
