---
title: "CRDC Arrest Social Media Post Results"
format:
  html:
    embed-resources: true
knitr:
  opts_chunk:
    message: false
    warning: false
    error: false
    echo: false
    fig.path: export/figures/socialmedia-
    dev: ragg_png
    dpi: 300
execute:
  freeze: auto
  warning: false
  message: false
  error: false
  fig-width: 11
  fig-height: 7
---

# CRDC Social Media Posts

```{r}
#| label: setup
#| message: false
#| warning: false

library(targets)
library(brms)
library(dplyr)
library(ggplot2)
library(ggdist)
library(marginaleffects)
library(tidybayes)
library(ggridges)
library(duckdb)
library(knitr)
library(civilytics)

```

 ## Read Data
```{r}
#| label: load-data
#| message: false
#| warning: false

# Source custom functions + artifact-reproduction helpers
source("R/funs.R")
source("R/spatial_utils.R")
source("R/crdc_path.R"); source("R/model_registry.R"); source("R/paper_figures.R")

# predicted_draws VIEW over the PUBLISHED draws parquet (CRDC_ARTIFACTS selects
# local "export" for the owner, hf:// for a new user). Existing
# get_prediction_summary(con, ...) calls work unchanged.
h <- open_draws_view(); con <- h$con

# Load model-input data from the published staged artifacts
tydata <- read_stage_df("stages/inputs/three_year_data.parquet")
rdata  <- read_stage_df("stages/inputs/recent_data.parquet")
ccd_ty <- read_stage_df("stages/inputs/combined_sch_data.parquet")
crdc   <- read_stage_df("stages/crdc/full_crdc_data_y2122.parquet")

# Logo overlay helper
logo_file <- system.file("img", "civilytics-wordmark.png", package = "civilytics")

# Output dimensions (must match ragg_png/knitr fig.width & fig.height)
fig_w <- 11  # inches
fig_h <- 7   # inches
dpi   <- 300

# Logo target size in inches → convert to plot fractions
logo_w_in <- 1.5   # how wide you want the logo
logo_h_in <- 0.375  # match your logo's aspect ratio (~4:1)
margin_in <- 0.05  # gap from edge

# add_logo_overlay <- function(p,
#                               side = "right",
#                               vertical = "bottom",
#                              fig_w    = 12,
#                              fig_h    = 8,
#                              logo_w_in = 1.5,
#                              logo_h_in = 0.375,
#                              margin_in = 0.15) {
#   logo_grob <- png::readPNG(logo_file) |> grid::rasterGrob(interpolate = TRUE)

#   # Convert inches → plot fractions
#   w <- logo_w_in / fig_w
#   h <- logo_h_in / fig_h
#   if ( side == "right") {
#     x <- 1 - w - (margin_in / fig_w)   # right-anchored
#   } else {
#     x <- 0.05
#  }
#      if(vertical == "bottom") {
#     y <- 0 #(margin_in / fig_h)
#   } else {
#     y <- 1 - (margin_in / fig_h)
#   }

#   cowplot::ggdraw(p) +
#     cowplot::draw_grob(logo_grob, x = x, y = y, width = w, height = h)
# }

# Add colors
color_fill_values <- c("WH" = "#c92d0e", "BL" = "#020684",
                        "AM" = "#ff8f43",
                       "TOTAL" = "#909090ff", "HI" = "#0791b6")
color_fill_values2 <- c("White" = "#c92d0e", "Black" = "#020684",
                        "Total" = "#909090ff", "Hispanic" = "#0791b6",
                        "Amer. Ind." = "#ff8f43")


model_palette <- c('Modeled' ="#000a9bff", 'Frequentist' = "#858585bb")



agresti_coull <- function(numerator, denominator, confidence_level = 0.95) {
    adj_star <- qnorm(1 - (1-confidence_level)/2)
    # critical value
    if (numerator > 0) {
        num_star <- numerator + adj_star
        denom_star <- denominator + (2 * adj_star)
        phat <- num_star / denom_star
        phat_se <- sqrt((phat/denom_star) * (1-phat))
        ci_upper <- (phat + (adj_star * phat_se)) * denom_star
        ci_lower <- (phat - (adj_star * phat_se)) * denom_star
        sd <- mean(c(ci_upper, ci_lower)) / confidence_level
        sd <- phat_se * denom_star
    } else {

        ci_upper <- denominator*(-log(1-confidence_level)/denominator)
        ci_lower <- 0
        sd <- mean(c(ci_upper, ci_lower)) / confidence_level
        phat_se <- 0
        phat <- mean(c(ci_upper, ci_lower))
    }
    return(c(ci_upper, ci_lower, sd, phat_se, phat))
}

# let's wrap model types in two dimensions

add_model_time <- function(x) {
    y <- rep("One year", length(x))
    y[x %in% c("unified_m3_mod", "unified_m4_mod", "unified_m5_mod")] <- "Three year"
    y[x %in% c("stratified_m3_mod", "stratified_m4_mod", "stratified_m5_mod")] <- "Three year"
    return(y)
}

add_model_form <- function(x) {
    y <- rep("Stratified", length(x))
    y[startsWith(x, "unified")] <- "Unified"
    return(y)
}

add_model_cov <- function(x) {
  y <- rep("Baseline", length(x))
  y[x %in% c("stratified_m2_mod", "unified_m4_mod", "unified_m5_mod", "unified_m2_mod", "stratified_m4_mod", "stratified_m5_mod" )] <- "Covariate"
  return(y)
}


print_model_name <- function(x, lbreak = FALSE) {
    y <- rep("Stratified", length(x))
    y[startsWith(x, "unified")] <- "Unified"
    x <- gsub("_mod", "", x)
    x <- gsub("unified_m", "", x)
    x <- gsub("stratified_m", "", x)
    if (lbreak) {
      y <- paste0(y, "\nModel:", x)
    }  else {
      y <- paste0(y, " Model:", x)

    }

    return(y)
}


```

<!-- DRIFT: series-intro-para -->
We were fortunate to receive a grant from the **American Educational Research Association** and the **National Science Foundation (NSF)** two years ago (and to receive a no-cost extension to complete the work in 2025!). We wrote the grant to use small area estimation strategies to "get the most" from the Civil Rights Data Collection (CRDC) school arrest data. We wanted to share findings from the grant with others and decided on a series of post we'll share over the next few weeks. We'll start with basic descriptive information about arrests (by state, by race, by district) before moving to the "getting the most" piece of the work and then will conclude with some reflections on what will be lost if national data collections like this don't continue.
<!-- /DRIFT -->

<!-- DRIFT: aera-nsf-disclaimer -->
For all posts, please know: *This research was supported by a grant from the American Educational Research Association which receives funds for its "AERA Grants Program" from the National Science Foundation under NSF award NSF-DRL #1749275. Opinions reflect those of the author and do not necessarily reflect those AERA or NSF.*
<!-- /DRIFT -->

## Post 1: State and National Totals

::: {.smpost slug="how-states-stack-up" title="Arrests in schools in 2021-22: How states stack up on student arrests" date="2025-02-01" status="gold" series="1"}

```{r}
#| label: calc_nat_totals
#| message: false
#| warning: false


dist_ref_arr <- crdc %>%
  group_by(YEAR,LEA_STATE, LEAID, RACE, SEX) %>%
  summarize(ARRESTS = sum(ARRESTS),
            REFERRALS = sum(REFERRALS),
            stu_enroll = sum(stu_enroll),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = ARRESTS / (stu_enroll / 1000),
         referral_rate = REFERRALS / (stu_enroll / 1000)) %>%
  group_by(YEAR, LEA_STATE, LEAID) |>
  mutate(total_enroll = sum(stu_enroll[RACE != "TOTAL" &
                                         SEX != "TOTAL"]))

state_map <- tigris::states(cb  = TRUE, resolution = "20m") |> tigris::shift_geometry()

dist_ref_arr |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) -> nat_totals

nat_referral_rate <- nat_totals$referrals / (nat_totals$enrollment / 1000)
nat_arrest_rate <- nat_totals$arrests / (nat_totals$enrollment / 1000)
nat_totals
```



```{r}
#| label: calc_state_totals
#| message: false
#| warning: false

dist_ref_arr |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEA_STATE) |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000)) |>
    mutate(arrest_nnh = enrollment / arrests) -> state_totals

state_map2 <- inner_join(state_map, state_totals, by = join_by(STUSPS == LEA_STATE))
```

<!-- DRIFT: post1-intro-paras -->
The Civil Rights Data Collection (CRDC) is the only source of data on school-related arrests for all the nation's public schools and districts. The most recent available data, released in January 2025, are for the 2021-22 school year.

Nearly 35,000 students were arrested in "school-related arrests" in 2021-22 (34,846 students). "School-related arrests" are defined as those taking place on school grounds, while students are taking school transportation, or due to a referral by a school official.

How many students were arrested "at school/in a school-related arrest" in each state?
<!-- /DRIFT -->

```{r}
#| label: arrest_chloropleth
#| message: false
#| warning: false

p1 <- ggplot(data = state_map2, aes(fill = arrests)) +
  geom_sf() +
  label_states(state_map2, "arrests", label_fn = civilytics::pretty_count,
              fill = "white", size = 3.5) +
  scale_fill_distiller("Arrests", palette = "YlOrRd", direction = 1, labels = scales::comma) +
  labs(title = "Students arrested at school by state", subtitle = "34,846 school-based arrests in 2021-22", caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22") +
  theme_civilytics_map() +
  theme(legend.text = element_text(size = 8),
        legend.position = "inside",
        legend.position.inside = c(0.9, 0.32))

  # theme_void() +
  # theme(
  #   #plot.margin = rep(unit(6, "pt"), 4),
  #       plot.background = element_rect(fill = "#FAFAF7", color = NA),
  #       legend.position.inside = c(0.9, 0.32),
  #       legend.position = "inside",
  #       legend.text = element_text(size = 8))




grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-right"))

```

<!-- DRIFT: post1-rate-intro -->
Of course, states have vastly different student populations, motivating questions about rates. The national school-related arrest rate in 2021-22 was 0.72 out of every 1,000 students. Which states had higher or lower school-related arrest rates?
<!-- /DRIFT -->

```{r}
#| label: arrest_rate_chloropleth
#| message: false
#| warning: false


p1 <- ggplot(data = state_map2, aes( fill = arrest_rate)) +
  geom_sf() +
  label_states(state_map2, "arrest_rate", label_fn = function(x) round(x, 2), size = 3.5,
    fill = "white") +
  scale_fill_distiller("Arrest rate", palette = "YlOrRd", direction = 1) +
   labs(title = "Arrest rate at school by state", subtitle = "Arrests per 1,000 students enrolled", caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22") +
  theme_civilytics_map() +
  theme(
    #plot.margin = rep(unit(6, "pt"), 4),
        plot.background = element_rect(fill = "#FAFAF7", color = NA),
        legend.position.inside = c(0.9, 0.32),
        legend.position = "inside",
        legend.text = element_text(size = 8))


grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-right"))
#plot(civilytics::add_logo(p1, civilytics::make_logo_grob(), margin_param = -12))


#rm(state_map2)
```

<!-- DRIFT: post1-nnh-intro -->
Another, perhaps more intuitive approach with small rates is to calculate the "number needed to harm" (NNH) or how many students it "takes" before 1 is arrested. Nationally, 1 out of every 1,395 students was arrested "at school" in 2021-22.^[Corrected from the originally posted "1 in 1,428." This number-needed-to-harm uses CRDC-reported enrollment (48,596,489 students ÷ 34,846 arrests = 1 in 1,395), the exact reciprocal of the 0.72-per-1,000 national rate from the same data. The original 1,428 reflected an external national-enrollment denominator.] In Kansas, 1 out of every 194 students was arrested in schools. In South Dakota, 1 out of every 348 students was arrested.
<!-- /DRIFT -->

```{r}
#| label: state_arrest_nnh_map
#| message: false
#| warning: false

dist_ref_arr |>
    ungroup() |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEA_STATE) |>
    summarize(
        all_lea_count = n_distinct(LEAID),
        lea_any_arr = n_distinct(LEAID[ARRESTS >0]),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000)) |>
    mutate(arrest_nnh = enrollment / arrests) -> state_totals

state_map2 <- inner_join(state_map, state_totals, by = join_by(STUSPS == LEA_STATE))

p1 <- ggplot(data = state_map2, aes( fill = arrest_nnh)) +
  geom_sf() +
  label_states(state_map2, "arrest_nnh", label_fn = pretty_count, size = 3.5,
    fill = "white") +
   scale_fill_distiller(palette = "YlOrRd", direction = -1,
  breaks = c(50000, 8000, 3000, 1000, 500),
  trans = "log", labels = scales::comma) +
  labs(title = "Students per arrested student by state",
  subtitle = "Number needed to harm (NNH)",
  caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22",
  fill = "NNH") +
  theme_void() +
  theme(
    #plot.margin = rep(unit(6, "pt"), 4),
        plot.background = element_rect(fill = "#FAFAF7", color = NA),
        legend.position.inside = c(0.9, 0.32),
        legend.position = "inside",
        legend.text = element_text(size = 8))

grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-left"))

#add_logo_overlay(p1, fig_w = 11, fig_h = 7, side = "left")
#plot(civilytics::add_logo(p1, civilytics::make_logo_grob(), margin_param = -12))

```

<!-- DRIFT: post1-closing-para -->
Part of what motivates this project is the challenges with rates and small Ns. Someone may reasonably think that the elevated arrest rates in South Dakota and Kansas are due at least in part to their much smaller populations (small Ns). While using rates rather than counts provides a sense of scale and standardization, rates' sensitivity to slight differences in counts (and to the denominator) magnifies differences that may not be substantively meaningful (sometimes called the base rate fallacy) and may imply a false precision in reported numbers. By using a combination of descriptive analysis and statistics (coming in future posts), we can more accurately convey true differences in school-based arrests that are consistent with the data collected. One way to help with this is to use multiple years of measurement to accumulate a greater population over which to measure the low incident rate. Our next post will look at how the addition of prior years of data changes, or doesn't change, this picture.
<!-- /DRIFT -->

:::

## Post 2: Trends

::: {.smpost slug="arrests-declining-everywhere" title="School-based arrests are declining but is this true everywhere?" date="2025-02-08" status="gold" series="2"}

<!-- DRIFT: post2-opening-para -->
Overall, the number of students with school-related arrests declined by \~44% in the last three CRDC data collection waves\*, from over 62,000 students arrested in 2015-16 to under 35,000 in 2021-22.
<!-- /DRIFT -->

```{r}
#| label: national_trend
#| message: false
#| warning: false

#combined_data <- read_stage_df("stages/inputs/combined_model_data.parquet")

# Unfiltered CRDC without requiring join to CCD
combined_data <- bind_rows(read_stage_df("stages/crdc/full_crdc_data_y2122.parquet"),
                            read_stage_df("stages/crdc/full_crdc_data_y1718.parquet"),
                            read_stage_df("stages/crdc/full_crdc_data_y1516.parquet"))


year_summary <- combined_data |>
  group_by(YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )

p1 <- ggplot(year_summary, aes(x = YEAR, y = total_arrests)) +
  geom_point() +
  geom_line(group = 1) +
  geom_text(aes(label = pretty_count(total_arrests)), nudge_x = 0.15) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma,
                  expand = expansion(mult = c(0, 0.1))) +
  labs(x = "CRDC Wave", y = "CRDC Arrests",
  title = "Arrest trends across CRDC waves",
  subtitle = "Total arrests nationally.",
  caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22, 2017-18, and 2015-16") +
  civilytics::theme_civilytics_dark_map()

grid::grid.draw(civilytics::civilytics_logo(p1, variant = "dark", position = "bottom-right"))


```


```{r}
#| label: national_arrest_rate_trend
#| message: false
#| warning: false


p1 <- ggplot(year_summary, aes(x = YEAR, y = total_arrests/(total_enrollment/1000))) +
  geom_point() +
  geom_line(group = 1) +
  geom_text(aes(label = paste0(round(total_arrests/(total_enrollment/1000), digits = 2), " arrests per \n1,000 students")), nudge_x = 0.15, vjust = 0, hjust = 0) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1)), labels = scales::comma) +
  labs(x = "CRDC Wave", y = "CRDC Arrests per 1,000", title = "Arrest rate trends across CRDC waves",
  caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22, 2017-18, and 2015-16") +
  civilytics::theme_civilytics(font_size =14)


grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-right"))
#plot(civilytics::add_logo(p1, civilytics::make_logo_grob(), margin_param = -12))


```

<!-- DRIFT: post2-ks-ca-intro -->
A big part of why we wanted to look at multiple years of data is not just to understand trends but because it's easy to dismiss one year of data as a fluke, a possible data error, or simply not reliable. For example, looking at multiple years of data allows us to ask whether Kansas just had more school-related arrests than California in 2021-22 (despite having a student population 1/12^th^ the size of CA) or if this is a consistent pattern across years?

In 2021-22 Kansas (allegedly) had 2,413 school-related arrests compared to 565 in 2017-18 and 521 in 2015-16.

In both of the prior years, Kansas did have arrest rates above the national average and above California's.[^post2-1] However, California, as expected, had many more total arrests than Kansas in both of those years.[^post2-2]
<!-- /DRIFT -->

```{r}
#| label: select_state_arrest_trend
#| message: false
#| warning: false

state_year_summary <- combined_data |>
  group_by(LEA_STATE, YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )


p1 <- ggplot(state_year_summary |> filter(LEA_STATE %in% c("CA", "KS")),
    aes(x = YEAR, y = total_arrests, group = LEA_STATE)) +
  geom_point() +
  geom_line(aes(linetype = LEA_STATE)) +
  annotate("text", x = 3.1, y = 2450, label = "Kansas") +
    annotate("text", x = 3.1, y = 1500, label = "California") +

   geom_text(aes(label = pretty_count(total_arrests)), nudge_y = 100, nudge_x = 0.1) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma,
  expand = expansion(mult = c(0, 0.1))) +
    guides(linetype = guide_none()) +
  labs(x = "CRDC Wave", y = "CRDC Arrests",
caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22, 2017-18, and 2015-16",
  title = "Comparing state arrest rate trends across CRDC waves", subtitle = "Kansas and California") +
  civilytics::theme_civilytics(font_size =14)


grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-right"))
#plot(civilytics::add_logo(p1, civilytics::make_logo_grob(), margin_param = -12))


```

<!-- DRIFT: post2-ks-conclusion -->
Taken together, the data indicate that Kansas does have a "higher than expected" school-based arrest rate and perhaps a more punitive approach to discipline than many states -- but that there were likely one or more data errors in Kansas's data for 2021-22.
<!-- /DRIFT -->

### Looking into Kansas

```{r}
#| label: kansas_lea
#| message: false
#| warning: false

combined_data |>
    group_by(LEA_STATE, LEA_NAME, LEAID, YEAR) |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    summarize(
        total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    JJ = sum(JJ == "Yes")
    ) -> lea_data

# TODO: join the LEA name for 21-22 onto the combined data to create a consistent
# name

recent_names <- lea_data |> filter(YEAR == "21-22") |> ungroup() |>
  select(LEAID, LEA_NAME) |> distinct_all()


lea_data <- left_join(lea_data |> ungroup() |> select(-LEA_NAME), recent_names,
                    by = join_by(LEAID))

```

<!-- DRIFT: post2-derby-para -->
Looking into those data more, 87% of Kansas's arrests in 2021-22 were (allegedly) from one district with \~6,500 students. Based on the reported data, the district (Derby Public Schools) would have an arrest rate of nearly 1 in 3 students in 2021-22. The district did have a very high reported arrest rate (1 in 63 students) in 2017-18 but not nearly as anomalous as in 2021-22.

Despite the [Office of Civil Rights](https://www.ed.gov/media/document/crdc-quality-informational-snapshot-january-2025-109165.pdf) taking steps before, during, and after data collection to ensure good data quality, with over 17,000 districts reporting hundreds of data elements, it's not feasible to identify and correct every possible data error. As we'll discuss in future posts, part of the motivation for this study was to use statistical methods to improve the quality of descriptive reporting on data particularly in the case of highly consequential but rare events like arrests.

People are rightly skeptical of the influence outliers or data errors can have on even state-level descriptive statistics when looking at relatively rare events like arrests. This skepticism limits the utility of a valuable data collection. We hope through statistical analysis we can explore the extent to which we can recover the utility of these data and increase confidence in using the numbers to inform the public about arrests in schools not just overall or in particular states but in districts as well.

\*A note on CRDC waves:

The CRDC is generally a biennial data collection. We examine data from the most recent year available, the 2021-22 school year, and from two prior collections, 2017-18 and 2015-16. There was also an off-cycle data collection in 2020-21 (due to the pandemic, the planned 2019-20 data collection was postponed). In 2020-21, only about a quarter as many arrests (9,738) and referrals (65,312) were reported as in 2021-22. We do not include 2020-21 data in these analyses because students' and schools' experiences during that time varied greatly depending on whether instruction was virtual, hybrid, or in-person along with other factors.
<!-- /DRIFT -->

[^post2-1]: 2017-18: KS: 1.2 arrests per 1,000, national: 1.03, CA: 0.3; 2015-16: KS: 1.1 arrests per 1,000, national: 1.23, CA: 0.5

[^post2-2]: 2017-18: 2,151 vs. 565, 2015-16: 3,424 vs. 521

:::


## Districts

::: {.smpost slug="districts-most-arrests" title="Some districts arrest hundreds of students and have arrest rates 10x the national average" date="2025-02-15" status="gold" series="3"}

<!-- DRIFT: post3-opening-para -->
We wrote this grant in large part because we wanted to examine district arrest rates in more detail. The district level is often the appropriate site of authority for the policies and procedures leading to school arrests and is where individual and collective actors (parents, students, school staff, and elected officials) can have the most immediate, direct, and demonstrable effect. Unfortunately, it's also where errors or anomalies are most obvious, often leading folks to dismiss what can be learned from data.

In 2021-22, 11.6% of districts reported at least one arrest. Let's look at which districts reported the most students arrested at "school" (i.e., with school-related arrests) in 2021-22. Because of the data error concern, let's also look at what these districts reported in previous waves of the CRDC.
<!-- /DRIFT -->

```{r}
#| label: national_lea_table
#| message: false
#| warning: false

library(flextable)

img_path <- "export/figures/socialmedia-national_lea_table.png"



table_data <- lea_data |>
  mutate(YEAR = dplyr::recode_values(YEAR,
    "21-22" ~ "2021_22",
    "15-16" ~ "2015_16",
    "17-18" ~ "2017_18")) |>
  group_by(LEAID) |>
  tidyr::pivot_wider(id_cols = c(LEA_STATE, LEAID, LEA_NAME),
    values_from = c(total_enrollment, total_arrests), names_from = YEAR) |>
  mutate(arrest_rate_2021_22 = total_arrests_2021_22 / (total_enrollment_2021_22 / 1000)) |>
  arrange(desc(total_arrests_2021_22)) |>
  ungroup() |>
  select(LEA_STATE, LEA_NAME, total_enrollment_2021_22,
    total_arrests_2021_22, arrest_rate_2021_22,
    total_arrests_2015_16, total_arrests_2017_18) |>
  slice_head(n = 20) |>
  mutate(
    LEA_NAME                 = stringr::str_to_title(LEA_NAME),
    total_enrollment_2021_22 = scales::comma(total_enrollment_2021_22, accuracy = 1),
    total_arrests_2021_22    = scales::comma(total_arrests_2021_22, accuracy = 1),
    arrest_rate_2021_22      = scales::comma(arrest_rate_2021_22, accuracy = 0.1),
    total_arrests_2015_16    = scales::comma(total_arrests_2015_16, accuracy = 1),
    total_arrests_2017_18    = scales::comma(total_arrests_2017_18, accuracy = 1)
  )

header_bg <- "#2c3e50"
header_fg <- "white"
alt_bg    <- "#f0f0eb"

ft <- flextable(table_data) |>
  set_header_labels(
    LEA_STATE                = "State",
    LEA_NAME                 = "District",
    total_enrollment_2021_22 = "Enrollment\n(2021-22)",
    total_arrests_2021_22    = "Arrests\n(2021-22)",
    arrest_rate_2021_22      = "Arrest Rate\n(per 1,000)",
    total_arrests_2015_16    = "Arrests\n(2015-16)",
    total_arrests_2017_18    = "Arrests\n(2017-18)"
  ) |>
  add_header_row(
    values    = c("", "", "", "", "Historical Arrests"),
    colwidths = c(1, 1, 1, 1, 3),
    top       = TRUE
  ) |>
  add_header_lines("Top 20 School Districts by Student Arrests — CRDC 2021-22") |>
  bg(bg = header_bg, part = "header") |>
  color(color = header_fg, part = "header") |>
  bold(part = "header") |>
  fontsize(i = 1, size = 14, part = "header") |>
  fontsize(size = 11, part = "body") |>
  bg(i = seq(2, 20, 2), bg = alt_bg, part = "body") |>
  font(fontname = "Arial", part = "all") |>
  align(j = c("LEA_STATE", "LEA_NAME"), align = "left",  part = "body") |>
  align(j = c("total_enrollment_2021_22", "total_arrests_2021_22",
              "arrest_rate_2021_22", "total_arrests_2015_16",
              "total_arrests_2017_18"), align = "right", part = "body") |>
  align(part = "header", align = "center") |>
  width(j = "LEA_STATE",                width = 0.55) |>
  width(j = "LEA_NAME",                 width = 2.4)  |>
  width(j = "total_enrollment_2021_22", width = 1.1)  |>
  width(j = "total_arrests_2021_22",    width = 1.0)  |>
  width(j = "arrest_rate_2021_22",      width = 1.05) |>
  width(j = "total_arrests_2015_16",    width = 1.0)  |>
  width(j = "total_arrests_2017_18",    width = 1.0)  |>
  border_outer(part = "all",  border = officer::fp_border(color = "#888888", width = 1)) |>
  border_inner_h(part = "body", border = officer::fp_border(color = "#cccccc", width = 0.5)) |>
  add_footer_lines("Source: U.S. Department of Education Civil Rights Data Collection") |>
  fontsize(size = 9, part = "footer") |>
  color(color = "#555555", part = "footer") |>
  set_table_properties(layout = "fixed")

# Export table as PNG
dims <- flextable_dim(ft)
ragg::agg_png(img_path,
  width = dims$width, height = dims$height + 0.4,  # extra space for logo
  units = "in", res = 300, scaling = 1
)
plot(ft)
dev.off()

# Composite logo onto saved PNG using magick
logo  <- magick::image_read(logo_file)
table_img <- magick::image_read(img_path)
img_w <- magick::image_info(table_img)$width
img_h <- magick::image_info(table_img)$height
logo_w <- round(img_w * 0.15)
logo   <- magick::image_scale(logo, paste0(logo_w, "x"))
logo_h <- magick::image_info(logo)$height
table_img <- magick::image_composite(
  table_img, logo,
  offset = paste0("+", img_w - logo_w - 20, "+", img_h - logo_h - 20)
)
magick::image_write(table_img, img_path, format = "png")

# Embed in rendered document
knitr::include_graphics(img_path)
```

<!-- DRIFT: post3-table-observations -->
The table highlights at least four things:

1.  There are wave-specific data errors for particular districts, such
    as the (implausibly) high number of arrests in DeKalb and Derby
    School Districts in 2021-22 and the alleged 0 arrests in Pinellas
    County, Anne Arundel County, NYC Public Schools, etc. in 2015-16.

2.  Some districts have pretty consistent numbers of students arrested
    at school each year (wave) -- for example, Pasadena (TX)
    consistently has around 400 students with school-based arrests,
    Jefferson Parish (LA) has around 200-300 students arrested, Rapid
    City (SD) consistently has around 200 students arrested.

3.  This is not just a list of the largest school districts. There were
    26 districts with \>100k students in 2021-22 but only 5 of these
    districts are in the top 20 here. Some districts that are not
    especially large -- like Waterbury School District in CT, Bibb
    County in GA, Chambersburg Area in PA -- have relatively large
    numbers of students arrested at school each year.

4.  Almost (but not) all of these districts have arrest rates above the
    national average. NYC and Miami-Dade have very large student
    populations and, thus, despite (seemingly) large numbers of students
    arrested, their rates are below the national average (0.72 per
    1,000). Most districts in this list arrested at least 5x as many
    students as the national average in 2021-22 though.

Some people might next ask: What are the top 20 districts by arrest
rates? Others might assume those are all very small districts (this is
true) but we will show in one of our next posts that we may still be
able to learn some things from districts with the highest arrest rates.

Other people may immediately ask: what are the demographics of these
districts? (or more bluntly: Do these districts primarily serve students
historically arrested at the highest rates?) We will also address this
in one of our next posts.
<!-- /DRIFT -->

:::

## High arrest rate districts and coefficient of variation

::: {.smpost slug="highest-arrest-rates" title="Districts with the highest arrest rates are small, serve special education students and Native Americans" date="2025-02-22" status="gold" series="4"}

<!-- DRIFT: post4-intro-paras -->
Another way to think about arrest rates is in terms of risk to
students--the risk of being arrested. Some serious offenses may lead to
arrests no matter what school students attend, but other actions -- like
fighting or disorderly conduct -- may be dealt with through referrals to
law enforcement at some schools and through school-based consequences at
others.

If you're a student trying to avoid an arrest, what's a particularly
risky district to attend?

It's not LAUSD or NYC. It might be Pekin Community High School District
or La-Salle Peru Township High School District (both in Illinois) or
Mingus Union High School District in Arizona. These districts
consistently have arrest rates 15 to 20x the national average. They
consistently arrest 20, 50, or even 100 students a year, despite having
fewer than 2,000 students in their districts.

By looking at the districts with the highest arrest rates we see that
that the risk isn't the greatest in districts with the most arrests
overall but in smaller districts -- some of which have abnormally high
arrest numbers.
<!-- /DRIFT -->

<!-- DRIFT: post4-table-observations -->
The data also show:

1.  Many very small districts with high arrest rates primarily serve
    students with special education needs. Research has long
    demonstrated that students with disabilities are disproportionately
    likely to be arrested (e.g., [GAO
    2024](https://www.gao.gov/products/gao-24-106294); [Prison Policy
    Initiative](https://www.prisonpolicy.org/research/disability/)).

2.  At least one of these districts serves almost exclusively American
    Indians. American Indian students have disproportionately high
    arrest rates ([GAO
    2024](https://www.gao.gov/products/gao-24-106294)).

3.  Several of these districts exclusively serve high school students.
    Of course, age is an extremely important factor in arrest rates.

4.  Data errors (or perhaps anomalies) are more obvious with smaller
    populations. For example, several schools on the list reported
    relatively large numbers of arrests one year but 0-2 in other years.

5.  These are mainly but not exclusively very small districts: 11 of the
    districts in the top 20 have \<300 students. Inherently, smaller
    districts will have greater variation in their rates.

> Tomorrow, we'll look more systematically at how enrollment size and
> arrest rates are related and how we can use the coefficient of
> variation, a measure of precision, to quantify this.
<!-- /DRIFT -->

```{r}
#| label: national_lea_table_higharrestrates
#| message: false
#| warning: false


img_path <- "export/figures/socialmedia-national_lea_table_higharrestrate.png"

table_data <- lea_data |>
  mutate(YEAR = dplyr::recode_values(YEAR,
    "21-22" ~ "2021_22",
    "15-16" ~ "2015_16",
    "17-18" ~ "2017_18")) |>
  mutate(arrest_rate = total_arrests / (total_enrollment/1000)) |>
  group_by(LEAID) |>
  tidyr::pivot_wider(id_cols = c(LEA_STATE, LEAID, LEA_NAME),
    values_from = c(total_enrollment, total_arrests, arrest_rate), names_from = YEAR) |>
  arrange(desc(arrest_rate_2021_22)) |>
  ungroup() |>
  select(LEA_STATE, LEA_NAME, arrest_rate_2021_22, arrest_rate_2017_18,
    arrest_rate_2015_16,
  total_enrollment_2021_22,
    total_arrests_2021_22, total_arrests_2017_18,
    total_arrests_2015_16) |>
  slice_head(n = 20) |>
  mutate(
    LEA_NAME                 = stringr::str_to_title(LEA_NAME),
    arrest_rate_2021_22      = scales::comma(arrest_rate_2021_22, accuracy = 0.1),
    arrest_rate_2017_18      = scales::comma(arrest_rate_2017_18, accuracy = 0.1),
    arrest_rate_2015_16      = scales::comma(arrest_rate_2015_16, accuracy = 0.1),
    total_enrollment_2021_22 = scales::comma(total_enrollment_2021_22, accuracy = 1),
    total_arrests_2021_22    = scales::comma(total_arrests_2021_22, accuracy = 1),
    total_arrests_2015_16    = scales::comma(total_arrests_2015_16, accuracy = 1),
    total_arrests_2017_18    = scales::comma(total_arrests_2017_18, accuracy = 1)
  )

header_bg <- "#2c3e50"
header_fg <- "white"
alt_bg    <- "#f0f0eb"

ft <- flextable(table_data) |>
  set_header_labels(
    LEA_STATE                = "State",
    LEA_NAME                 = "District",
    arrest_rate_2021_22      = "Rate\n(2021-22)",
    arrest_rate_2017_18      = "Rate\n(2017-18)",
    arrest_rate_2015_16      = "Rate\n(2015-16)",
    total_enrollment_2021_22 = "Enrollment\n(2021-22)",
    total_arrests_2021_22    = "Arrests\n(2021-22)",
    total_arrests_2017_18    = "Arrests\n(2017-18)",
    total_arrests_2015_16    = "Arrests\n(2015-16)"
  ) |>
  add_header_row(
    values    = c("", "", "Arrest Rate (per 1,000 students)", "Current Year", "Historical Arrests"),
    colwidths = c(1, 1, 3, 2, 2),
    top       = TRUE
  ) |>
  add_header_lines("Top 20 School Districts by Student Arrest Rate — CRDC 2021-22") |>
  bg(bg = header_bg, part = "header") |>
  color(color = header_fg, part = "header") |>
  bold(part = "header") |>
  fontsize(i = 1, size = 14, part = "header") |>
  fontsize(size = 11, part = "body") |>
  bg(i = seq(2, 20, 2), bg = alt_bg, part = "body") |>
  font(fontname = "Arial", part = "all") |>
  align(j = c("LEA_STATE", "LEA_NAME"), align = "left",  part = "body") |>
  align(j = c("arrest_rate_2021_22", "arrest_rate_2017_18", "arrest_rate_2015_16",
              "total_enrollment_2021_22", "total_arrests_2021_22",
              "total_arrests_2017_18", "total_arrests_2015_16"),
        align = "right", part = "body") |>
  align(part = "header", align = "center") |>
  width(j = "LEA_STATE",                width = 0.55) |>
  width(j = "LEA_NAME",                 width = 2.0)  |>
  width(j = "arrest_rate_2021_22",      width = 0.9)  |>
  width(j = "arrest_rate_2017_18",      width = 0.9)  |>
  width(j = "arrest_rate_2015_16",      width = 0.9)  |>
  width(j = "total_enrollment_2021_22", width = 1.0)  |>
  width(j = "total_arrests_2021_22",    width = 0.9)  |>
  width(j = "total_arrests_2017_18",    width = 0.9)  |>
  width(j = "total_arrests_2015_16",    width = 0.9)  |>
  border_outer(part = "all",  border = officer::fp_border(color = "#888888", width = 1)) |>
  border_inner_h(part = "body", border = officer::fp_border(color = "#cccccc", width = 0.5)) |>
  add_footer_lines("Source: U.S. Department of Education Civil Rights Data Collection") |>
  fontsize(size = 9, part = "footer") |>
  color(color = "#555555", part = "footer") |>
  set_table_properties(layout = "fixed")

# Export table as PNG
dims <- flextable_dim(ft)
ragg::agg_png(img_path,
  width = dims$width, height = dims$height + 0.4,  # extra space for logo
  units = "in", res = 300, scaling = 1
)
plot(ft)
dev.off()

# Composite logo onto saved PNG using magick
logo  <- magick::image_read(logo_file)
table_img <- magick::image_read(img_path)
img_w <- magick::image_info(table_img)$width
img_h <- magick::image_info(table_img)$height
logo_w <- round(img_w * 0.15)
logo   <- magick::image_scale(logo, paste0(logo_w, "x"))
logo_h <- magick::image_info(logo)$height
table_img <- magick::image_composite(
  table_img, logo,
  offset = paste0("+", img_w - logo_w - 20, "+", img_h - logo_h - 20)
)
magick::image_write(table_img, img_path, format = "png")

# Embed in rendered document
knitr::include_graphics(img_path)

```

:::

::: {.smpost slug="coefficient-of-variation" title="Basic facts — coefficient of variation" date="2025-03-01" status="gold" series="5"}

<!-- DRIFT: post5-cov-blockquote -->
> The coefficient of variation is a standardized metric for the relative
> precision of a measurement. The more imprecise a measure is, the more
> difficult it is to attribute differences among observations to true
> underlying differences as opposed to variation resulting from
> measurement alone. In the case of arrest rates, the limitations of
> time (a single school year) and population (the district enrollment)
> mean that the arrest rate as captured by the Civil Rights Data
> Collection (CRDC) may be limited in its ability to describe true
> differences in arrest patterns.
>
> We can see this three ways in the figure below.
>
> First, the coefficients of variation are higher for arrests than
> referrals -- school-based referrals to law enforcement are more
> frequent than school-based arrests, so there is less dispersion around
> their measurement.
>
> Second, the CV decreases for both referrals and arrests as the student
> enrollment decile increases (i.e., as districts are larger), but does
> so much more consistently for referrals.
>
> Third, student groups with smaller populations such as American Indian
> and Black students generally have less precise arrest rates than their
> counterparts in larger student groups.
>
> These disparities in precision were one of the initial motivators for
> us to look into statistical methods to improve the precision of arrest
> rates and allow for more meaningful district-student group-level
>
> comparisons.
<!-- /DRIFT -->

```{r}
#| label: cov_plot_quintile_total
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9
#| include: false
# exploratory: figure superseded by cov_plot_decile_byraceeth; chunk retained (include: false) because it builds shared tabdf/plotdf used below

#dist_ref_arr

# For clarity, let's look at total district size and then cut it and show CV for
# overall district size, and then for Black, White, and Hispanic students

dist_ref_arr %>%
  group_by(LEA_STATE, LEAID) %>%
  filter(stu_enroll > 0) |>
  mutate(total_enroll = sum(stu_enroll[SEX == "TOTAL" & RACE == "TOTAL"]),
         total_arrests = ARRESTS[SEX == "TOTAL" & RACE == "TOTAL"],
         total_referrals = REFERRALS[SEX == "TOTAL" & RACE == "TOTAL"]
  ) -> tabdf

total_enroll_quint <- quantile(tabdf$total_enroll[tabdf$SEX == "TOTAL" & tabdf$RACE == "TOTAL"],
         probs = seq(0, 1,0.2))

quint_labels <- c(
  paste0(format(total_enroll_quint[1], big.mark = ","), " - ",
         format(total_enroll_quint[2] - 1, big.mark = ",")),
  paste0(format(total_enroll_quint[2], big.mark = ","), " - ",
         format(total_enroll_quint[3] - 1, big.mark = ",")),
  paste0(format(total_enroll_quint[3], big.mark = ","), " - ",
         format(total_enroll_quint[4] - 1, big.mark = ",")),
  paste0(format(total_enroll_quint[4], big.mark = ","), " - ",
         format(total_enroll_quint[5] - 1, big.mark = ",")),
  paste0(">", format(total_enroll_quint[5], big.mark = ","))
)
# cut by enrollment quintiles

plotdf <- tabdf %>%
    ungroup() |>
  mutate(size = cut(total_enroll, total_enroll_quint, labels = quint_labels,
      ordered_result = TRUE)
  ) %>%
  #mutate(size = as.numeric(size)) %>%
  filter(SEX == "TOTAL") %>%
  filter(!is.na(size)) %>%
  group_by(size, RACE) %>%
  summarize(count = n(),
            all_enroll = sum(total_enroll),
            mean_arr_rate = mean(arrest_rate, na.rm =TRUE),
            sd_arr_rate = sd(arrest_rate, na.rm = TRUE),
            mean_ref_rate = mean(referral_rate, na.rm = TRUE),
            sd_ref_rate = sd(referral_rate, na.rm = TRUE)) %>%
  mutate(cov_arrests = sd_arr_rate / mean_arr_rate * 100,
         cov_referrals = sd_ref_rate / mean_ref_rate * 100)

plotdf$race_lab <- crdc_race_recode(plotdf$RACE)


p1 <- ggplot(plotdf[plotdf$RACE== "TOTAL",],
       aes(x = size, y = cov_arrests/100, color = race_lab, group = race_lab)) +
  geom_smooth(method = "loess", show.legend = FALSE) +
  geom_point(size = 4) +
  scale_color_manual(values = color_fill_values2) +
  scale_y_continuous(labels = scales::label_percent(big.mark = ","),
  limits = c(0, 30)) +
  civilytics::theme_civilytics(font_size = 14) +
  theme(legend.position = "bottom",
  axis.title.x = element_text(margin = margin(t = 15))) +
  labs(x = "Total Enrollment Quintile",
      color = "",
       title = "CV for Arrest Rates by Total Enrollment",
       y = "CV (% scale)")

grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-right"))

```

```{r}
#| label: cov_plot_decile_total
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9
#| include: false
# exploratory: figure superseded by cov_plot_decile_byraceeth; chunk retained (include: false) because it builds shared tabdf/plotdf used below

total_enroll_decile <- quantile(
  tabdf$total_enroll[tabdf$SEX == "TOTAL" & tabdf$RACE == "TOTAL"],
  probs = seq(0, 1, 0.1)
)

decile_labels <- c(
  sapply(1:9, function(i) {
    paste0(format(total_enroll_decile[i], big.mark = ","), " - ",
           format(total_enroll_decile[i + 1] - 1, big.mark = ","))
  }),
  paste0(">", format(total_enroll_decile[10], big.mark = ","))
)

plotdf <- tabdf %>%
  ungroup() |>
  mutate(size = cut(total_enroll, total_enroll_decile, labels = decile_labels,
      ordered_result = TRUE)
  ) %>%
  filter(SEX == "TOTAL") %>%
  filter(!is.na(size)) %>%
  group_by(size, RACE) %>%
  summarize(count = n(),
            all_enroll = sum(total_enroll),
            mean_arr_rate = mean(arrest_rate, na.rm = TRUE),
            sd_arr_rate = sd(arrest_rate, na.rm = TRUE),
            mean_ref_rate = mean(referral_rate, na.rm = TRUE),
            sd_ref_rate = sd(referral_rate, na.rm = TRUE)) %>%
  mutate(cov_arrests = sd_arr_rate / mean_arr_rate * 100,
         cov_referrals = sd_ref_rate / mean_ref_rate * 100)

plotdf$race_lab <- crdc_race_recode(plotdf$RACE)

p1 <- ggplot(plotdf[plotdf$RACE == "TOTAL",],
       aes(x = size, y = cov_arrests/100, color = race_lab, group = race_lab)) +
  geom_smooth(method = "loess", show.legend = FALSE, se=FALSE) +
  geom_point(size = 4) +
  scale_color_manual(values = color_fill_values2) +
  scale_y_continuous(labels = scales::label_percent(big.mark = ","),
    limits = c(0, 25), breaks = c(0, 5, 10, 15, 20, 25),
    expand = expansion(mult = c(0, 0.1))) +
  civilytics::theme_civilytics(font_size = 14) +
  theme(legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 15)),
    axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Total Enrollment Decile",
       color = "",
       title = "Arrest Rate Coefficient of Variation",
       subtitle = "By enrollment deciles",
       caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22",
       y = "CV (% scale)")

grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-right"))


```


```{r}
#| label: cov_plot_decile_byraceeth
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9



p1 <- ggplot(plotdf[plotdf$RACE %in% c("AM","BL", "WH", "TOTAL", "HI"),],
       aes(x = size, y = cov_arrests/100, color = race_lab, group = race_lab)) +
  geom_smooth(method = "loess", show.legend = FALSE, se = FALSE) +
  geom_point(size = 4) +
  scale_color_manual(values = color_fill_values2) +
  scale_y_continuous(labels = scales::label_percent(big.mark = ","),
  breaks = c(0, 5, 10, 15, 20, 30, 40),
  limits = c(0, 40), expand = expansion(mult = c(0, 0.1))) +
  civilytics::theme_civilytics(font_size = 12) +
  theme(legend.position = "bottom",
  axis.title.x = element_text(margin = margin(t = 15))) +
  labs(x = "Total Enrollment Decile",
       title = "Arrests",
       y = "CV (% scale)",
       color = "Race / Ethnicity")

  # #theme(legend.position = c(0.8, 0.7))+
  # labs(x = "Total Enrollment Quantile",
  #      y = "CV (% scale)",
  #      title = "Figure 2: Coefficient of Variation for Arrest Rates by District Population",
  #      color = "Race")


p2 <- ggplot(plotdf[plotdf$RACE %in% c("AM", "BL", "WH", "TOTAL", "HI"),],
       aes(x = size, y = cov_referrals/100, color = race_lab, group = race_lab)) +
  geom_smooth(method = "loess", show.legend = FALSE, se = FALSE) +
  geom_point(size = 4) +
  scale_color_manual(values = color_fill_values2) +
  scale_y_continuous(labels = scales::label_percent(big.mark = ","),
  breaks = c(0, 5, 10, 15, 20, 30, 40),
  limits = c(0, 40), expand = expansion(mult = c(0, 0.1)))  +
  civilytics::theme_civilytics(font_size = 12) +
  theme(legend.position = "bottom",
  axis.title.x = element_text(margin = margin(t = 15)))  +
  labs(x = "Total Enrollment Decile",
       title = "Referrals",
       y = "",
       color = "Race / Ethnicity")

# +
#   labs(x = "Total Enrollment Quantile",
#        y = "CV (% scale)",
#        color = "Race")


library(patchwork)

pw <- p1 + p2


pw <- pw + plot_annotation(title = "Coefficient of Variation by District Total Enrollment",
      subtitle = "By race/ethnicity\nLoess smooth and observed values shown for each student group",
       caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22",) +
  plot_layout(guides = "collect") &
  civilytics::theme_civilytics(font_size = 11) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = -20),
        panel.grid.major.y = element_line(color = "gray80"))

grid::grid.draw(civilytics::civilytics_logo(pw, position = "bottom-right"))

```

<!-- DRIFT: post5-cov-closing-note -->
Another note: You may also notice that, for arrests, there is an uptick
in the coefficient of variation for the largest districts. In a future
post we will look at one potential source for this increased variance:
underreporting of student arrests.
<!-- /DRIFT -->

:::

## Arrests by race

```{r}
dist_ref_arr |>
    ungroup() |>
    filter(RACE != "TOTAL" & SEX != "TOTAL") |>
    group_by(RACE, SEX) |>
    summarize(
        lea_count = n_distinct(LEAID),
        enrollment = sum(stu_enroll),
        arrests = sum(ARRESTS),
        referrals = sum(REFERRALS)) |>
    mutate(arrest_rate = arrests / (enrollment / 1000),
            referral_rate = referrals / (enrollment/1000)) -> nat_sg_totals

nat_sg_totals |>
    filter(RACE %in% c("AM", "WH", "HI", "BL"))
```



```{r}
#| label: sg_summary_arrestNNH
#| fig-width: 9
#| warning: false
#| message: false
# candidate add-on for Post 6 NNH sentence; not in posted version

pw <- ggplot(nat_sg_totals |> mutate(SEX = ifelse(SEX == "M", "Male students", "Female students")) |>
   filter(RACE %in% c("AM", "WH", "HI", "BL"))) +
  aes(x = crdc_race_recode(RACE), y = 1000 / arrest_rate,
  fill = crdc_race_recode(RACE)) +
    scale_fill_manual(values = color_fill_values2) +
  geom_col() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0))),
              vjust = -0.25) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 12) +
  labs(x = "", y = "NNH", fill = "Race / Ethnicity",
  title = "Enrolled students per arrest by race/ethnicity and sex",
  subtitle = "Number needed to harm (NNH), higher numbers are lower arrest rates\nBy race/ethnicity and sex",
       caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())

grid::grid.draw(civilytics::civilytics_logo(pw, position = "bottom-right"))



```




::: {.smpost slug="arrest-rates-by-race" title="Basic facts — arrest rates by race" date="2025-03-08" status="gold" series="6"}

<!-- DRIFT: post6-basic-facts-race -->
Over 50% of school-based arrests of American Indian/Alaskan Native
students in 2021-22 occurred in just 7 districts.

Disparities in arrest rates by race/ethnicity, gender, and disability
status are widely known and have been documented for years (e.g., for a
small sampling of this work, see [Darling-Hammond and Ho
2024](https://journals.sagepub.com/doi/full/10.1177/23328584241293411);
[Homer and Fisher
2019](https://www.tandfonline.com/doi/full/10.1080/15388220.2019.1604377);
[Whitaker et. al
2019](https://www.aclu.org/publications/cops-and-no-counselors);
[Civilytics' 2017-18 CRDC
analysis](https://www.tableau.com/foundation/data-equity/equitable-education/student-retention-referrals)).

There are many different ways to describe these inequities.

We can talk about *raw numbers*: more Black female students were
arrested than White female students despite there being 3 times as many
White female students enrolled in U.S. public schools.

We can talk about *arrest rates*: White students had 0.51 arrests per
1,000 students, Hispanic students 0.66, American Indian/Alaskan Native
students 1.16, and Black students 1.62 arrests per 1,000 students.

We can highlight the "number needed to harm." While 1 in about 1,500
White males were arrested at school in 2021-22, 1 in about 500 Black
males and 1 in about 700 American Indian/Alaskan Native males were
arrested at school in 2021-22.

And we can talk about the geographic concentration of students by race
-- and therefore the geographic concentration of arrests by race. That
leads us back to where this post started: over 50% of school-based
arrests of American Indian/Alaskan Native students occurred in just 7
districts.
<!-- /DRIFT -->

```{r}
#| label: arrests_concentration_by_sg
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
p1 <- dist_ref_arr %>%
    filter(RACE != "TOTAL" & SEX  == "TOTAL") %>%
    filter(RACE %in% c("BL", "WH", "HI", "AM")) |>
  group_by(RACE) %>%
  arrange(RACE, desc(ARRESTS)) %>%
  mutate(total_arr = sum(ARRESTS),
         total_pop = sum(stu_enroll)) %>%
  filter(ARRESTS > 0) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_pop) %>%
  mutate(dist_per_cum = cumsum(dist_per)) %>%
  ggplot(aes(x = dist_per_cum, y = cum_sum_per, color = crdc_race_recode(RACE),
             group = crdc_race_recode(RACE))) +
  geom_line(linewidth = 1.1) +
  geom_hline(yintercept = 0.5) +
  scale_color_manual(values = color_fill_values2) +
  scale_x_continuous("% of national student population",
                     labels = scales::percent) +
  scale_y_continuous("% of all arrests", labels = scales::percent) +
 civilytics::theme_civilytics(font_size =12) +
  labs(title = "Share of cumulative arrests by student population",
       subtitle = "Horizontal line at 50% of all arrests",
       caption = "Source: U.S. Department of Education Civil Rights Data Collection, 2021-22",
       color = "Race/Eth.") +
  theme(axis.ticks.x = element_blank(),
        panel.grid = element_blank()) +
  theme(legend.position = "inside",
    legend.position.inside = c(0.8, 0.2))

grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-right"))
```


:::

::: {.smpost slug="american-indian-arrests-models" title="A closer look: American Indian student arrests and what models tell us" date="" draft="true" status="draft" series="7"}

<!-- DRIFT: post7-intro -->
Post 6 ended with a striking fact: more than half of all American
Indian/Alaska Native student arrests in 2021-22 were concentrated in
just 7 school districts. Nationwide, 529 American Indian students were
reported arrested -- a small number, but one that is geographically
extreme in its concentration.

The table below shows those 7 districts. Two things jump out immediately.
First, these are small communities: the largest district enrolls just
over 2,000 American Indian students; the smallest enrolls only 58. Second,
their arrest rates are far above the national average of 1.16 arrests
per 1,000 American Indian students -- in some cases by an order of
magnitude.

Extreme values in small populations are worth examining carefully. They
can reflect real patterns, data errors, or one-time events. Before we
accept or dismiss these numbers, we should look closer.
<!-- /DRIFT -->

```{r}
#| label: am_districts_table
#| message: false
#| warning: false

img_path <- "export/figures/socialmedia-am_districts_table.png"

# Your data pipeline with flextable conversion
table_data <- dist_ref_arr %>%
    filter(RACE != "TOTAL" & SEX  == "TOTAL") %>%
    filter(RACE %in% c("AM")) |>
  group_by(RACE) %>%
  arrange(RACE, desc(ARRESTS)) %>%
  mutate(total_arr = sum(ARRESTS),
         total_pop = sum(stu_enroll)) %>%
  filter(ARRESTS > 0) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_pop) %>%
  mutate(dist_per_cum = cumsum(dist_per)) |>
  filter(cum_sum_per <= 0.518) |>
  mutate(arr_rate = ARRESTS / (stu_enroll/1000)) |>
  ungroup () |>
  select(YEAR, LEA_STATE, LEAID, ARRESTS, REFERRALS, stu_enroll, arr_rate, total_enroll) |>
  inner_join(recent_names) |>
  select(YEAR, LEA_STATE, LEA_NAME, everything())

# Store row count before formatting
n_rows <- nrow(table_data)

# Format the data for flextable
table_data <- table_data |>
  select(-LEAID) |>
  mutate(
    LEA_NAME                 = stringr::str_to_title(LEA_NAME),
    YEAR                     = dplyr::recode_values(YEAR,
      "21-22" ~ "2021-22",
      "17-18" ~ "2017-18",
      "15-16" ~ "2015-16"),
    ARRESTS                  = scales::comma(ARRESTS, accuracy = 1),
    REFERRALS                = scales::comma(REFERRALS, accuracy = 1),
    arr_rate                 = round(arr_rate, 2),
    stu_enroll               = scales::comma(stu_enroll, accuracy = 1),
    total_enroll             = scales::comma(total_enroll, accuracy = 1)
  )

header_bg <- "#2c3e50"
header_fg <- "white"
alt_bg    <- "#f0f0eb"

ft <- flextable(table_data) |>
  set_header_labels(
    YEAR                     = "Year",
    LEA_STATE                = "State",
    LEA_NAME                 = "District",
    ARRESTS                  = "Arrests",
    REFERRALS                = "Referrals",
    arr_rate                 = "Arrests per 1,000",
    stu_enroll               = "Enrollment",
    total_enroll             = "Total Enrollment"
  ) |>
  add_header_lines("Top Districts for American Indian/Alaska Native Student Arrests") |>
  add_header_lines("Districts contributing to 51.8% of arrests") |>
  bg(bg = header_bg, part = "header") |>
  color(color = header_fg, part = "header") |>
  bold(part = "header") |>
  fontsize(i = 1, size = 14, part = "header") |>
  fontsize(size = 10, part = "body") |>
  #bg(i = seq(2, n_rows + 1, 2), bg = alt_bg, part = "body") |>
  font(fontname = "Arial", part = "all") |>
  align(j = c("LEA_STATE", "LEA_NAME"), align = "left", part = "body") |>
  align(j = c("ARRESTS", "REFERRALS", "arr_rate","stu_enroll", "total_enroll"), align = "right", part = "body") |>
  align(part = "header", align = "center") |>
  width(j = "YEAR",                width = 0.8) |>
  width(j = "LEA_STATE",           width = 0.6)  |>
  width(j = "LEA_NAME",            width = 2.5)  |>
  width(j = "ARRESTS",             width = 0.9)  |>
  width(j = "REFERRALS",           width = 0.9)  |>
  width(j = "arr_rate",           width = 0.9)  |>
  width(j = "stu_enroll",          width = 1.2)  |>
  width(j = "total_enroll",        width = 1.2)  |>
  border_outer(part = "all",  border = officer::fp_border(color = "#888888", width = 1)) |>
  border_inner_h(part = "body", border = officer::fp_border(color = "#cccccc", width = 0.5)) |>
  add_footer_lines("Source: U.S. Department of Education Civil Rights Data Collection, 2021-22") |>
  add_footer_lines("Nationally, 529 American Indian / Alaska Native student arrests were reported.") |>
  fontsize(size = 9, part = "footer") |>
  color(color = "#555555", part = "footer") |>
  set_table_properties(layout = "fixed")

# Export table as PNG
dims <- flextable_dim(ft)
ragg::agg_png(img_path,
  width = dims$width, height = dims$height + 0.4,
  units = "in", res = 300, scaling = 1
)
plot(ft)
dev.off()

# Composite logo onto saved PNG using magick
logo  <- magick::image_read(logo_file)
table_img <- magick::image_read(img_path)
img_w <- magick::image_info(table_img)$width
img_h <- magick::image_info(table_img)$height
logo_w <- round(img_w * 0.15)
logo   <- magick::image_scale(logo, paste0(logo_w, "x"))
logo_h <- magick::image_info(logo)$height
table_img <- magick::image_composite(
  table_img, logo,
  offset = paste0("+", img_w - logo_w - 20, "+", img_h - logo_h - 20)
)
magick::image_write(table_img, img_path, format = "png")

# Embed in rendered document
knitr::include_graphics(img_path)

```

<!-- DRIFT: post7-three-year-intro -->
A first step is to look at these same districts across multiple waves of
CRDC data. If the 2021-22 numbers reflect a true increase -- not a
reporting anomaly -- we would expect to see some consistency or a clear
trend over time. The table below shows arrest counts for each of these 7
districts across the three available CRDC waves: 2015-16, 2017-18, and
2021-22.
<!-- /DRIFT -->

```{r}
#| label: am_districts_table_3yr

img_path <- "export/figures/socialmedia-am_districts_table_3yr.png"
threewave <- bind_rows(read_stage_df("stages/crdc/full_crdc_data_y2122.parquet"), read_stage_df("stages/crdc/full_crdc_data_y1718.parquet"), read_stage_df("stages/crdc/full_crdc_data_y1516.parquet"))

threewave <- threewave %>%
  group_by(YEAR,LEA_STATE, LEAID, RACE, SEX) %>%
  summarize(ARRESTS = sum(ARRESTS),
            REFERRALS = sum(REFERRALS),
            stu_enroll = sum(stu_enroll),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = ARRESTS / (stu_enroll / 1000),
         referral_rate = REFERRALS / (stu_enroll / 1000)) %>%
  group_by(YEAR, LEA_STATE, LEAID) |>
  mutate(total_enroll = sum(stu_enroll[RACE != "TOTAL" &
                                         SEX != "TOTAL"]))



# Your data pipeline with flextable conversion
table_data <- dist_ref_arr %>%
    filter(RACE != "TOTAL" & SEX  == "TOTAL") %>%
    filter(RACE %in% c("AM")) |>
  group_by(RACE) %>%
  arrange(RACE, desc(ARRESTS)) %>%
  mutate(total_arr = sum(ARRESTS),
         total_pop = sum(stu_enroll)) %>%
  filter(ARRESTS > 0) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_pop) %>%
  mutate(dist_per_cum = cumsum(dist_per)) |>
  filter(cum_sum_per <= 0.518) |>
  mutate(arr_rate = ARRESTS / (stu_enroll/1000)) |>
  ungroup () |>
  select(YEAR, LEA_STATE, LEAID, ARRESTS, REFERRALS, stu_enroll, arr_rate, total_enroll) |>
  inner_join(recent_names) |>
  select(YEAR, LEA_STATE, LEA_NAME, everything())


# Your data pipeline with flextable conversion
table_data3 <- threewave %>%
    filter(RACE != "TOTAL" & SEX  == "TOTAL") %>%
    filter(RACE %in% c("AM")) |>
  group_by(RACE) %>%
  arrange(RACE, desc(ARRESTS)) %>%
  filter(LEAID %in% table_data$LEAID) |>
  mutate(total_arr = sum(ARRESTS),
         total_pop = sum(stu_enroll)) %>%
  #filter(ARRESTS > 0) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_pop) %>%
  mutate(dist_per_cum = cumsum(dist_per)) |>
  mutate(arr_rate = ARRESTS / (stu_enroll/1000)) |>
  ungroup () |>
  select(YEAR, LEA_STATE, LEAID, ARRESTS, REFERRALS, stu_enroll, arr_rate, total_enroll) |>
  inner_join(recent_names) |>
  group_by(LEAID) |> arrange(LEAID, YEAR) |>
  select(YEAR, LEA_STATE, LEA_NAME, everything())

table_data3 <- table_data3 |> ungroup() |>
  select(YEAR, LEA_STATE, LEA_NAME, ARRESTS, stu_enroll) |>
  tidyr::pivot_wider(names_from = YEAR, values_from = c(ARRESTS, stu_enroll)) |>
  mutate(LEA_NAME                 = stringr::str_to_title(LEA_NAME)) |>
  ungroup() |>
  arrange(desc(`ARRESTS_21-22`))
  # |>
  # sleect(LEA_STATE, LEA_NAME, )
  # select(LEA_STATE, LEA_NAME, ends_with("15-16"), ends_with("17-18"), everything())




ft <- flextable(table_data3) |>
  set_header_labels(
    LEA_STATE                = "State",
    LEA_NAME                 = "District",
    `ARRESTS_15-16`                  = "Arrests 2015-16",
    `ARRESTS_17-18`                  = "Arrests 2017-18",
    `ARRESTS_21-22`                  = "Arrests 2021-22",
    `stu_enroll_15-16`               = "Enrollment 2015-16",
    `stu_enroll_17-18`                = "Enrollment 2017-18",
    `stu_enroll_21-22`                 = "Enrollment 2021-22"
  ) |>
  add_header_lines("For districts contributing to 51.8% of arrests in 2021-22") |>
  add_header_lines("Three year data on American Indian/Alaska Native Student Arrests") |>
   bg(bg = header_bg, part = "header") |>
  color(color = header_fg, part = "header") |>
  bold(part = "header") |>
  fontsize(i = 1, size = 14, part = "header") |>
  fontsize(size = 10, part = "body") |>
  #bg(i = seq(2, n_rows + 1, 2), bg = alt_bg, part = "body") |>
  font(fontname = "Arial", part = "all") |>
  align(j = c("LEA_STATE", "LEA_NAME"), align = "left", part = "body") |>
  align(j = c("`ARRESTS_15-16`", "`ARRESTS_17-18`", "`ARRESTS_21-22`","`stu_enroll_15-16`", "`stu_enroll_17-18`", "`stu_enroll_21-22`"), align = "right", part = "body") |>
  align(part = "header", align = "center") |>
  # width(j = "YEAR",                width = 0.8) |>
  width(j = "LEA_STATE",           width = 0.6)  |>
  width(j = "LEA_NAME",            width = 2.5)  |>
  # width(j = "ARRESTS",             width = 0.9)  |>
  # width(j = "REFERRALS",           width = 0.9)  |>
  # width(j = "arr_rate",           width = 0.9)  |>
  # width(j = "stu_enroll",          width = 1.2)  |>
  # width(j = "total_enroll",        width = 1.2)  |>
  border_outer(part = "all",  border = officer::fp_border(color = "#888888", width = 1)) |>
  border_inner_h(part = "body", border = officer::fp_border(color = "#cccccc", width = 0.5)) |>
  add_footer_lines("Source: U.S. Department of Education Civil Rights Data Collection, 2021-22") |>
  add_footer_lines("Nationally, 529 American Indian / Alaska Native student arrests were reported in 2021-22.") |>
  fontsize(size = 9, part = "footer") |>
  color(color = "#555555", part = "footer") |>
  set_table_properties(layout = "fixed")

# Export table as PNG
dims <- flextable_dim(ft)
ragg::agg_png(img_path,
  width = dims$width, height = dims$height + 0.4,
  units = "in", res = 300, scaling = 1
)
plot(ft)
dev.off()

# Composite logo onto saved PNG using magick
logo  <- magick::image_read(logo_file)
table_img <- magick::image_read(img_path)
img_w <- magick::image_info(table_img)$width
img_h <- magick::image_info(table_img)$height
logo_w <- round(img_w * 0.15)
logo   <- magick::image_scale(logo, paste0(logo_w, "x"))
logo_h <- magick::image_info(logo)$height
table_img <- magick::image_composite(
  table_img, logo,
  offset = paste0("+", img_w - logo_w - 20, "+", img_h - logo_h - 20)
)
magick::image_write(table_img, img_path, format = "png")

# Embed in rendered document
knitr::include_graphics(img_path)

```


<!-- DRIFT: post7-three-year-observations -->
In every case, arrest counts in 2021-22 are the highest across the three
waves. In some districts -- like Douglas County, Nevada -- this is the
first wave with any reported AM arrests. In others, like Rapid City,
South Dakota, numbers have grown steadily across waves. This is notable
because nationally, arrest totals fell in 2021-22 compared to prior
years. These 7 districts are moving opposite to the national trend.

Descriptive patterns can only take us so far. What we really want to know
is: how *plausible* are these values? How likely are they to reflect the
true arrest totals versus data quality issues? One direct approach would
be to contact the districts. But with dozens of districts reporting
elevated arrests for small student groups, we also want statistical
tools that help us evaluate plausibility at scale. That's where
statistical models come in.
<!-- /DRIFT -->

<!-- DRIFT: post7-model-intro -->
We'll use Sioux Falls, South Dakota as an illustrative example. Sioux
Falls reported 0 AM student arrests in both 2015-16 and 2017-18, then
reported over 10 arrests in 2021-22. That's a striking jump, and it
raises a natural question: does the data rule out a reporting error?

As a first check, we can apply the binomial distribution -- treating
enrollment as the number of trials and constructing a 95% confidence
interval around the reported arrest count. The bright blue pointrange in
the figure below shows this frequentist interval. It does not overlap 0:
using only the 2021-22 enrollment and arrest data, we can rule out zero
arrests.

But that frequentist interval is wide and uses only a single year. The
panel figure below shows multiple model estimates for Sioux Falls side by
side, ranging from a simple single-year, no-covariate model (top left) to
a richer model incorporating all three waves of CRDC data plus referrals
to law enforcement as a covariate and accounting for state and student
group effects (bottom right). Each panel shows the full probability
distribution of predicted arrests; the colored region is the 95% highest
posterior density interval.

Two patterns emerge clearly. First, adding more years of data (moving
from column 1 to column 2) *depresses* the estimate -- models that
incorporate prior waves where arrests were zero assign more probability
mass to low counts, making 0 more plausible in 2021-22 than the raw data
suggest. Second, adding covariates (moving from row 1 to row 2)
*increases* the estimated plausible range -- covariates like referrals
add evidence that some arrests did occur, pushing estimates upward.

In every panel, model estimates are lower than the reported data, and
the frequentist interval includes values -- arrest counts above 20 --
that no model considers likely. The spikiness in the right panels is not
a display artifact: these models output discrete arrest counts, and when
only a few values have meaningful probability, the smoothed density shows
sharp peaks.

Which set of estimates is most trustworthy? Why do models diverge, and
what should we do when they do? We'll unpack that in future posts.[^post7-1]
<!-- /DRIFT -->

```{r}
#| label: arrests_model_comparison_am
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 8
focal_dist <- sample(unique(table_data$LEAID), 1)
focal_dist <- "4659820" # rapid city
focal_dist <- "4666270" #sioux falls
obsv_focal <- rdata |> filter(LEAID == focal_dist) |> filter(RACE == "AM")
# Make plot title
unique(obsv_focal$LEA_NAME)
unique(obsv_focal$LEA_STATE)
dist_name_for_title <- paste0(tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))),
                            " in ", state.name[state.abb==unique(obsv_focal$LEA_STATE)])

dist_test <- get_prediction_summary(con, confidence_level = 0.95, RACE = "AM",
      central_tendency = "median", YEAR = "21-22", LEAID = focal_dist)

dist_draws <- get_prediction_draws(con,  YEAR = "21-22", RACE = "AM", LEAID = focal_dist)


obsv_plot <- obsv_focal |> ungroup() |> group_by(LEAID, YEAR) |>
    summarize(arrests = sum(ARRESTS),
             enroll = sum(stu_enroll)) |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Frequentist")

plot_draws <- dist_draws |>
  filter(!(grepl("_m5", model_id))) |>
  filter(!(grepl("sg_", model_id))) |>
  group_by(model_id, draw_id) |>
              summarize(pred = sum(pred)) |>
  group_by(model_id) |>
  arrange(model_id, desc(pred)) |>
  mutate( ordvar = 1:n()) |>
  mutate(fillvar = abs(ordvar - 250))



# Starting with your filtered and summarized data
result <- dist_draws |>
  filter(!(grepl("_m5", model_id))) |>
  filter(!(grepl("sg_", model_id))) |>
  group_by(model_id, draw_id) |>
  summarize(pred = sum(pred), .groups = "drop") |>
  group_by(model_id) |>
  # Add rank (descending by pred value)
  mutate(
    rank = rank(-pred, ties.method = "min")
  )

# Calculate HPD intervals for each model_id using base R
# First, get unique model_ids
model_ids <- unique(result$model_id)

# Initialize vectors to store HPD bounds
lower_hpd_vec <- numeric(nrow(result))
upper_hpd_vec <- numeric(nrow(result))
within_hpd_vec <- logical(nrow(result))

# Loop through each model_id to calculate HPD
for (i in seq_along(model_ids)) {
  current_model <- model_ids[i]
  model_rows <- which(result$model_id == current_model)
  model_preds <- result$pred[model_rows]

  # Sort predictions for this model
  sorted_preds <- sort(model_preds)

  # Find the narrowest window containing 95% of values
  n_values <- length(sorted_preds)
  window_size <- ceiling(0.95 * n_values)

  # Calculate all possible windows
  window_widths <- numeric(n_values - window_size + 1)
  window_starts <- numeric(n_values - window_size + 1)
  window_ends <- numeric(n_values - window_size + 1)

  for (j in 1:(n_values - window_size + 1)) {
    window_starts[j] <- sorted_preds[j]
    window_ends[j] <- sorted_preds[j + window_size - 1]
    window_widths[j] <- window_ends[j] - window_starts[j]
  }

  # Find the narrowest window
  min_width_idx <- which.min(window_widths)
  lower_hpd <- window_starts[min_width_idx]
  upper_hpd <- window_ends[min_width_idx]

  # Assign HPD bounds and within_hpd flag to all rows for this model
  lower_hpd_vec[model_rows] <- lower_hpd
  upper_hpd_vec[model_rows] <- upper_hpd
  within_hpd_vec[model_rows] <- model_preds >= lower_hpd & model_preds <= upper_hpd
}

# Add HPD information to the result data frame
result$lower_hpd <- lower_hpd_vec
result$upper_hpd <- upper_hpd_vec
result$within_hpd <- within_hpd_vec

# Arrange by model_id and rank
result <- result |>
  arrange(model_id, rank)

# To filter only rows within HPD
hpd_only <- result |> filter(within_hpd)

result <- result |>
  mutate(model_time = add_model_time(model_id),
    model_label = paste0(add_model_time(model_id), "\n", add_model_cov(model_id)))




p1 <- ggplot(result, aes(x = pred, y = model_label, group = model_label)) +
  geom_density_ridges(scale = 0.9, rel_min_height = 0.01) +
    geom_density_ridges_gradient(data = result |> filter(within_hpd),
    aes(fill = after_stat(x)), scale = 0.9, gradient_lwd = 1.0, rel_min_height = 0.01) +
    geom_pointrange(data = obsv_plot %>%
                   tidyr::crossing(model_id = unique(result$model_id)) |>
                   mutate(model_label = paste0(add_model_time(model_id), "\n", add_model_cov(model_id)),
                            model_time = add_model_time(model_id)),
  aes(y = model_label, x = arrests, xmin = ci_lower, fill = NULL, xmax = ci_upper), color = I("#2993fd"),
  position = position_nudge(y = 0.2)) +
     scale_fill_viridis_c(name = "Arrests", option = "C", guide = guide_none()) +
        scale_y_discrete("Model", limits = rev,
                #labels = function(x) print_model_name(x, lbreak = TRUE),
                expand = expansion(mult = c(0.1, 0.15))) +
    facet_wrap(~add_model_time(model_id), ncol = 2, scales = "free") +
    labs(title = stringr::str_wrap(paste0("Predicted arrests probability density interval for ", dist_name_for_title, " for American Indian students"), 90),
    x = "Predicted arrests", y = "Model type",
    color = "Model or observed",
    subtitle = stringr::str_wrap("Observed arrests and frequentist interval shown in bright blue pointrange. Colored fill represents the 95% highest posterior density for each model, shaded to emphasize number of arrests.", 90)) +
    theme_ridges(grid = FALSE, font_size = 14) +
    theme(legend.position = "bottom",
    panel.grid.major.y = element_line(color = "gray50"),
    axis.text.y = element_text(angle = 90, hjust = -0.5))

grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-left"))

```

[^post7-1]: The question of which model is most appropriate involves
choices about how much to trust prior years, which covariates to
include, and how to handle the particular data-generating process for
school arrests. Future posts will make those choices explicit.

:::

## Post 8 Expected vs Observed Arrests

::: {.smpost slug="misreported-zeros-descriptive" title="Failure to report, or a real difference in policy? Populous districts reporting zero arrests" date="" draft="true" status="draft" series="8"}

<!-- DRIFT: post8-intro -->
In 2021-22, 2,060 school districts -- just 11.6% of all districts in
the data -- reported at least one school-based arrest. The most common
value, by far, is zero. Yet that small share of arresting districts
enrolled 45% of all students, meaning nearly half of all students
attended a school where an arrest occurred.

When 88% of districts report zero arrests, a key question arises: are
these true zeros, or failures to report?

The table below shows why the question matters. The national average
arrest rate in 2021-22 was roughly 1 arrest per 1,428 students. For a
district with 20,000 or more students, that implies around 14 expected
arrests. Yet 125 out of 376 large districts (33.2%) reported exactly
zero arrests -- a share that would be nearly impossible if the national
rate applied uniformly to everyone. If arrests followed a binomial
process at the national rate, we would expect essentially every district
of this size to report at least one arrest.
<!-- /DRIFT -->

```{r}
#| label: compute-expected-arrests
thresh <- c(0, 1000, 10000, 20000, Inf)
labs <- c("0-999", "1,000-9,999", "10,000-19,999","20,000+")

dist_size_summary <- dist_ref_arr %>%
   ungroup() %>%
     filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
   mutate(popcut = cut(stu_enroll,
                       breaks = thresh,
                       labels = labs,
                       include.lowest= TRUE)) %>%
   group_by(popcut) %>%
 mutate(expec_arrest = 1 - pbinom(0, stu_enroll, nat_arrest_rate/1000,
                              lower.tail = TRUE)) %>% select(-jj_count, - RACE, - SEX) |>
   summarize(dists = n(),
             dists_w_arrests = sum(ARRESTS > 0),
             expec_distw_arrests = as.integer(round(sum(expec_arrest),0))) %>%
   mutate(dist_w_arrest_per = pretty_per(dists_w_arrests/dists),
          expect_w_arrest_per = pretty_per(expec_distw_arrests/dists))
```

```{r}
#| label: display-expected-arrests

#dist_size_summary |> knitr::kable()

img_path <- "export/figures/socialmedia-expected_arrest_table.png"

ft <- flextable(dist_size_summary) |>
  set_header_labels(
    popcut                = "Enrollment Range",
    dists                 = "Districts",
    dists_w_arrests                     = "Districts reporting an arrest",
    expec_distw_arrests                  = "Districts expected with an arrest",
    dist_w_arrest_per                    = "% of districts with an arrest",
    expect_w_arrest_per =     "% districts expected with an arrest"
  ) |>
  add_header_lines("Expected Arrests based on National Arrest Rate") |>
  #add_header_lines("Three year data on American Indian/Alaska Native Student Arrests") |>
   bg(bg = header_bg, part = "header") |>
  color(color = header_fg, part = "header") |>
  bold(part = "header") |>
  fontsize(i = 1, size = 14, part = "header") |>
  fontsize(size = 10, part = "body") |>
  #bg(i = seq(2, n_rows + 1, 2), bg = alt_bg, part = "body") |>
  font(fontname = "Arial", part = "all") |>
   align(j = c("popcut"), align = "left", part = "body") |>
  align(j = c("dists","dists_w_arrests", "expec_distw_arrests", "dist_w_arrest_per", "expect_w_arrest_per"), align = "right", part = "body") |>
  align(part = "header", align = "center") |>
  width(j = "popcut",                width = 1.7) |>
  width(j = "dists",           width = 1.1)  |>
  width(j = "dists_w_arrests",            width = 1.1)  |>
  width(j = "dist_w_arrest_per",            width = 1.1)  |>
  width(j = "expect_w_arrest_per",            width = 1.1)  |>
  # width(j = "ARRESTS",             width = 0.9)  |>
  # width(j = "REFERRALS",           width = 0.9)  |>
  # width(j = "arr_rate",           width = 0.9)  |>
  # width(j = "stu_enroll",          width = 1.2)  |>
  # width(j = "total_enroll",        width = 1.2)  |>
  border_outer(part = "all",  border = officer::fp_border(color = "#888888", width = 1)) |>
  border_inner_h(part = "body", border = officer::fp_border(color = "#cccccc", width = 0.5)) |>
  add_footer_lines("Source: U.S. Department of Education Civil Rights Data Collection") |>
  add_footer_lines("National arrest rate in 2021-22 was 0.71 arrests per 1,000 students.") |>
  fontsize(size = 9, part = "footer") |>
  color(color = "#555555", part = "footer") |>
  set_table_properties(layout = "fixed")

# Export table as PNG
dims <- flextable_dim(ft)
ragg::agg_png(img_path,
  width = dims$width, height = dims$height + 0.4,
  units = "in", res = 300, scaling = 1
)
plot(ft)
dev.off()

# Composite logo onto saved PNG using magick
logo  <- magick::image_read(logo_file)
table_img <- magick::image_read(img_path)
img_w <- magick::image_info(table_img)$width
img_h <- magick::image_info(table_img)$height
logo_w <- round(img_w * 0.15)
logo   <- magick::image_scale(logo, paste0(logo_w, "x"))
logo_h <- magick::image_info(logo)$height
table_img <- magick::image_composite(
  table_img, logo,
  offset = paste0("+", img_w - logo_w - 20, "+", img_h - logo_h - 20)
)
magick::image_write(table_img, img_path, format = "png")

# Embed in rendered document
knitr::include_graphics(img_path)

```

<!-- DRIFT: post8-suspicious-zeros -->
This level of zeros is suspicious.

To illustrate what suspicious reporting looks like, consider Paterson
Public School District in New Jersey, with roughly 18,000 students
enrolled in grades 7-12. The table below shows Paterson's arrest
reporting across all three CRDC waves available in these data: 47
arrests in 2015-16, 11 in 2017-18, and 0 in 2021-22. A district of
this size, with a documented history of arrests, reporting zero in the
most recent wave is difficult to explain by policy change alone.

Despite steps taken before, during, and after data collection to ensure
good data quality ([Office of Civil Rights
2025](https://www.ed.gov/media/document/crdc-quality-informational-snapshot-january-2025-109165.pdf)),
data errors exist. The number of reported -- but possibly erroneous --
zeros poses challenges not just for understanding arrest patterns in
particular districts, but for our whole understanding of national
arrest rates.[^post8-1]
<!-- /DRIFT -->

```{r}
#| label: big0_example_dist_table

focal_dist <- "3412690"  # Paterson NJ

img_path <- "export/figures/socialmedia-big0_example_dist_table.png"

table_data3 <- tydata |> filter(LEAID == focal_dist) |>
  group_by(LEA_STATE, LEA_NAME, YEAR) |>
  summarize(arrests = sum(ARRESTS), enroll = sum(stu_enroll))

ft <- flextable(table_data3) |>
  set_header_labels(
    LEA_STATE                = "State",
    LEA_NAME                 = "District",
    YEAR                     = "CRDC Wave",
    arrests                  = "Arrests",
    enroll                    = "Enrollment"
  ) |>
  add_header_lines("Suspiciously Low Reported Arrest Counts") |>
  #add_header_lines("Three year data on American Indian/Alaska Native Student Arrests") |>
   bg(bg = header_bg, part = "header") |>
  color(color = header_fg, part = "header") |>
  bold(part = "header") |>
  fontsize(i = 1, size = 14, part = "header") |>
  fontsize(size = 10, part = "body") |>
  #bg(i = seq(2, n_rows + 1, 2), bg = alt_bg, part = "body") |>
  font(fontname = "Arial", part = "all") |>
  align(j = c("YEAR", "LEA_STATE", "LEA_NAME"), align = "left", part = "body") |>
  align(j = c("arrests", "enroll"), align = "right", part = "body") |>
  align(part = "header", align = "center") |>
  width(j = "YEAR",                width = 0.8) |>
  width(j = "LEA_STATE",           width = 0.6)  |>
  width(j = "LEA_NAME",            width = 2.5)  |>
  # width(j = "ARRESTS",             width = 0.9)  |>
  # width(j = "REFERRALS",           width = 0.9)  |>
  # width(j = "arr_rate",           width = 0.9)  |>
  # width(j = "stu_enroll",          width = 1.2)  |>
  # width(j = "total_enroll",        width = 1.2)  |>
  border_outer(part = "all",  border = officer::fp_border(color = "#888888", width = 1)) |>
  border_inner_h(part = "body", border = officer::fp_border(color = "#cccccc", width = 0.5)) |>
  add_footer_lines("Source: U.S. Department of Education Civil Rights Data Collection") |>
  add_footer_lines("Enrollment shown is for grades 7-12.") |>
  fontsize(size = 9, part = "footer") |>
  color(color = "#555555", part = "footer") |>
  set_table_properties(layout = "fixed")

# Export table as PNG
dims <- flextable_dim(ft)
ragg::agg_png(img_path,
  width = dims$width, height = dims$height + 0.4,
  units = "in", res = 300, scaling = 1
)
plot(ft)
dev.off()

# Composite logo onto saved PNG using magick
logo  <- magick::image_read(logo_file)
table_img <- magick::image_read(img_path)
img_w <- magick::image_info(table_img)$width
img_h <- magick::image_info(table_img)$height
logo_w <- round(img_w * 0.15)
logo   <- magick::image_scale(logo, paste0(logo_w, "x"))
logo_h <- magick::image_info(logo)$height
table_img <- magick::image_composite(
  table_img, logo,
  offset = paste0("+", img_w - logo_w - 20, "+", img_h - logo_h - 20)
)
magick::image_write(table_img, img_path, format = "png")

# Embed in rendered document
knitr::include_graphics(img_path)
```

[^post8-1]: Potentially erroneous zeros skew the distribution of arrest
rates and reduce the value of summary statistics like the mean, since
few districts actually sit near that average.

:::

```{r}

# Get 100 highest arrests
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      arrange(desc(total_arrests)) |>
      ungroup() |>
      slice_head(n= 100) -> bigarr

# Get 100 largest districts
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      arrange(desc(total_enroll)) |>
      ungroup() |>
      slice_head(n= 100) -> big100

# Get largest districts with 0 arrests
rdata |> group_by(LEA_STATE, LEAID, LEA_NAME) |>
      summarize(total_enroll = sum(stu_enroll),
            total_arrests = sum(ARRESTS)) |>
      filter(total_arrests == 0) |>
      arrange(desc(total_enroll)) |>
      ungroup() |>
      slice_head(n= 100) -> big0
```


```{r}
# focal_dist <- "4823640"
# focal_dist <- "4826580"
# focal_dist <- "2502790"
focal_dist <- "3412690" # patterson

obsv_focal <- rdata |> filter(LEAID == focal_dist)
# Make plot title
unique(obsv_focal$LEA_NAME)
unique(obsv_focal$LEA_STATE)
dist_name_for_title <- paste0(tools::toTitleCase(tolower(unique(obsv_focal$LEA_NAME))),
                            " in ", state.name[state.abb==unique(obsv_focal$LEA_STATE)])


dist_test <- get_prediction_summary(con, confidence_level = 0.95,
      central_tendency = "median", YEAR = "21-22", LEAID = focal_dist)

dist_draws <- get_prediction_draws(con,  YEAR = "21-22", LEAID = focal_dist)


obsv_plot <- obsv_focal |> ungroup() |>
    summarize(arrests = sum(ARRESTS),
             enroll = sum(stu_enroll)) |>
    mutate(ci_upper = agresti_coull(arrests, enroll, confidence_level = 0.95)[1],
            ci_lower = agresti_coull(arrests, enroll, confidence_level = 0.95)[2]) |>
    mutate(type = "Frequentist")

model_plot <- dist_draws |>
    group_by(model_id, draw_id) |>
    summarize(arrestst = sum(pred)) |>
    group_by(model_id) |>
    summarize(arrests = median(arrestst),
            ci_upper = quantile(arrestst, 1 - (1 - 0.95)/2),
            ci_lower = quantile(arrestst, (1 - 0.95)/2),
            enroll = obsv_plot$enroll[1]) |>
    rename(type = model_id)

plotdf <- bind_rows(obsv_plot %>%
  tidyr::crossing(model_id = unique(model_plot$type)),
  model_plot |> mutate(model_id = type)) |>
    mutate(type = ifelse(type == "Frequentist", type, "Modeled"))





```

```{r}
#| label: paterson_model_comparison
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 8

# Figure 1 of Post 9 (Posts 7-x): predicted-arrest density per model vs the
# frequentist interval for Paterson NJ (focal_dist = 3412690), mirroring
# arrests_model_comparison_am. dist_draws/obsv_plot are from the prep chunk above.
res_pat <- dist_draws |>
  filter(!grepl("_m5", model_id), !grepl("sg_", model_id)) |>
  group_by(model_id, draw_id) |>
  summarize(pred = sum(pred), .groups = "drop") |>
  group_by(model_id) |>
  mutate(model_time = add_model_time(model_id),
         model_label = paste0(add_model_time(model_id), "\n", add_model_cov(model_id)))

obsv_pat <- obsv_plot |>
  tidyr::crossing(model_id = unique(res_pat$model_id)) |>
  mutate(model_label = paste0(add_model_time(model_id), "\n", add_model_cov(model_id)),
         model_time  = add_model_time(model_id))

p1 <- ggplot(res_pat, aes(x = pred, y = model_label, group = model_label)) +
  ggridges::geom_density_ridges(scale = 0.9, rel_min_height = 0.01) +
  geom_pointrange(data = obsv_pat,
    aes(y = model_label, x = arrests, xmin = ci_lower, xmax = ci_upper),
    color = I("#2993fd"), position = position_nudge(y = 0.2)) +
  scale_y_discrete("Model", limits = rev,
                   expand = expansion(mult = c(0.1, 0.15))) +
  facet_wrap(~add_model_time(model_id), ncol = 2, scales = "free") +
  labs(title = stringr::str_wrap(paste0("Predicted arrests probability density for ",
        dist_name_for_title), 90),
       x = "Predicted arrests", y = "Model type",
       subtitle = stringr::str_wrap("Observed arrests and frequentist interval shown in bright blue pointrange (0 reported arrests).", 90)) +
  ggridges::theme_ridges(grid = FALSE, font_size = 14) +
  theme(legend.position = "bottom",
        panel.grid.major.y = element_line(color = "gray50"),
        axis.text.y = element_text(angle = 90, hjust = -0.5))

grid::grid.draw(civilytics::civilytics_logo(p1, position = "bottom-left"))
```

```{r}
#| label: zero_arrest_counts_draws
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

plot_draws <- dist_draws |> group_by(model_id, draw_id) |>
              summarize(pred = sum(pred))
ndraws <- 500 # draws per model per group, 500 * 8

plot_draws <- plot_draws |>
              mutate(pred = ifelse(pred > 16, 16, pred))



p1 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
             # filter(grepl("nat_", model_id)) |>
            filter(add_model_cov(model_id) == "Baseline") |>
            filter(add_model_time(model_id) == "Three year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 3, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.1))) +
    scale_x_continuous(breaks = c(5:16), limits = c(4, 17),
    expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "Three year models", subtitle = "Baseline", x ="", y = "") +
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))


p2 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
  #filter(grepl("nat_", model_id)) |>
  filter(add_model_cov(model_id) == "Baseline") |>
     filter(add_model_time(model_id) == "One year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 3, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.2)), name = "Model type") +
    scale_x_continuous(breaks = c(0:8), limits = c(-0.5, 8),
    expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "One year models", subtitle = "Baseline", x = "", y = "Model type")+
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))

p3 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
 # filter(grepl("nat_", model_id)) |>
            filter(add_model_cov(model_id) == "Covariate") |>
            filter(add_model_time(model_id) == "Three year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 3, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.1))) +
    scale_x_continuous(breaks = c(0:12), limits = c(-0.5, 13),
   expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "Three year models", subtitle = "Covariates",
            x = "Predicted Arrests", y ="") +
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))


p4 <- ggplot(plot_draws |> filter(!(grepl("_m5", model_id))) |>
#  filter(grepl("nat_", model_id)) |>
      filter(add_model_cov(model_id) == "Covariate") |>
     filter(add_model_time(model_id) == "One year"),
    aes(x = pred, y = model_id, group = model_id)) +
    geom_density_ridges2(stat = "binline", scale= 0.9, binwidth = 1,
                          color = "#000a9bff") +
      geom_text(
    stat = "bin",
    aes(
      y = group + (0.9 * after_stat(count/max(count))),
      label = ifelse(stat(count) > 25, pretty_per(after_stat(count/ndraws)), "")
      #label = ifelse(stat(count) > 0, pretty_per(stat(count/max(count)), ndigit = 3), "")
    ),
    nudge_y = 0.0,
    vjust = -0.25, size = 3, color = "black", binwidth = 1
  ) +
    scale_y_discrete(labels = function(x) print_model_name(x, lbreak = TRUE),
    expand = expansion(add = c(0.25, 1.2))) +
    scale_x_continuous(breaks = c(0:8), limits = c(-0.5, 8),
    expand = expansion(add = c(0.25, 1), mult = c(0, 0.2))) +
    theme_ridges(grid = FALSE, font_size = 14) +
    labs(title = "One year models", subtitle = "Covariates",
     y = "", x = "Predicted Arrests")+
    theme(axis.text.y = element_text(angle = 90, hjust = -0.5))

pfull <- p2+ p1 + p4 + p3  + plot_annotation(
  title = paste0("All model draws for ", dist_name_for_title),
  caption = "Observed arrests in 2021-22 are 0.\nArrests are top-coded at 16 or more for visual clarity.\nValues with fewer than 5% predicted likelihood are not labeled.",
  subtitle = 'Frequency of predicted arrests from 500 draws of posterior for each model'
)

grid::grid.draw(civilytics::civilytics_logo(pfull, position = "bottom-left"))

```

In our first example, we look at how Bayesian prediction models differ from
frequentist intervals in the case of no observed arrests. The Paterson New Jersey
school district has `r pretty_count(obsv_focal$total_enroll[1])`
in our dataset - so having 0 arrests is suspicious.

In the case of no arrests, the observed interval is constant - from 0 to 3, and shown in gray
alongside the Bayesian prediction interval in blue for
each of our 10 models. The models are sorted into single-year specifications
on the left and three year specifications on the right. We see that the single-year
models, except for stratfiied model 1, have much narrower intervals than the
frequentist interval - confirming the advantage of the modeling methodology for increasing
precision.

However, these models still suggest that 0 arrests is the median estimate (the point).
Here the three year models disagree - the intervals are all wider
than the frequentist interval and the median estimated arrests is greater than 0.
In particular models without covariates (Unifed Model 3
and Stratified Model 3) suggest that more than 10 arrests are most liklely. The
reason becomes clear when we look at the multiple years of data and see that
in 15-16 Paterson had 47 arrests reported and in 17-18 it had 11 arrests reported.


```{r}
#| label: cleanup
#| include: false

close_draws_view(h)
```
