---
title: "Supplementary Materials: Equity Analysis at a Large Scale (CRDC School Arrests)"
author:
  - name: "Jared E. Knowles"
    affiliation: "Civilytics Consulting"
  - name: "Hannah Miller"
    affiliation: "Civilytics Consulting"
format:
  html:
    theme: theme/civilytics.scss
    css: theme/extras.css
    toc: true
    toc-location: right
    toc-depth: 3
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
  error: false
knitr:
  opts_chunk:
    fig.path: export/figures/supplement-
---

This document collects the supplementary materials for the white paper *Equity
Analysis at a Large Scale: Using Small Area Estimation to Get the Most from the
CRDC School Arrest Data*. It has four parts: exploratory analysis of the combined
three-year CRDC data, the data-construction and sample-restriction details,
model performance and diagnostics, and additional applied prediction-interval
examples. The headline tables and figures, and the national descriptive totals
by race and sex, are reported in the paper itself (`white_paper.html`); this
companion holds the supporting detail. Like the paper, every section reads the
published artifacts via `crdc_path()` and renders standalone.

```{r}
#| label: setup
#| include: false

# Libraries (union of the three merged source documents)
library(targets)
library(brms)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggdist)
library(marginaleffects)
library(tidybayes)
library(ggridges)
library(patchwork)
library(duckdb)
library(knitr)

options(scipen = 9)
options(width = 120)
if (!dir.exists("export/figures")) dir.create("export/figures", recursive = TRUE)

# --- Artifact-access layer (documented reproduction interface; kept sourced) ---
source("R/funs.R")
source("R/crdc_path.R"); source("R/model_registry.R"); source("R/paper_figures.R")
cv_apply_branding()                    # civilytics theme on all figures
h <- open_draws_view(); con <- h$con   # predicted_draws view over published parquet

# --- Inputs (loaded once for all sections) ---
tydata <- read_stage_df("stages/inputs/three_year_data.parquet")
rdata  <- read_stage_df("stages/inputs/recent_data.parquet")
y2122  <- read_stage_df("stages/crdc/full_crdc_data_y2122.parquet")

# National 2021-22 arrest rate used by the District Concentration analysis (Part 2)
nat_totals <- y2122 |>
  ungroup() |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarize(
    all_lea_count = n_distinct(LEAID),
    lea_any_arr   = n_distinct(LEAID[ARRESTS > 0]),
    enrollment    = sum(stu_enroll),
    arrests       = sum(ARRESTS),
    referrals     = sum(REFERRALS))
nat_arrest_rate <- nat_totals$arrests / (nat_totals$enrollment / 1000)

# Combined three-year frame + canonical (most-recent-wave) LEA name (Part 1 EDA)
combined_data <- read_stage_df("stages/inputs/combined_model_data.parquet")
combined_data |>
  select(LEAID, LEA_NAME, YEAR) |>
  distinct_all() |>
  group_by(LEAID) |>
  mutate(last_name = ifelse(any(YEAR == "21-22"), LEA_NAME[YEAR == "21-22"],
                     ifelse(any(YEAR == "17-18"), LEA_NAME[YEAR == "17-18"],
                            LEA_NAME[YEAR == "15-16"]))) |>
  ungroup() |>
  select(LEAID, last_name) |>
  distinct_all() -> lea_names_canonical
combined_data <- left_join(combined_data |> select(-LEA_NAME),
                           lea_names_canonical |> rename(LEA_NAME = last_name),
                           by = join_by(LEAID))
rm(lea_names_canonical)

# State shapes for the EDA choropleths (Part 1)
state_map <- tigris::states(cb = TRUE, resolution = "20m") |> tigris::shift_geometry()

# --- Colour palettes (inlined for transparency) ---
color_fill_values  <- c("WH" = "#c92d0e", "BL" = "#020684", "AM" = "#ff8f43",
                        "TOTAL" = "#909090ff", "HI" = "#0791b6")
color_fill_values2 <- c("White" = "#c92d0e", "Black" = "#020684",
                        "Total" = "#909090ff", "Hispanic" = "#0791b6",
                        "Amer. Ind." = "#ff8f43")
model_palette <- c("Modeled" = "#000a9bff", "Frequentist" = "#858585bb")
group_palette <- c("WH" = "#09d2ecc2", "BL" = "#ab00a9f1", "HI" = "#00ab17f1")

# --- Model-labelling helpers (inlined for transparency; used by Parts 3-4) ---
add_model_time <- function(x) {
  y <- rep("One year", length(x))
  y[x %in% c("nat_m3_mod", "nat_m4_mod", "nat_m5_mod",
             "sg_m3_mod", "sg_m4_mod", "sg_m5_mod")] <- "Three year"
  y
}
add_model_form <- function(x) {
  y <- rep("Stratified", length(x)); y[startsWith(x, "nat")] <- "Unified"; y
}
add_model_cov <- function(x) {
  y <- rep("Baseline", length(x))
  y[x %in% c("sg_m2_mod", "nat_m4_mod", "nat_m5_mod",
             "nat_m2_mod", "sg_m4_mod", "sg_m5_mod")] <- "Covariate"
  y
}
print_model_name <- function(x, lbreak = FALSE) {
  y <- rep("Stratified", length(x)); y[startsWith(x, "nat")] <- "Unified"
  x <- gsub("_mod", "", x); x <- gsub("nat_m", "", x); x <- gsub("sg_m", "", x)
  if (lbreak) paste0(y, "\nModel:", x) else paste0(y, " Model:", x)
}

# --- Agresti-Coull frequentist interval (inlined; used by Part 4 examples) ---
agresti_coull <- function(numerator, denominator, confidence_level = 0.95) {
  adj_star <- qnorm(1 - (1 - confidence_level) / 2)
  if (numerator > 0) {
    num_star   <- numerator + adj_star
    denom_star <- denominator + (2 * adj_star)
    phat       <- num_star / denom_star
    phat_se    <- sqrt((phat / denom_star) * (1 - phat))
    ci_upper   <- (phat + (adj_star * phat_se)) * denom_star
    ci_lower   <- (phat - (adj_star * phat_se)) * denom_star
    sd         <- phat_se * denom_star
  } else {
    ci_upper <- denominator * (-log(1 - confidence_level) / denominator)
    ci_lower <- 0
    sd       <- mean(c(ci_upper, ci_lower)) / confidence_level
    phat_se  <- 0
    phat     <- mean(c(ci_upper, ci_lower))
  }
  c(ci_upper, ci_lower, sd, phat_se, phat)
}
```

# Part 1 — Exploratory Data Analysis

This part explores the combined three-year CRDC data (2015-16, 2017-18,
2021-22). Counts and rates here are pooled across all three waves unless stated
otherwise, and so differ from the single-wave 2021-22 figures in the paper.

## Basic Data Structure

```{r}
#| label: data-structure
# Basic structure and dimensions
cat("Data dimensions:", nrow(combined_data), "rows,", ncol(combined_data), "columns\n\n")

# Column names and types
str(combined_data)

# First few rows
head(combined_data) |> kable()
```

## Summary Statistics by Year

```{r}
#| label: year-summary
# Summary statistics by year
year_summary <- combined_data |>
  group_by(YEAR) |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    mean_enrollment = mean(stu_enroll, na.rm = TRUE),
    mean_arrests = mean(ARRESTS, na.rm = TRUE),
    mean_referrals = mean(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )

year_summary |> knitr::kable(digits = 2)



```

### Summary Plot of Trends

```{r}
#| label: national-arrest-trend-plots

year_summary <- combined_data |>
  group_by(YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(year_summary, aes(x = YEAR, y = total_arrests)) +
  geom_point() +
  geom_line(group = 1) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  labs(x = "School Year", y = "CRDC Arrests", title = "Arrest Trends Across CRDC Collections") +
  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE)

ggplot(year_summary, aes(x = YEAR, y = total_arrests/(total_enrollment/1000))) +
  geom_point() +
  geom_line(group = 1) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  labs(x = "School Year", y = "CRDC Arrests per 1,000", title = "Arrest Rate Trends Across CRDC Collections") +
  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE)




```

## State Variation Summing over 3 years

```{r}
#| label: state_map_arrest_rate
#| echo: false
#| warning: false
#| message: false

state_summary <- combined_data |>
  group_by(LEA_STATE) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate =  total_arrests / (total_enrollment / 1000)) |>
    mutate(arrest_nnh = total_enrollment / total_arrests)



state_map2 <- inner_join(state_map, state_summary, by = join_by(STUSPS == LEA_STATE))

ggplot(data = state_map2, aes( fill = arrest_rate)) +
  geom_sf() +
  geom_sf_label(aes(label = round(arrest_rate, 1))) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  labs(title = "Arrests per 1,000 for all students by state") +
  theme_void()

#rm(state_map2)

```



```{r}
#| label: state_map_arrestsnnh
#| echo: false
#| warning: false
#| message: false
ggplot(data = state_map2, aes( fill = arrest_nnh)) +
  geom_sf() +
  geom_sf_label(aes(label = civilytics::pretty_count(round(arrest_nnh, 0)))) +
  scale_fill_distiller(palette = "YlOrRd", direction = -1,
  breaks = c(50000, 8000, 3000, 1000, 500),
  trans = "log", labels = scales::comma) +
  labs(title = "Students per arrested student by state", fill = "NNH") +
  theme_void() +
  theme(legend.position.inside = c(0.1, 0.8), legend.position = "inside")
```

## Race and Sex Distribution

These counts and rates are **pooled across all three CRDC waves (2015-16,
2017-18, 2021-22)** and are shown here as number-needed-to-harm (NNH, students
per arrested student). They therefore differ from the single-wave 2021-22
totals and per-1,000 rates reported in `white_paper.qmd`; they are not the same
figures and should not be read as such.

```{r}
#| label: race-sex-distribution
# Distribution across race and sex categories
race_sex_summary <- combined_data |>
  filter(RACE != "TOTAL", SEX != "TOTAL") |>
  group_by(RACE, SEX) |>
  summarise(
    n_observations = n(),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate = total_arrests / (total_enrollment / 1000)) |>
  arrange(desc(total_arrests))

race_sex_summary |> knitr::kable(digits = 3)
```



```{r}
#| label: nnh_arrest_counts_3yr

ggplot(race_sex_summary |> filter(RACE %in% c("AM", "WH", "HI", "BL"))) +
  aes(x = crdc_race_recode(RACE), y = 1000 / arrest_rate,
  fill = crdc_race_recode(RACE)) +
    scale_fill_manual(values = color_fill_values2) +
  geom_col() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0))),
              vjust = -0.25) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE) +
  labs(x = "", y = "NNH (Arrests)", fill = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())

```

### Trends by group

```{r}
#| label: race-sex-arrestratetrends
race_sex_trends <- combined_data |>
  filter(RACE != "TOTAL", SEX != "TOTAL") |>
  group_by(YEAR, RACE, SEX) |>
  summarise(
    n_observations = n(),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate = total_arrests / (total_enrollment / 1000)) |>
  arrange(desc(total_arrests))



ggplot(race_sex_trends |> filter(RACE %in% c("AM", "WH", "HI", "BL")) |>
  mutate(group_var = paste(RACE, SEX, sep = "_"))) +
  aes(x = YEAR, y = 1000 / arrest_rate, group = group_var,
  color = crdc_race_recode(RACE)) +
    scale_color_manual(values = color_fill_values2) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0)),
                color = NULL),
              vjust = -0.25, show.legend = FALSE) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE) +
  labs(x = "", y = "NNH (Arrests)", color = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4000),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())


ggplot(race_sex_trends |> filter(RACE %in% c("AM", "WH", "HI", "BL")) |>
  mutate(group_var = paste(RACE, SEX, sep = "_"))) +
  aes(x = YEAR, y = arrest_rate, group = group_var,
  color = crdc_race_recode(RACE)) +
    scale_color_manual(values = color_fill_values2) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(round(arrest_rate, 2)),
                color = NULL),
              vjust = -0.25, show.legend = FALSE) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE) +
  labs(x = "", y = "Arrests per 1,000", color = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())


```


## State Trends

```{r}
#| label: state-trend-comparison

state_trend_summary <- combined_data |>
  filter(LEA_STATE %in% c("KS", "CA")) |>
  group_by(LEA_STATE, YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate =  total_arrests / (total_enrollment / 1000)) |>
    mutate(arrest_nnh = total_enrollment / total_arrests)

ggplot(state_trend_summary, aes(x = YEAR, y = total_arrests, group = LEA_STATE)) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(total_arrests)), nudge_y = -200) +
  geom_text(data = state_trend_summary |> filter(YEAR == "21-22"),
            aes(label = LEA_STATE), nudge_x = 0.25) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4000),
                    labels = scales::comma) +
    labs(x = "CRDC Wave", y = "Total arrests", title = "State-level total arrest trend comparison") +
    civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE)



```
## LEA Comparison

If we aggregate to the LEA level, what patterns do we see?

```{r}
#| label: lea-panel-balance

combined_data |>
    filter(total_enroll > 0) |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEAID) |>
    summarize(
          nyears = n_distinct(YEAR),
              nnames = n_distinct(LEA_NAME)
    ) -> lea_samp_summary

table(lea_samp_summary$nyears) |> prop.table()
nrow(lea_samp_summary)



combined_data |>
    filter(total_enroll > 0) |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEAID) |>
    mutate(
          nyears = n_distinct(YEAR)
    ) |>
      group_by(nyears) |>
      summarize(
                obs = n_distinct(LEAID),
                enrollment = sum(stu_enroll),
                ARRESTS = sum(ARRESTS)) -> lea_samp_summary


(lea_samp_summary$enrollment / lea_samp_summary$nyears) / sum(lea_samp_summary$enrollment / lea_samp_summary$nyears)


```


### Rankings over time

```{r}
#| label: lea-arrest-count-leaderboard

combined_data |>
#  filter(total_enroll > 0) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  group_by(YEAR, LEAID, LEA_NAME, LEA_STATE) |>
  summarize(enroll = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS, na.rm = TRUE)) |>
  group_by(YEAR) |>
  mutate(arrest_rank = rank(arrests)) |>
  #arrange(desc(arrest_rank), YEAR) |>
  tidyr::pivot_wider(
        values_from = c(enroll, arrests, arrest_rank),
                    names_from = YEAR) |>
  ungroup() |>
  filter(!is.na(`arrests_21-22`)) |>
  arrange(desc(`arrests_21-22`)) |>
  #mutate(LEA_NAME = stringr::str_trunc(LEA_NAME, width = 24)) |>
  mutate(arrest_rate_2122 = round(`arrests_21-22` / (`enroll_21-22` / 1000), 2)) |>
  select(LEA_STATE, LEA_NAME, `enroll_21-22`, `arrests_21-22`,
          arrest_rate_2122,  `arrests_17-18`, `arrests_15-16`) -> arrest_rank_leaderboard


  write.csv(arrest_rank_leaderboard, file = "export/arrest_rank_leader_table.csv")



  arrest_rank_leaderboard |>
    rename(State = LEA_STATE,
           `LEA Name` = LEA_NAME,
           `Enrollment (21-22)` = `enroll_21-22`,
           `Arrests (21-22)` = `arrests_21-22`,
           `Arrest Rate (21-22)` = `arrest_rate_2122`,
           `Arrests (17-18)` = `arrests_17-18`,
           `Arrests (15-16)` = `arrests_15-16`
           ) |>
     head(20) |> knitr::kable()


```

### Rankings of arrest rates

```{r}
#| label: lea-arrest-rate-leaderboard

combined_data |>
#  filter(total_enroll > 0) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  group_by(YEAR, LEAID, LEA_NAME, LEA_STATE) |>
  summarize(enroll = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS, na.rm = TRUE),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = round(arrests / (enroll / 1000), 1)) |>
  group_by(YEAR) |>
  #arrange(desc(arrest_rank), YEAR) |>
  tidyr::pivot_wider(
        values_from = c(enroll, arrests, arrest_rate, jj_count),
                    names_from = YEAR) |>
  ungroup() |>
  filter(!is.na(`arrests_21-22`)) |>
  arrange(desc(`arrest_rate_21-22`)) |>
  #mutate(LEA_NAME = stringr::str_trunc(LEA_NAME, width = 24)) |>
  select(LEA_STATE, LEA_NAME, `jj_count_21-22`,
        `arrest_rate_21-22`, `arrest_rate_17-18`,
        `arrest_rate_15-16`,
        `enroll_21-22`, `arrests_21-22`,
        `arrests_17-18`,
        `arrests_15-16`) -> arrest_rate_rank_leaderboard


  arrest_rate_rank_leaderboard |>
    select(-`jj_count_21-22`) |>
    rename(State = LEA_STATE,
           `LEA Name` = LEA_NAME,
           `Enrollment (21-22)` = `enroll_21-22`,
           `Arrests (21-22)` = `arrests_21-22`,
           `Arrest Rate (21-22)` = `arrest_rate_21-22`,
           `Arrest Rate (17-18)` = `arrest_rate_17-18`,
           `Arrest Rate (15-16)` = `arrest_rate_15-16`,
           `Arrests (17-18)` = `arrests_17-18`,
           `Arrests (15-16)` = `arrests_15-16`
           ) |>
     head(20) |> knitr::kable()

```


```{r}
#| label: badcells
#| eval: false
#| include: false
# Check waterbury
combined_data |> filter(LEAID == "0634320") |> View()
old_data <- read_stage_df("stages/crdc/model_data_y1516.parquet")
old_data |> filter(LEAID == "0634320") |> View()

old_data |> filter(LEA_STATE == "CA") |>
  filter(SEX == "F", RACE == "BL") |>
  group_by(LEAID, LEA_NAME) |>
  summarize(stu_enroll = sum(stu_enroll),
            nschls = n(),
            arrests = sum(ARRESTS)) |>
  filter(stu_enroll > 2000)
```

```{r}
#| label: eda-free-memory
#| include: false
# The combined three-year frame and state shapes are only used by Part 1; free
# them before the heavier draws queries in Parts 3-4.
rm(combined_data, state_map); gc()
```
