---
title: "Supplementary Materials: Equity Analysis at a Large Scale (CRDC School Arrests)"
author:
  - name: "Jared E. Knowles"
    affiliation: "Civilytics Consulting"
  - name: "Hannah Miller"
    affiliation: "Civilytics Consulting"
format:
  html:
    theme: theme/civilytics.scss
    css: theme/extras.css
    toc: true
    toc-location: right
    toc-depth: 3
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
  error: false
knitr:
  opts_chunk:
    fig.path: export/figures/supplement-
---

This document collects the supplementary materials for the white paper *Equity
Analysis at a Large Scale: Using Small Area Estimation to Get the Most from the
CRDC School Arrest Data*. It has four parts: exploratory analysis of the combined
three-year CRDC data, the data-construction and sample-restriction details,
model performance and diagnostics, and additional applied prediction-interval
examples. The headline tables and figures, and the national descriptive totals
by race and sex, are reported in the paper itself (`white_paper.html`); this
companion holds the supporting detail. Like the paper, every section reads the
published artifacts via `crdc_path()` and renders standalone.

```{r}
#| label: setup
#| include: false

# Libraries (union of the three merged source documents)
library(targets)
library(brms)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggdist)
library(marginaleffects)
library(tidybayes)
library(ggridges)
library(patchwork)
library(duckdb)
library(knitr)

options(scipen = 9)
options(width = 120)
if (!dir.exists("export/figures")) dir.create("export/figures", recursive = TRUE)

# --- Artifact-access layer (documented reproduction interface; kept sourced) ---
source("R/funs.R")
source("R/crdc_path.R"); source("R/model_registry.R"); source("R/paper_figures.R")
cv_apply_branding()                    # civilytics theme on all figures
h <- open_draws_view(); con <- h$con   # predicted_draws view over published parquet

# --- Inputs (loaded once for all sections) ---
tydata <- read_stage_df("stages/inputs/three_year_data.parquet")
rdata  <- read_stage_df("stages/inputs/recent_data.parquet")
y2122  <- read_stage_df("stages/crdc/full_crdc_data_y2122.parquet")

# National 2021-22 arrest rate used by the District Concentration analysis (Part 2)
nat_totals <- y2122 |>
  ungroup() |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarize(
    all_lea_count = n_distinct(LEAID),
    lea_any_arr   = n_distinct(LEAID[ARRESTS > 0]),
    enrollment    = sum(stu_enroll),
    arrests       = sum(ARRESTS),
    referrals     = sum(REFERRALS))
nat_arrest_rate <- nat_totals$arrests / (nat_totals$enrollment / 1000)

# Combined three-year frame + canonical (most-recent-wave) LEA name (Part 1 EDA)
combined_data <- read_stage_df("stages/inputs/combined_model_data.parquet")
combined_data |>
  select(LEAID, LEA_NAME, YEAR) |>
  distinct_all() |>
  group_by(LEAID) |>
  mutate(last_name = ifelse(any(YEAR == "21-22"), LEA_NAME[YEAR == "21-22"],
                     ifelse(any(YEAR == "17-18"), LEA_NAME[YEAR == "17-18"],
                            LEA_NAME[YEAR == "15-16"]))) |>
  ungroup() |>
  select(LEAID, last_name) |>
  distinct_all() -> lea_names_canonical
combined_data <- left_join(combined_data |> select(-LEA_NAME),
                           lea_names_canonical |> rename(LEA_NAME = last_name),
                           by = join_by(LEAID))
rm(lea_names_canonical)

# State shapes for the EDA choropleths (Part 1)
state_map <- tigris::states(cb = TRUE, resolution = "20m") |> tigris::shift_geometry()

# --- Colour palettes (inlined for transparency) ---
color_fill_values  <- c("WH" = "#c92d0e", "BL" = "#020684", "AM" = "#ff8f43",
                        "TOTAL" = "#909090ff", "HI" = "#0791b6")
color_fill_values2 <- c("White" = "#c92d0e", "Black" = "#020684",
                        "Total" = "#909090ff", "Hispanic" = "#0791b6",
                        "Amer. Ind." = "#ff8f43")
model_palette <- c("Modeled" = "#000a9bff", "Frequentist" = "#858585bb")
group_palette <- c("WH" = "#09d2ecc2", "BL" = "#ab00a9f1", "HI" = "#00ab17f1")

# --- Model-labelling helpers (inlined for transparency; used by Parts 3-4) ---
add_model_time <- function(x) {
  y <- rep("One year", length(x))
  y[x %in% c("nat_m3_mod", "nat_m4_mod", "nat_m5_mod",
             "sg_m3_mod", "sg_m4_mod", "sg_m5_mod")] <- "Three year"
  y
}
add_model_form <- function(x) {
  y <- rep("Stratified", length(x)); y[startsWith(x, "nat")] <- "Unified"; y
}
add_model_cov <- function(x) {
  y <- rep("Baseline", length(x))
  y[x %in% c("sg_m2_mod", "nat_m4_mod", "nat_m5_mod",
             "nat_m2_mod", "sg_m4_mod", "sg_m5_mod")] <- "Covariate"
  y
}
print_model_name <- function(x, lbreak = FALSE) {
  y <- rep("Stratified", length(x)); y[startsWith(x, "nat")] <- "Unified"
  x <- gsub("_mod", "", x); x <- gsub("nat_m", "", x); x <- gsub("sg_m", "", x)
  if (lbreak) paste0(y, "\nModel:", x) else paste0(y, " Model:", x)
}

# --- Agresti-Coull frequentist interval (inlined; used by Part 4 examples) ---
agresti_coull <- function(numerator, denominator, confidence_level = 0.95) {
  adj_star <- qnorm(1 - (1 - confidence_level) / 2)
  if (numerator > 0) {
    num_star   <- numerator + adj_star
    denom_star <- denominator + (2 * adj_star)
    phat       <- num_star / denom_star
    phat_se    <- sqrt((phat / denom_star) * (1 - phat))
    ci_upper   <- (phat + (adj_star * phat_se)) * denom_star
    ci_lower   <- (phat - (adj_star * phat_se)) * denom_star
    sd         <- phat_se * denom_star
  } else {
    ci_upper <- denominator * (-log(1 - confidence_level) / denominator)
    ci_lower <- 0
    sd       <- mean(c(ci_upper, ci_lower)) / confidence_level
    phat_se  <- 0
    phat     <- mean(c(ci_upper, ci_lower))
  }
  c(ci_upper, ci_lower, sd, phat_se, phat)
}
```

# Part 1 — Exploratory Data Analysis

This part explores the combined three-year CRDC data (2015-16, 2017-18,
2021-22). Counts and rates here are pooled across all three waves unless stated
otherwise, and so differ from the single-wave 2021-22 figures in the paper.

## Basic Data Structure

```{r}
#| label: data-structure
# Basic structure and dimensions
cat("Data dimensions:", nrow(combined_data), "rows,", ncol(combined_data), "columns\n\n")

# Column names and types
str(combined_data)

# First few rows
head(combined_data) |> kable()
```

## Summary Statistics by Year

```{r}
#| label: year-summary
# Summary statistics by year
year_summary <- combined_data |>
  group_by(YEAR) |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    mean_enrollment = mean(stu_enroll, na.rm = TRUE),
    mean_arrests = mean(ARRESTS, na.rm = TRUE),
    mean_referrals = mean(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )

year_summary |> knitr::kable(digits = 2)



```

### Summary Plot of Trends

```{r}
#| label: national-arrest-trend-plots

year_summary <- combined_data |>
  group_by(YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(year_summary, aes(x = YEAR, y = total_arrests)) +
  geom_point() +
  geom_line(group = 1) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  labs(x = "School Year", y = "CRDC Arrests", title = "Arrest Trends Across CRDC Collections") +
  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE)

ggplot(year_summary, aes(x = YEAR, y = total_arrests/(total_enrollment/1000))) +
  geom_point() +
  geom_line(group = 1) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  labs(x = "School Year", y = "CRDC Arrests per 1,000", title = "Arrest Rate Trends Across CRDC Collections") +
  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE)




```

## State Variation Summing over 3 years

```{r}
#| label: state_map_arrest_rate
#| echo: false
#| warning: false
#| message: false

state_summary <- combined_data |>
  group_by(LEA_STATE) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate =  total_arrests / (total_enrollment / 1000)) |>
    mutate(arrest_nnh = total_enrollment / total_arrests)



state_map2 <- inner_join(state_map, state_summary, by = join_by(STUSPS == LEA_STATE))

ggplot(data = state_map2, aes( fill = arrest_rate)) +
  geom_sf() +
  geom_sf_label(aes(label = round(arrest_rate, 1))) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  labs(title = "Arrests per 1,000 for all students by state") +
  theme_void()

#rm(state_map2)

```



```{r}
#| label: state_map_arrestsnnh
#| echo: false
#| warning: false
#| message: false
ggplot(data = state_map2, aes( fill = arrest_nnh)) +
  geom_sf() +
  geom_sf_label(aes(label = civilytics::pretty_count(round(arrest_nnh, 0)))) +
  scale_fill_distiller(palette = "YlOrRd", direction = -1,
  breaks = c(50000, 8000, 3000, 1000, 500),
  trans = "log", labels = scales::comma) +
  labs(title = "Students per arrested student by state", fill = "NNH") +
  theme_void() +
  theme(legend.position.inside = c(0.1, 0.8), legend.position = "inside")
```

## Race and Sex Distribution

These counts and rates are **pooled across all three CRDC waves (2015-16,
2017-18, 2021-22)** and are shown here as number-needed-to-harm (NNH, students
per arrested student). They therefore differ from the single-wave 2021-22
totals and per-1,000 rates reported in `white_paper.qmd`; they are not the same
figures and should not be read as such.

```{r}
#| label: race-sex-distribution
# Distribution across race and sex categories
race_sex_summary <- combined_data |>
  filter(RACE != "TOTAL", SEX != "TOTAL") |>
  group_by(RACE, SEX) |>
  summarise(
    n_observations = n(),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate = total_arrests / (total_enrollment / 1000)) |>
  arrange(desc(total_arrests))

race_sex_summary |> knitr::kable(digits = 3)
```



```{r}
#| label: nnh_arrest_counts_3yr

ggplot(race_sex_summary |> filter(RACE %in% c("AM", "WH", "HI", "BL"))) +
  aes(x = crdc_race_recode(RACE), y = 1000 / arrest_rate,
  fill = crdc_race_recode(RACE)) +
    scale_fill_manual(values = color_fill_values2) +
  geom_col() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0))),
              vjust = -0.25) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE) +
  labs(x = "", y = "NNH (Arrests)", fill = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())

```

### Trends by group

```{r}
#| label: race-sex-arrestratetrends
race_sex_trends <- combined_data |>
  filter(RACE != "TOTAL", SEX != "TOTAL") |>
  group_by(YEAR, RACE, SEX) |>
  summarise(
    n_observations = n(),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate = total_arrests / (total_enrollment / 1000)) |>
  arrange(desc(total_arrests))



ggplot(race_sex_trends |> filter(RACE %in% c("AM", "WH", "HI", "BL")) |>
  mutate(group_var = paste(RACE, SEX, sep = "_"))) +
  aes(x = YEAR, y = 1000 / arrest_rate, group = group_var,
  color = crdc_race_recode(RACE)) +
    scale_color_manual(values = color_fill_values2) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(round(1000/arrest_rate,0)),
                color = NULL),
              vjust = -0.25, show.legend = FALSE) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE) +
  labs(x = "", y = "NNH (Arrests)", color = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4000),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())


ggplot(race_sex_trends |> filter(RACE %in% c("AM", "WH", "HI", "BL")) |>
  mutate(group_var = paste(RACE, SEX, sep = "_"))) +
  aes(x = YEAR, y = arrest_rate, group = group_var,
  color = crdc_race_recode(RACE)) +
    scale_color_manual(values = color_fill_values2) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(round(arrest_rate, 2)),
                color = NULL),
              vjust = -0.25, show.legend = FALSE) +
  facet_wrap(~SEX) +
  civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE) +
  labs(x = "", y = "Arrests per 1,000", color = "Race:") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4),
                    labels = scales::comma) +
  theme(legend.position = "bottom", axis.ticks.x = element_blank())


```


## State Trends

```{r}
#| label: state-trend-comparison

state_trend_summary <- combined_data |>
  filter(LEA_STATE %in% c("KS", "CA")) |>
  group_by(LEA_STATE, YEAR) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  summarise(
    n_observations = n(),
    n_schools = n_distinct(COMBOKEY),
    n_districts = n_distinct(LEAID),
    total_enrollment = sum(stu_enroll, na.rm = TRUE),
    total_arrests = sum(ARRESTS, na.rm = TRUE),
    total_referrals = sum(REFERRALS, na.rm = TRUE),
    .groups = "drop"
  ) |>
    mutate(arrest_rate =  total_arrests / (total_enrollment / 1000)) |>
    mutate(arrest_nnh = total_enrollment / total_arrests)

ggplot(state_trend_summary, aes(x = YEAR, y = total_arrests, group = LEA_STATE)) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = civilytics::pretty_count(total_arrests)), nudge_y = -200) +
  geom_text(data = state_trend_summary |> filter(YEAR == "21-22"),
            aes(label = LEA_STATE), nudge_x = 0.25) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                    limits = c(0, 4000),
                    labels = scales::comma) +
    labs(x = "CRDC Wave", y = "Total arrests", title = "State-level total arrest trend comparison") +
    civilytics::theme_civilytics(font_size = 14, paper_bg = FALSE)



```
## LEA Comparison

If we aggregate to the LEA level, what patterns do we see?

```{r}
#| label: lea-panel-balance

combined_data |>
    filter(total_enroll > 0) |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEAID) |>
    summarize(
          nyears = n_distinct(YEAR),
              nnames = n_distinct(LEA_NAME)
    ) -> lea_samp_summary

table(lea_samp_summary$nyears) |> prop.table()
nrow(lea_samp_summary)



combined_data |>
    filter(total_enroll > 0) |>
    filter(RACE == "TOTAL" & SEX == "TOTAL") |>
    group_by(LEAID) |>
    mutate(
          nyears = n_distinct(YEAR)
    ) |>
      group_by(nyears) |>
      summarize(
                obs = n_distinct(LEAID),
                enrollment = sum(stu_enroll),
                ARRESTS = sum(ARRESTS)) -> lea_samp_summary


(lea_samp_summary$enrollment / lea_samp_summary$nyears) / sum(lea_samp_summary$enrollment / lea_samp_summary$nyears)


```


### Rankings over time

```{r}
#| label: lea-arrest-count-leaderboard

combined_data |>
#  filter(total_enroll > 0) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  group_by(YEAR, LEAID, LEA_NAME, LEA_STATE) |>
  summarize(enroll = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS, na.rm = TRUE)) |>
  group_by(YEAR) |>
  mutate(arrest_rank = rank(arrests)) |>
  #arrange(desc(arrest_rank), YEAR) |>
  tidyr::pivot_wider(
        values_from = c(enroll, arrests, arrest_rank),
                    names_from = YEAR) |>
  ungroup() |>
  filter(!is.na(`arrests_21-22`)) |>
  arrange(desc(`arrests_21-22`)) |>
  #mutate(LEA_NAME = stringr::str_trunc(LEA_NAME, width = 24)) |>
  mutate(arrest_rate_2122 = round(`arrests_21-22` / (`enroll_21-22` / 1000), 2)) |>
  select(LEA_STATE, LEA_NAME, `enroll_21-22`, `arrests_21-22`,
          arrest_rate_2122,  `arrests_17-18`, `arrests_15-16`) -> arrest_rank_leaderboard


  write.csv(arrest_rank_leaderboard, file = "export/arrest_rank_leader_table.csv")



  arrest_rank_leaderboard |>
    rename(State = LEA_STATE,
           `LEA Name` = LEA_NAME,
           `Enrollment (21-22)` = `enroll_21-22`,
           `Arrests (21-22)` = `arrests_21-22`,
           `Arrest Rate (21-22)` = `arrest_rate_2122`,
           `Arrests (17-18)` = `arrests_17-18`,
           `Arrests (15-16)` = `arrests_15-16`
           ) |>
     head(20) |> knitr::kable()


```

### Rankings of arrest rates

```{r}
#| label: lea-arrest-rate-leaderboard

combined_data |>
#  filter(total_enroll > 0) |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  group_by(YEAR, LEAID, LEA_NAME, LEA_STATE) |>
  summarize(enroll = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS, na.rm = TRUE),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = round(arrests / (enroll / 1000), 1)) |>
  group_by(YEAR) |>
  #arrange(desc(arrest_rank), YEAR) |>
  tidyr::pivot_wider(
        values_from = c(enroll, arrests, arrest_rate, jj_count),
                    names_from = YEAR) |>
  ungroup() |>
  filter(!is.na(`arrests_21-22`)) |>
  arrange(desc(`arrest_rate_21-22`)) |>
  #mutate(LEA_NAME = stringr::str_trunc(LEA_NAME, width = 24)) |>
  select(LEA_STATE, LEA_NAME, `jj_count_21-22`,
        `arrest_rate_21-22`, `arrest_rate_17-18`,
        `arrest_rate_15-16`,
        `enroll_21-22`, `arrests_21-22`,
        `arrests_17-18`,
        `arrests_15-16`) -> arrest_rate_rank_leaderboard


  arrest_rate_rank_leaderboard |>
    select(-`jj_count_21-22`) |>
    rename(State = LEA_STATE,
           `LEA Name` = LEA_NAME,
           `Enrollment (21-22)` = `enroll_21-22`,
           `Arrests (21-22)` = `arrests_21-22`,
           `Arrest Rate (21-22)` = `arrest_rate_21-22`,
           `Arrest Rate (17-18)` = `arrest_rate_17-18`,
           `Arrest Rate (15-16)` = `arrest_rate_15-16`,
           `Arrests (17-18)` = `arrests_17-18`,
           `Arrests (15-16)` = `arrests_15-16`
           ) |>
     head(20) |> knitr::kable()

```


```{r}
#| label: badcells
#| eval: false
#| include: false
# Check waterbury
combined_data |> filter(LEAID == "0634320") |> View()
old_data <- read_stage_df("stages/crdc/model_data_y1516.parquet")
old_data |> filter(LEAID == "0634320") |> View()

old_data |> filter(LEA_STATE == "CA") |>
  filter(SEX == "F", RACE == "BL") |>
  group_by(LEAID, LEA_NAME) |>
  summarize(stu_enroll = sum(stu_enroll),
            nschls = n(),
            arrests = sum(ARRESTS)) |>
  filter(stu_enroll > 2000)
```

```{r}
#| label: eda-free-memory
#| include: false
# The combined three-year frame and state shapes are only used by Part 1; free
# them before the heavier draws queries in Parts 3-4.
rm(combined_data, state_map); gc()
```

# Part 2 — Data Construction & Sample Restrictions

This part documents how the analytic sample is built from the raw CRDC and the
magnitude of each restriction the paper describes. The national descriptive
totals and rates by race and sex are reported in the paper (`white_paper.html`).

## Sample Continuity Analysis for CRDC Waves

Look at how many LEAs are in multiple years of data and how many LEAs drop
out of the data when we add successive waves of the data.

```{r}
#| label: load-historical-data

y1718 <- read_stage_df("stages/crdc/full_crdc_data_y1718.parquet")
y1516 <- read_stage_df("stages/crdc/full_crdc_data_y1516.parquet")
```

```{r}
#| label: match-test-1718-2122

match_test(y1718$LEAID, y2122$LEAID)
```

```{r}
#| label: match-test-1516-1718

match_test(y1516$LEAID, y1718$LEAID[y1718$LEAID %in% unique(y2122$LEAID)])
```

## District Concentration Analysis

For this analysis we look at the most recent CRDC wave.

### District Size Categories

```{r}
#| label: compute-district-categories

thresh <- c(0, 1000, 10000, 20000, Inf)
labs <- c("0-999", "1,000-9,999", "10,000-19,999","20,000+")

dist_ref_arr <- y2122 %>%
  group_by(YEAR, LEA_STATE, LEAID, RACE, SEX) %>%
  summarize(ARRESTS = sum(ARRESTS),
            REFERRALS = sum(REFERRALS),
            stu_enroll = sum(stu_enroll),
            jj_count = sum(JJ == "Yes")) |>
  mutate(arrest_rate = ARRESTS / (stu_enroll / 1000),
         referral_rate = REFERRALS / (stu_enroll / 1000)) %>%
  group_by(YEAR, LEA_STATE, LEAID) |>
  mutate(total_enroll = sum(stu_enroll[RACE != "TOTAL" &
                                         SEX != "TOTAL"]))
```

### Expected vs Observed Arrests

```{r}
#| label: compute-expected-arrests

dist_size_summary <- dist_ref_arr %>%
   ungroup() %>%
     filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
   mutate(popcut = cut(stu_enroll,
                       breaks = thresh,
                       labels = labs,
                       include.lowest= TRUE)) %>%
   group_by(popcut) %>%
 mutate(expec_arrest = pbinom(1, stu_enroll, nat_arrest_rate/1000,
                              lower.tail = FALSE)) %>%
   summarize(dists = n(),
             dists_w_arrests = sum(ARRESTS>0),
             expec_distw_arrests = round(sum(expec_arrest),1)) %>%
   mutate(dist_w_arrest_per = pretty_per(dists_w_arrests/dists),
          expect_w_arrest_per = pretty_per(expec_distw_arrests/dists))
```

```{r}
#| label: display-expected-arrests

dist_size_summary |> knitr::kable()
```

### Concentration Visualization

This visualization shows the relative concentration of arrests compared in LEAs
compared to the concentration of population in LEAs.
```{r}
#| label: compute-concentration-data

total_enroll_national <- sum(dist_ref_arr$total_enroll[
  dist_ref_arr$RACE == "TOTAL" & dist_ref_arr$SEX == "TOTAL"
])

concentration_plot_data <- dist_ref_arr %>%
  ungroup %>%
  filter(ARRESTS > 0) %>%
  filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
  arrange(desc(stu_enroll)) %>%
  mutate(total_arr = sum(ARRESTS)) %>%
  mutate(cum_sum = cumsum(ARRESTS)) %>%
  mutate(cum_sum_per = cum_sum / total_arr) %>%
  mutate(dist_per = stu_enroll / total_enroll_national) %>%
  mutate(dist_per_cum = cumsum(dist_per))
```

```{r}
#| label: plot-concentration
#| fig-cap: "Share of total arrests by cumulative district enrollment"
#| fig-width: 10
#| fig-height: 6

concentration_plot <- concentration_plot_data %>%
  ggplot(aes(x = dist_per_cum, y = cum_sum_per)) +
  geom_line() +
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5, linetype = 2) +
  scale_x_continuous("% of national student population",
                     labels = scales::percent, limits = c(0, 0.7)) +
  scale_y_continuous("% of all arrests", labels = scales::percent) +
  civilytics::theme_civilytics(font_size = 12, paper_bg = FALSE) +
  labs(title = "Share of total arrests by cumulative district enrollment",
       subtitle = "Horizontal line at 50% of all arrests, vertical line at 50% of all students") +
  theme(axis.ticks.x = element_blank(),
        panel.grid = element_blank())

print(concentration_plot)

# Save high-resolution figure
ggsave("export/figures/arrest_concentration_by_district_enrollment.png",
       plot = concentration_plot,
       width = 10, height = 6, dpi = 300, bg = "white")
```

### Enrollment by Arrest Status

```{r}
#| label: compute-enrollment-by-arrest

enrollment_by_arrest <- dist_ref_arr %>%
  ungroup %>%
  filter(RACE == "TOTAL" & SEX == "TOTAL") %>%
  mutate(has_arrest = ifelse(ARRESTS > 0, "Arrests", "No arrests")) |>
  group_by(has_arrest) |>
  summarize(enrollment = sum(total_enroll)) |>
  mutate(enrollment_per = enrollment / total_enroll_national)

enrollment_by_arrest |> knitr::kable()
```


## Sample Restriction Impact

Here we tabulate the figures about sample restrictions in the paper.

### Section 504 Data

```{r}
#| label: compute-504-data

sch_ref <- read_stage_df("stages/crdc/lerefs_y2122.parquet")

section_504_summary <- sch_ref %>%
  select(!matches("TOT_")) %>%
  select(matches("_504_")) |>
  ungroup() |>
  summarize_all(~ sum(crdc_sub(.x), na.rm = TRUE))
```

```{r}
#| label: display-504-data

section_504_summary |> knitr::kable()
```

### Missing Data Patterns

Law enforcement referral missingness data.

```{r}
#| label: compute-missing-patterns

# Drop non COMBOKEY identifier columns
sch_ref <- sch_ref %>% select(COMBOKEY:last_col())
names(sch_ref) <- gsub("_IDEA_", "_", names(sch_ref))

# Check for reserve values
missing_pattern_table <- sch_ref %>%
  select(!matches("TOT_")) %>%
  select(!matches("_504_")) %>%
  select(!matches("_LEP_")) %>%
  select(!matches("_EL_")) %>%
  tidyr::pivot_longer(cols = matches("SCH_DIS"), names_prefix = "SCH_",
              names_to = c("DISAB", "REF_TYPE", "RACE", "SEX"),
              names_sep = "_") %>%
  group_by(COMBOKEY, DISAB, REF_TYPE, RACE, SEX) |>
  mutate(value_mod = ifelse(value < 0, TRUE, FALSE)) |>
  pull(value_mod) |>
  table()
```

```{r}
#| label: display-missing-patterns

missing_pattern_table
```

```{r}
#| label: display-missing-proportions

prop.table(missing_pattern_table)
```

```{r}
#| label: cleanup-sch-ref

rm(sch_ref)
```

### Section 504 Enrollment

```{r}
#| label: compute-504-enrollment

enrollment <- read_stage_df("stages/crdc/popcounts_y2122.parquet")

enrollment_504_summary <- enrollment %>%
  select(!matches("TOT_")) %>%
  select(matches("_504")) |>
  select(1:2) |>
  summarize_all(~ sum(crdc_sub(.x), na.rm = TRUE))
```

```{r}
#| label: display-504-enrollment

enrollment_504_summary |> knitr::kable()
```

```{r}
#| label: cleanup-enrollment

rm(enrollment)
```

### Grade Level Restrictions

This reports the impact of restricting the analysis to schools that offer
a grade 7 or above.

```{r}
#| label: compute-grade-restrictions

gr_data <- read_stage_df("stages/inputs/combined_model_data.parquet") |> filter(YEAR == "21-22")
gr_data2 <- gr_data |> filter(highest_grade_offered >= 7)

grade_restriction_summary <- gr_data |>
  filter(YEAR == "21-22") |>
  filter(RACE == "TOTAL") |>
  filter(SEX == "TOTAL") |>
  mutate(hs = ifelse(highest_grade_offered >= 7, "Yes", "No")) |>
  group_by(hs) |>
  summarize(enrollment = sum(enrollment, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-grade-restrictions

grade_restriction_summary |> knitr::kable()
```

### CCD Matching

By matching to CCD data we further restrict the CRDC universe. This analysis
shows the impact of that.

```{r}
#| label: compute-ccd-matching

ccd <- read_stage_df("stages/crdc/ccd_sch_geo_y2122.parquet")
crdc <- read_stage_df("stages/crdc/full_crdc_data_y2122.parquet")

ccd$COMBOKEY <- stringr::str_pad(ccd$ncessch_num, width = 12,
                                side = "left", pad = "0")
ccd <- ccd |> select(COMBOKEY, highest_grade_offered, lowest_grade_offered,
                     latitude, longitude, enrollment)
```

```{r}
#| label: display-ccd-match-test

match_test(crdc$COMBOKEY, ccd$COMBOKEY)
```

```{r}
#| label: compute-ccd-match-summary

ccd_match_summary <- crdc |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  ungroup() |>
  mutate(in_ccd = ifelse(COMBOKEY %in% unique(ccd$COMBOKEY), "Yes", "No")) |>
  group_by(in_ccd) |>
  summarize(enrollment = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-ccd-match-summary

ccd_match_summary |> knitr::kable()
```

```{r}
#| label: cleanup-ccd-crdc

rm(ccd, crdc)
gc()
```

### Enrollment Thresholds

We restrict the sample by enrollment size, and this reports the impact of that
restriction.

```{r}
#| label: compute-enrollment-thresholds

thr_data <- read_stage_df("stages/inputs/combined_model_data.parquet") |> filter(YEAR == "21-22")
thr_data <- crdc_lea_collapse(thr_data)

enrollment_threshold_summary <- thr_data |>
  filter(RACE == "TOTAL" & SEX == "TOTAL") |>
  mutate(filtered_out = ifelse(total_enroll < 30 | stu_enroll == 0, "Yes", "No")) |>
  group_by(filtered_out) |>
  summarize(dists = n(),
            enrollment = sum(stu_enroll, na.rm = TRUE),
            arrests = sum(ARRESTS),
            referrals = sum(REFERRALS))
```

```{r}
#| label: display-enrollment-thresholds

enrollment_threshold_summary |> knitr::kable()
```

### Censored or truncated enrollment data

Here we report the impact of suppressed records on enrollment.


```{r}
#| label: compute-incomplete-enrollments

enrollment <- read_stage_df("stages/crdc/popcounts_y2122.parquet")
zz <- incomplete_enrollments(enrollment, year = "21-22")
zb <- read_stage_df("stages/crdc/schenrollraw_y2122.parquet")

incomplete_proportion <- nrow(zz) / sum(nrow(zz), nrow(zb))
incomplete_proportion
```
